




Notebook to test the preservedCoexp R package and add some plots to main figure 6. Showing how long it takes to generate co-expression matrices as the number of cells increases
Showing how long it takes to compute preserved co-expression scores of genes as the number of genes increases
Potentially run on my macbook pro and compare the times, our lab servers are pretty memory heavy, might not be the case in every lab


```{r}
library(Seurat)
library(preservedCoexp)
library(dplyr)
library(parallel)
library(ggplot2)

```

Use the same organoid dataset to test as in Figure 6, organoid data 110
```{r}

load('../data/seurat_objs/all_data_just_meta_seurat_with_addl_features.Rdata')
all_sample_meta
```
```{r}

data = load(all_sample_meta$processed_seurat_path[110])
data = get(data)
rm(dataset)
gc()
data
```



```{r}
library(Matrix)
exp_data = Matrix(data@assays$RNA@data, sparse = T)

dim(exp_data)
```

Save just the expression matrix for testing on other machines
```{r}
save(exp_data, file = 'r_package_data/test_exp_matrix.Rdata')

```




Fit to Go and get co-expresssion network
```{r}
data('go_genes', package = 'preservedCoexp')

exp_data_GO = fit_to_GO(exp_data)              #Fit to GO gene annotations
rank_mat = get_spearman(exp_data_GO)   #Get co-expression matrix
rank_mat = rank_coexpression(rank_mat)  #Get rank standardized co-expression matrix

```


```{r}
data('meta_presCoexp_df', package = 'preservedCoexp')
data('fetal_meta_markers', package = 'preservedCoexp')
aggregated_fetal_network = load_fetal_coexp() 


plot_results = plot_meta_results(aggregated_fetal_network, rank_mat, fetal_meta_markers, meta_presCoexp_df, parallel = T)
plot_results[[1]]
plot_results[[2]]
plot_results[[3]]
```


Get the time it takes to generate co-expression networks for 10k cells and then increase cell number to 100k

Make a matrix with 120k cells
```{r}
ncol(exp_data)

exp_data = cbind(exp_data, exp_data[ ,1:7043])
ncol(exp_data)

exp_data = cbind(exp_data[ ,1:20000], cbind(exp_data, exp_data))
dim(exp_data)
```

```{r}

rm(data,aggregated_fetal_network, exp_data_GO, rank_mat )
gc()
```




```{r}
num_samples = 10
sample_vec = seq(10000, 100000, 10000)
time_mat = matrix(data=NA, nrow = num_samples, ncol = 10,dimnames = list(paste('Sample', 1:num_samples), paste0(seq(10,100,10), 'k cells')) ) 

for(j in 1:length(sample_vec)){
  start_top = Sys.time()
  for(i in 1:num_samples){
    #Random sample some cells
    rand_index = sample(1:ncol(exp_data), sample_vec[j])
    exp_data_temp = exp_data[ ,rand_index]
    #Make a co-expression network
    start = proc.time()
    exp_data_temp = fit_to_GO(exp_data_temp)      
    rank_mat = get_spearman(exp_data_temp)   
    rank_mat = rank_coexpression(rank_mat)
    time_took = proc.time() - start
    time_mat[i, j] = time_took[['elapsed']]                #Time in seconds
    print(sprintf('Done with %ith sample of %i cells...', i, sample_vec[j]))
    print(Sys.time() - start_top)
    rm(exp_data_temp, rank_mat)
    gc()
  }
  
}


time_mat

```


Plot cells on the x, time on the y, error bars as SD
Also run this on my Mac, probably won't be able to go up to 100k cells
include in plot
```{r}

colMeans(time_mat)/60



```

```{r}

save(time_mat, file = '../data/package_time/coexp_time_sample_mats.Rdata')

```




Get the time it takes to compute preserved co_expression for sets of genes, try using the 100k dataset

```{r}
#Get a 100k co-expression matrix
exp_data_temp = fit_to_GO(exp_data)      
rank_mat = get_spearman(exp_data_temp)   
rank_mat = rank_coexpression(rank_mat)

rm(exp_data_temp)
gc()
```


Load up the aggregate fetal network again
```{r}
aggregated_fetal_network = load_fetal_coexp()

```


```{r}


num_samples = 10
sample_vec = seq(1000, 20000 , 1000)
time_mat_genes = matrix(data=NA, nrow = num_samples, ncol = length(sample_vec),dimnames = list(paste('Sample', 1:num_samples), paste(sample_vec, 'genes')) ) 

for(j in 1:length(sample_vec)){

  for(i in 1:num_samples){
    
    gene_index = sample(1:length(go_genes), sample_vec[j])
    start = proc.time()
    gene_aurocs = get_geneset_preserved_coexp(aggregated_fetal_network, rank_mat, go_genes[gene_index], parallel = T)
    time_took = proc.time() - start
    time_mat_genes[i, j ] = time_took[['elapsed']]
  
    
  }
  
  print(sprintf('Done with sampling %i genes...', sample_vec[j]))
  
}


time_mat_genes
```


```{r}


colMeans(time_mat_genes)

```


```{r}

save(time_mat_genes, file = '../data/package_time/presCoexp_time_sample_mats.Rdata')

```


```{r}

rugen_time = load(file = '../data/package_time/coexp_time_sample_mats.Rdata')
rugen_time = get(rugen_time)
rugen_time


mac_time = load(file = '../data/package_time/mac_time_mat.Rdata')
mac_time = get(mac_time)
mac_time

colnames(mac_time) = c('10k cells','15k cells','20k cells','25k cells','30k cells','35k cells','40k cells')

```




Plot using SD as error bars

```{r}

rugen_time_sd = apply(rugen_time/60, 2, sd)
rugen_time_mean = colMeans(rugen_time/60)

mac_time_sd = apply(mac_time/60, 2, sd)
mac_time_mean = colMeans(mac_time/60)

label_vec = c(rep('rugen', length = length(rugen_time_mean)), rep('macBook', length = length(mac_time_mean)))
time_mat_df = data.frame(num_cells = c(names(rugen_time_mean), names(mac_time_mean)), mean_time_mins = c(rugen_time_mean, mac_time_mean), sd_mins = c(rugen_time_sd, mac_time_sd),
                         label = label_vec)
time_mat_df$num_cells = factor(time_mat_df$num_cells, levels = c("10k cells","15k cells","20k cells", "25k cells","30k cells", "35k cells","40k cells","50k cells","60k cells",
                                                                 "70k cells","80k cells","90k cells","100k cells"))
time_mat_df


ggplot(time_mat_df, aes(x = num_cells, y = mean_time_mins, group = label, color = label)) + geom_line(linetype = 'dashed') + geom_point(size = 2.5, color = 'black') +
  geom_errorbar(aes(ymin=mean_time_mins-sd_mins, ymax=mean_time_mins+sd_mins), width=.75) +
  scale_color_manual(values = c('macBook' = 'blue', 'rugen' = 'red')) +
  scale_x_discrete(guide = guide_axis(n.dodge=2)) +
  xlab('Number of cells') + ylab('Time (minutes)') + ggtitle('Computing co-expression networks')
ggsave(filename = 'make_coexp_nets_cells_by_time.pdf', path = 'graphs/package_time/', device = 'pdf', useDingbats = F, height = 4, width = 6)


```

```{r}

rugen_genes = load(file = '../data/package_time/presCoexp_time_sample_mats.Rdata')
rugen_genes = get(rugen_genes)
rugen_genes
colnames(rugen_genes) = seq(1000, 20000, 1000)


mac_genes = load(file = '../data/package_time/mac_time_genes_mat.Rdata')
mac_genes = get(mac_genes)
mac_genes
colnames(mac_genes) = seq(1000, 20000, 1000)
```





```{r}

rugen_mat_genes_sd = apply(rugen_genes/60, 2, sd)
rugen_mat_genes_mean = colMeans(rugen_genes/60)

mac_mat_genes_sd = apply(mac_genes/60, 2, sd)
mac_mat_genes_mean = colMeans(mac_genes/60)

label_vec = c(rep('rugen', length = length(rugen_mat_genes_mean)), rep('mac', length = length(mac_mat_genes_mean)))

time_mat_genes_df = data.frame(num_genes = c(names(rugen_mat_genes_mean), names(mac_mat_genes_mean)), mean_time_mins = c(rugen_mat_genes_mean, mac_mat_genes_mean), 
                               sd_mins = c(rugen_mat_genes_sd, mac_mat_genes_sd), label = label_vec)
time_mat_genes_df$num_genes = factor(time_mat_genes_df$num_genes, levels = unique(time_mat_genes_df$num_genes))

time_mat_genes_df


ggplot(time_mat_genes_df, aes(x = num_genes, y = mean_time_mins, group = label, color = label)) + geom_line(linetype = 'dashed') + geom_point(size = 2.5, color = 'black') +
  geom_errorbar(aes(ymin=mean_time_mins-sd_mins, ymax=mean_time_mins+sd_mins), width=.75) +
  scale_color_manual(values = c('mac' = 'blue', 'rugen' = 'red')) +
  scale_x_discrete(guide = guide_axis(n.dodge=2)) +
  xlab('Number of genes') + ylab('Time (minutes)') + ggtitle('Computing preserved co-expression')
ggsave(filename = 'compute_presCoexp_genes_by_time.pdf', path = 'graphs/package_time/', device = 'pdf', useDingbats = F, height = 4, width = 6)



```








