

Testing the adult brain coexpression  (MTG human data; https://www.biorxiv.org/content/10.1101/2022.09.19.508480v1.full.pdf)


Get adult metamarkers and an adult coexpression network

Test the conserved coexpression of the fetal and organoid neurons and non-neurons

Compare against the fetal results

Expecting performance to drop compared to the fetal, but if the results are highly correlated, suggests that an adult reference can still be used to score organoids
Useful for organoid models that do not have a lot of exisiting comparable fetal data
Means we could also theoretically use all the exisiting adult data

```{r, setup, include=FALSE}
knitr::opts_knit$set(root.dir = '/home/werner/projects/meta_qc_organoid/code')
```

```{r}
library(Seurat)
library(data.table)
library(ggplot2)
library(MetaMarkers)
library(dplyr)
library(MetBrewer)
library(ComplexHeatmap)
library(reticulate)
Sys.setenv(RETICULATE_PYTHON = "/usr/bin/python3.6")
library(reticulate)
np <- reticulate::import("numpy")
library(parallel)
```

Now filter the networks to just include the go genes, same gene universe across all datasets/coexpression networks

```{r}
#library(org.Hs.eg.db)
#library(EGAD)
#library(AnnotationDbi)
load('../../../data/go_annotations/go_annotations.Rdata')
```


```{r}
human_class_path = '/data/fischer/biccn_markers/human/class_markers_top1000.csv.gz'
human_subclass_path = '/data/fischer/biccn_markers/human/subclass_markers_top1000.csv.gz'
human_type_path = '/data/fischer/biccn_markers/human/type_markers_top1000.csv.gz'

human_class_markers = fread(human_class_path)

human_subclass_markers = fread(human_subclass_path)

human_type_markers = fread(human_type_path)
```

```{r}
get_celltype_ranked_markers = function(metamarkers, celltype, num_markers, metric = 'rank'){
  
  if(metric != 'rank'){
    celltype_data = filter(metamarkers, cell_type == celltype) %>% arrange(desc(!!as.name(metric)))
    return(celltype_data[1:num_markers, ])}
  else{
    return(filter(metamarkers, cell_type == celltype & rank <= num_markers))
  }
}
```


```{r}

table(human_class_markers$cell_type)
filter(human_class_markers, rank <= 100 & cell_type == 'GABAergic')
filter(human_class_markers, rank <= 100 & cell_type == 'Glutamatergic')
filter(human_class_markers, rank <= 100 & cell_type == 'Non-neuronal')

human_subclass_markers %>% filter(rank <= 10)

adult_gaba = get_celltype_ranked_markers(human_class_markers, 'GABAergic', 100)$gene
adult_glut = get_celltype_ranked_markers(human_class_markers, 'Glutamatergic', 100)$gene
adult_nonN = get_celltype_ranked_markers(human_class_markers, 'Non-neuronal', 100)$gene

#Trying astro and oliugo specific markers for adult
adult_astro = get_celltype_ranked_markers(human_subclass_markers, 'Astro', 50)$gene
adult_oligo = get_celltype_ranked_markers(human_subclass_markers, 'Oligo', 50)$gene
adult_astro_oligo = c(adult_astro, adult_oligo)

adult_gaba_go = adult_gaba[adult_gaba %in% go_genes]
adult_glut_go = adult_glut[adult_glut %in% go_genes]
adult_nonN_go = adult_nonN[adult_nonN %in% go_genes]
adult_astro_oligo_go = adult_astro_oligo[adult_astro_oligo %in% go_genes]

length(adult_gaba_go)
length(adult_glut_go)
length(adult_nonN_go)
length(adult_astro_oligo)
```


load in the adult data


```{r}


load('../data/human_biccn/human_biccn_expression.Rdata')
human_exp_data
```


Exclude the Endo, VLMC, Micro-PVM cell types from the Non-neuronal class

```{r}

human_exp_data$class[human_exp_data$subclass %in% c('Endo','Micro-PVM','VLMC')] = 'other'
```



```{r}

table(human_exp_data$subclass[human_exp_data$class == 'Non-neuronal'])

human_exp_data[[]]

table(human_exp_data$orig.ident)

median(human_exp_data$nFeature_RNA)

table(human_exp_data$nFeature_RNA >= 200 & human_exp_data$nFeature_RNA <= 6000)

human_exp_data[["percent.mt"]] <- PercentageFeatureSet(human_exp_data, pattern = "^MT-")

median(human_exp_data$percent.mt)

human_exp_data@assays$RNA@meta.features

```


```{r}

hist(human_exp_data$nCount_RNA)
hist(human_exp_data$nFeature_RNA)

```


Make a coexpression network

```{r}
source_python('python_coexpression.py')
```




```{r}
start = Sys.time()

dataset_genes = rownames(human_exp_data)
#Filter for the GO gene universe
dataset_genes_in_go = dataset_genes[dataset_genes %in% go_genes]
dataset_genes_missing_from_go = go_genes[! go_genes %in% dataset_genes]
#Subset the gene expression for the genes present in GO
dataset_gene_index = dataset_genes %in% dataset_genes_in_go

exp_data = as.matrix(human_exp_data@assays$RNA@data[dataset_gene_index, ])

#Add zero counts for the missing GO genes
go_missing = matrix(0, nrow = length(dataset_genes_missing_from_go), ncol = ncol(exp_data), 
                    dimnames = list(dataset_genes_missing_from_go, colnames(exp_data)))
exp_data = rbind(exp_data, go_missing)
r_genes = rownames(exp_data)

#Free up memory
rm(human_exp_data)
gc()




#Get coexpression data
test_py_exp = np$array(exp_data) #Turn R matrix into numpy array

rm(exp_data, go_missing)
gc()


python_script_test = get_coexpression(test_py_exp, r_genes) #Python method for spearman correlations, coexpression
corr_matrix = as.matrix(python_script_test) #Turn to matrix
#Convert Nans to 0, these are genes with 0 expression in any cells
corr_matrix[!is.finite(corr_matrix)] = 0
  
file_name = '../data/human_biccn/adult_brain_coexp.Rdata'
save(corr_matrix, file = file_name)
end = Sys.time()
print(end - start)
print(sprintf('Done getting coexpression from %i dataset...', i))


#Rank the coexpression data
#Rank everything, I want the minimum rank to be 1, so ties are minimum. Subtracting 1 sets the minimum rank to 0
rank_corr = frank(c(corr_matrix), ties.method = 'min') - 1
#Divide by max rank squeezes everything between 0 and 1
norm_rank_corr = rank_corr / max(rank_corr)
#Put back into a matrix
rank_corr_matrix = matrix(norm_rank_corr, nrow = nrow(corr_matrix), ncol = ncol(corr_matrix), 
                          dimnames = list(rownames(corr_matrix), colnames(corr_matrix)))
diag(rank_corr_matrix) = 1

file_name = '../data/human_biccn/ranked_adult_brain_coexp.Rdata'
save(rank_corr_matrix, file =file_name)

```





```{r}
load('../../../data/human_biccn/ranked_adult_brain_coexp.Rdata')
adult_rank_corr_mat = rank_corr_matrix
dim(adult_rank_corr_mat)
```


Test the conservation of coexpression using the adult ranked network as the reference

```{r}
compute_auc = function(scores, label){
  label = as.logical(label)
  n1 = as.numeric(sum(label))
  n2 = as.numeric(sum(!label))
  R1 = sum(rank(scores)[label])
  U1 = R1 - n1 * (n1 + 1)/2
  auc = U1/(n1 * n2)
  return(auc)
}


#Compute ROC curve for plotting
#Input is a list of true/false labels (celltype identity) ranked by some metric (gene expression). 
roc_curve = function(ranked_labels){

  #Get total number of positives and negatives
  num_pos = sum(ranked_labels)
  num_neg = sum(!ranked_labels)
  num_total = length(ranked_labels)
  #Cumulative sum of positives and negatives
  tp = cumsum(ranked_labels)
  fp = cumsum(!ranked_labels)
  
  fpr = fp/num_neg
  tpr = tp/num_pos
  
  return(list('fpr' = fpr, 'tpr' = tpr))
}

get_conserved_coexp = function(network1, network2, test_gene, num_top = 10){

  #Get the top 10 coexpressed genes
  test_gene_coexp = network1[test_gene, ]
  test_gene_coexp = test_gene_coexp[names(test_gene_coexp) != test_gene]
  top_coexp = names(test_gene_coexp[order(test_gene_coexp, decreasing = T)][1:num_top])
  
  #Calculate the AUC using the organoid data
  scores = network2[test_gene, ]
  scores = scores[names(scores) != test_gene]
  labels = names(scores) %in% top_coexp
  
  #Get the auc
  auc = compute_auc(scores, labels)
  return(auc)
}

```


organoid dataset metadata

```{r}

all_sample_meta = read.csv('../../../data/seurat_objs/all_data_just_meta_seurat_with_addl_features_v2.csv')
all_sample_meta
```

```{r}

top_genes = 10

glut_cons_matrix_1 = matrix(nrow = length(adult_glut_go), ncol = nrow(all_sample_meta), dimnames = list(adult_glut_go, all_sample_meta$sample_ids))
gaba_cons_matrix_1 = matrix(nrow = length(adult_gaba_go), ncol = nrow(all_sample_meta), dimnames = list(adult_gaba_go, all_sample_meta$sample_ids))
nonN_cons_matrix_1 = matrix(nrow = length(adult_nonN_go), ncol = nrow(all_sample_meta), dimnames = list(adult_nonN_go, all_sample_meta$sample_ids))
astro_oligo_cons_matrix_1 = matrix(nrow = length(adult_astro_oligo_go), ncol = nrow(all_sample_meta), dimnames = list(adult_astro_oligo_go, all_sample_meta$sample_ids))

for(i in 1:nrow(all_sample_meta)){
  load(sprintf('../../../data/coexpression/organoid/6_26_24_GO_coexp/ranked_%s_coexp.Rdata', all_sample_meta$sample_ids[i]))
  

  glut_cons_matrix_1[ ,i] = sapply(1:length(adult_glut_go), function(i) get_conserved_coexp(adult_rank_corr_mat, rank_corr_matrix, adult_glut_go[i],top_genes ) )
  gaba_cons_matrix_1[ ,i] = sapply(1:length(adult_gaba_go), function(i) get_conserved_coexp(adult_rank_corr_mat, rank_corr_matrix, adult_gaba_go[i],top_genes ) )
  nonN_cons_matrix_1[ ,i] = sapply(1:length(adult_nonN_go), function(i) get_conserved_coexp(adult_rank_corr_mat, rank_corr_matrix, adult_nonN_go[i],top_genes ) )
  astro_oligo_cons_matrix_1[ ,i] = sapply(1:length(adult_astro_oligo_go), function(i) get_conserved_coexp(adult_rank_corr_mat, rank_corr_matrix, adult_astro_oligo_go[i],top_genes ) )
  print(i)

}


```


```{r}

glut_cons_matrix_1


```




```{r}

glut_cons_matrix = reshape2::melt(t(glut_cons_matrix_1))
mean_order_df = glut_cons_matrix %>% group_by(Var2) %>% summarize(median = median(value)) %>% arrange(desc(median))
glut_cons_matrix$Var2  = factor(glut_cons_matrix$Var2, levels = c(as.character(mean_order_df$Var2)))
ggplot(glut_cons_matrix, aes(x = Var2, y = value)) + geom_boxplot(outlier.size = .5, outlier.alpha = .5) +
  ylab('Conservation of Adult coexpression (AUC)') + xlab('Top 100 Adult Glutamatergic markers') + 
  theme(axis.text.x = element_blank()) 

gaba_cons_matrix = reshape2::melt(t(gaba_cons_matrix_1))
mean_order_df = gaba_cons_matrix %>% group_by(Var2) %>% summarize(median = median(value)) %>% arrange(desc(median))
gaba_cons_matrix$Var2  = factor(gaba_cons_matrix$Var2, levels = c(as.character(mean_order_df$Var2)))
ggplot(gaba_cons_matrix, aes(x = Var2, y = value)) + geom_boxplot(outlier.size = .5, outlier.alpha = .5) +
  ylab('Conservation of Adult coexpression (AUC)') + xlab('Top 100 Adult GABAergic markers') + 
  theme(axis.text.x = element_blank()) 

nonN_cons_matrix = reshape2::melt(t(nonN_cons_matrix_1))
mean_order_df = nonN_cons_matrix %>% group_by(Var2) %>% summarize(median = median(value)) %>% arrange(desc(median))
nonN_cons_matrix$Var2  = factor(nonN_cons_matrix$Var2, levels = c(as.character(mean_order_df$Var2)))
ggplot(nonN_cons_matrix, aes(x = Var2, y = value)) + geom_boxplot(outlier.size = .5, outlier.alpha = .5) +
  ylab('Conservation of Adult coexpression (AUC)') + xlab('Top 100 Adult Non-neuronal markers') + 
  theme(axis.text.x = element_blank()) 

astro_oligo_cons_matrix = reshape2::melt(t(astro_oligo_cons_matrix_1))
mean_order_df = astro_oligo_cons_matrix %>% group_by(Var2) %>% summarize(median = median(value)) %>% arrange(desc(median))
astro_oligo_cons_matrix$Var2  = factor(astro_oligo_cons_matrix$Var2, levels = c(as.character(mean_order_df$Var2)))
ggplot(astro_oligo_cons_matrix, aes(x = Var2, y = value)) + geom_boxplot(outlier.size = .5, outlier.alpha = .5) +
  ylab('Conservation of Adult coexpression (AUC)') + xlab('Top 100 Adult Astro/Oligo markers') + 
  theme(axis.text.x = element_blank()) 


```



get the conserved coexpression score per organoid dataset and check the distributions





```{r}

adult_org_glut_cons_coexp_score = colSums(glut_cons_matrix_1) / length(adult_glut_go)
adult_org_gaba_cons_coexp_score = colSums(gaba_cons_matrix_1) / length(adult_gaba_go)
adult_org_nonN_cons_coexp_score = colSums(nonN_cons_matrix_1) / length(adult_nonN_go)
adult_org_astro_oligo_cons_coexp_score = colSums(astro_oligo_cons_matrix_1) / length(adult_astro_oligo_go)

hist(adult_org_glut_cons_coexp_score)
hist(adult_org_gaba_cons_coexp_score)
hist(adult_org_nonN_cons_coexp_score)
hist(adult_org_astro_oligo_cons_coexp_score)


par(pty = 's')
plot(adult_org_glut_cons_coexp_score, adult_org_gaba_cons_coexp_score, xlab = 'Adult Glut conservation coexp score', ylab = 'Adult GABA conservation coexp score')
plot(adult_org_glut_cons_coexp_score, adult_org_nonN_cons_coexp_score, xlab = 'Adult Glut conservation coexp score', ylab = 'Adult NonN conservation coexp score')
plot(adult_org_nonN_cons_coexp_score, adult_org_gaba_cons_coexp_score, xlab = 'Adult NonN conservation coexp score', ylab = 'Adult GABA conservation coexp score')
plot(adult_org_nonN_cons_coexp_score, adult_org_astro_oligo_cons_coexp_score, xlab = 'Adult NonN conservation coexp score', ylab = 'Adult Astro/Oligo conservation coexp score')

```
Compare to the fetal conserved coexpression scores



```{r}

#pdf(file = 'graphs/coexpression/conserved_coexpression/adult_fetal_consCoexp_in_orgs_glut_v5.pdf', useDingbats = F, height = 4, width = 4)
#par(pty = 's')
#plot(all_sample_meta$glut_cons_coexp_metric, adult_org_glut_cons_coexp_score, 
#     xlab = 'Org. conservation of fetal Glut coexp', ylab = 'Org. conservation of Adult Glut coexp',
#     xlim = c(.3,1), ylim = c(.3,1), main = 'Glutamatergic')
#abline(a = 0, b = 1, col = 'red')
#dev.off()
cor(all_sample_meta$glut_cons_coexp_metric, adult_org_glut_cons_coexp_score, method = 'spearman')
par(pty = 's')
plot(all_sample_meta$glut_cons_coexp_metric, adult_org_glut_cons_coexp_score, 
     xlab = 'Org. conservation of fetal Glut coexp', ylab = 'Org. conservation of Adult Glut coexp',
     xlim = c(.3,1), ylim = c(.3,1), main = 'Glutamatergic')
abline(a = 0, b = 1, col = 'red')

#pdf(file = 'graphs/coexpression/conserved_coexpression/adult_fetal_consCoexp_in_orgs_gaba_v5.pdf', useDingbats = F, height = 4, width = 4)
#par(pty = 's')
#plot(all_sample_meta$gaba_cons_coexp_metric, adult_org_gaba_cons_coexp_score, 
#     xlab = 'Org. conservation of fetal GABA coexp', ylab = 'Org. conservation of Adult GABA coexp',
#     xlim = c(.3,1), ylim = c(.3,1), main = 'GABAergic')
#abline(a = 0, b = 1, col = 'red')
#dev.off()
cor(all_sample_meta$gaba_cons_coexp_metric, adult_org_gaba_cons_coexp_score, method = 'spearman')
par(pty = 's')
plot(all_sample_meta$gaba_cons_coexp_metric, adult_org_gaba_cons_coexp_score, 
     xlab = 'Org. conservation of fetal GABA coexp', ylab = 'Org. conservation of Adult GABA coexp',
     xlim = c(.3,1), ylim = c(.3,1), main = 'GABAergic')
abline(a = 0, b = 1, col = 'red')


#pdf(file = 'graphs/coexpression/conserved_coexpression/adult_fetal_consCoexp_in_orgs_nonN_v5.pdf', useDingbats = F, height = 4, width = 4)
#par(pty = 's')
#plot(all_sample_meta$nonN_cons_coexp_metric, adult_org_nonN_cons_coexp_score, 
#     xlab = 'Org. conservation of fetal nonN coexp', ylab = 'Org. conservation of Adult nonN coexp',
#     xlim = c(.3,1), ylim = c(.3,1), main = 'Non-neuronal')
#abline(a = 0, b = 1, col = 'red')
#dev.off()
cor(all_sample_meta$nonN_cons_coexp_metric, adult_org_nonN_cons_coexp_score, method = 'spearman')
par(pty = 's')
plot(all_sample_meta$nonN_cons_coexp_metric, adult_org_nonN_cons_coexp_score, 
     xlab = 'Org. conservation of fetal nonN coexp', ylab = 'Org. conservation of Adult nonN coexp',
     xlim = c(.3,1), ylim = c(.3,1), main = 'Non-neuronal')
abline(a = 0, b = 1, col = 'red')


#pdf(file = 'graphs/coexpression/conserved_coexpression/adult_fetal_consCoexp_in_orgs_astro_oligo_v5.pdf', useDingbats = F, height = 4, width = 4)
#par(pty = 's')
#plot(all_sample_meta$nonN_cons_coexp_metric, adult_org_astro_oligo_cons_coexp_score, 
#     xlab = 'Org. conservation of fetal nonN coexp', ylab = 'Org. conservation of Adult Astro/Oligo coexp',
#     xlim = c(.3,1), ylim = c(.3,1), main = 'Astro/Oligo')
#abline(a = 0, b = 1, col = 'red')
#dev.off()
cor(all_sample_meta$nonN_cons_coexp_metric, adult_org_astro_oligo_cons_coexp_score, method = 'spearman')
par(pty = 's')
plot(all_sample_meta$nonN_cons_coexp_metric, adult_org_astro_oligo_cons_coexp_score, 
     xlab = 'Org. conservation of fetal nonN coexp', ylab = 'Org. conservation of Adult Astro/Oligo coexp',
     xlim = c(.3,1), ylim = c(.3,1), main = 'Astro/Oligo')
abline(a = 0, b = 1, col = 'red')

```



```{r}
all_sample_meta$adult_gaba_cons_coexp = adult_org_gaba_cons_coexp_score
all_sample_meta$adult_glut_cons_coexp = adult_org_glut_cons_coexp_score
all_sample_meta$adult_nonN_cons_coexp = adult_org_nonN_cons_coexp_score
all_sample_meta$adult_astro_oligo_cons_coexp = adult_org_astro_oligo_cons_coexp_score

all_sample_meta %>% filter(adult_glut_cons_coexp >= .8 & glut_cons_coexp_metric >= .7 )

```


For the best and worst organoid datasets, show the global distributions of conserved co-expression aurocs against the aggregate fetal network
Include the adult performance against the fetal network

```{r}
load('../../../data/coexpression/fetal/ranked_agg_fetal_coexp_network_8_17_23.Rdata')
dim(fetal_ranked_aggregate)
dim(adult_rank_corr_mat)

```

get the adult performance for all genes against the fetal data

```{r}
top_genes = 10
start = Sys.time()
adult_to_fetal_consCoexp = unlist(mclapply(1:length(go_genes), function(i) get_conserved_coexp(fetal_ranked_aggregate, adult_rank_corr_mat, go_genes[i],top_genes ), mc.cores = 10 ))
mean_adult_to_fetal = mean(adult_to_fetal_consCoexp)
end = Sys.time()
end - start


hist(adult_to_fetal_consCoexp, breaks = seq(0,1, .001))
density_adult = density(adult_to_fetal_consCoexp)
plot(density_adult)

mean_adult_to_fetal
```



Now get a distribution of the mean conserved co-expression aurocs across all genes for each organoid data, using the fetal data as a reference
Will use to highlight the organoid datasets that perform worse than the adult co-expression. More removed from fetal data than adult, clearly not a good developmental model

Already did this, stored in the organoid metadata



```{r}

hist(all_sample_meta$genome_cons_coexp_score, breaks = seq(.5,1,.01))

```


```{r}
#pdf(file = 'graphs/coexpression/conserved_coexpression/mean_org_and_adult_consCoexp_hist_v3.pdf', 
#   height = 4, width = 4, useDingbats = F)
#par(pty = 's')
#hist(all_sample_meta$genome_cons_coexp_score, breaks = seq(.5,1,.01), xlim = c(.5,1), 
#     xlab = 'Mean Fetal Conserved Co-expression AUROC', main = '')
#legend("topright", c("Organoid", "Adult"), col=c("grey", "red"), lwd=10, cex = .75)
#abline(v = mean_adult_to_fetal, col = 'red', lwd = 3)
#dev.off()

par(pty = 's')
hist(all_sample_meta$genome_cons_coexp_score, breaks = seq(.5,1,.01), xlim = c(.5,1), 
     xlab = 'Mean Fetal Conserved Co-expression AUROC', main = '')
legend("topright", c("Organoid", "Adult"), col=c("grey", "red"), lwd=10, cex = .75)
abline(v = mean_adult_to_fetal, col = 'red', lwd = 3)
```


```{r}

save( adult_org_glut_cons_coexp_score,  adult_org_gaba_cons_coexp_score,  adult_org_nonN_cons_coexp_score, mean_adult_to_fetal,
      file = '/home/werner/projects/meta_qc_organoid/data/coexpression/adult_coexpression/adult_pres_coexp_scores.Rdata')

```




```{r}
table(all_sample_meta$genome_cons_coexp_score <= mean_adult_to_fetal)


```

compare fetal consCoexp global score with organoid cell number
Include the downsampled fetal data

```{r}
load('../../../data/coexpression/fetal_downsample_consCoexp_fetal_matrix_v2.Rdata')
downsample_mat
```

```{r}

cor_permute = function(x,y, method_cor){
  permute_y = sample(y,size=length(y), replace=FALSE)
  corr = suppressWarnings(cor(x,permute_y, method=method_cor))
  return(corr)
}

num_permutes = 10000

```





```{r}

bad_org_index = all_sample_meta$genome_cons_coexp_score <= mean_adult_to_fetal

all_sample_meta[bad_org_index, ]
hist(log10(as.numeric(all_sample_meta$post_qc_cell_number[!bad_org_index])), breaks = 24, xlim = c(0,6))
hist(log10(as.numeric(all_sample_meta$post_qc_cell_number[bad_org_index])), add = T, col = 'blue', breaks = 24)

#pdf(file = 'graphs/coexpression/conserved_coexpression/mean_org_consCoexp_and_cellNum_scatter_v3.pdf', 
#    height = 4, width = 4, useDingbats = F)
#par(pty = 's')
#plot(log10(as.numeric(all_sample_meta$post_qc_cell_number)), all_sample_meta$genome_cons_coexp_score, ylim = c(.5, 1), 
#     xlab = 'log10(number of cells)', ylab = 'Mean Fetal Conserved Co-expression AUROC')

#dev.off()

par(pty = 's')
plot(log10(as.numeric(all_sample_meta$post_qc_cell_number)), all_sample_meta$genome_cons_coexp_score, ylim = c(.5, 1), 
     xlab = 'log10(number of cells)', ylab = 'Mean Fetal Conserved Co-expression AUROC')

original_cor = cor(as.numeric(all_sample_meta$post_qc_cell_number), all_sample_meta$genome_cons_coexp_score, method = 'spearman')
#Do permutation to get a correlation pvalue
permute_cors = unlist(mclapply(1:num_permutes, function(i) 
  cor_permute(as.numeric(all_sample_meta$post_qc_cell_number),  all_sample_meta$genome_cons_coexp_score, 'spearman'), mc.cores=5))
p_val = sum(abs(permute_cors) >= abs(original_cor) ) / num_permutes 
original_cor
p_val


hist(permute_cors, breaks = 64)

```

```{r}

mean_fetal_auroc = colMeans(downsample_mat)
sd_fetal_auroc = apply(downsample_mat, 2, sd)
fetal_downsamp_fetal_consCoexp_df = data.frame(mean_auroc = mean_fetal_auroc, cell_num = c(100,500,1000,5000,10000,50000,100000), 
                                               top_error = mean_fetal_auroc + sd_fetal_auroc,
                                               bottom_error = mean_fetal_auroc - sd_fetal_auroc)

org_global_fetal_consCoexp_df = data.frame(mean_auroc =  all_sample_meta$genome_cons_coexp_score, cell_num =as.numeric(all_sample_meta$post_qc_cell_number))


ggplot(fetal_downsamp_fetal_consCoexp_df, aes(x = log10(cell_num), y = mean_auroc, color = 'Fetal')) + geom_point() + geom_line() +
  geom_errorbar(aes(ymin = bottom_error, ymax = top_error, color = 'Fetal'), width = .3) +
  geom_hline(yintercept = mean_adult_to_fetal,color = 'red') +
  geom_point(data = org_global_fetal_consCoexp_df, aes(x = log10(cell_num), y = mean_auroc, color = 'Organoid'), size = 2 ) +
  xlab('log10( Number of cells )') + ylab('Mean global fetal preserved co-expression score') + ylim(.5, .85) +
  scale_color_manual(breaks = c('Organoid','Fetal','Adult'), 
                     values = c('Organoid' = 'black', 'Fetal'= 'blue', 'Adult' ='red')) +
  theme(axis.text = element_text(size = 12))
#ggsave(filename = 'mean_org_consCoexp_and_cellNum_scatter_downsamp_fetal_v5.pdf', path = 'graphs/coexpression/conserved_coexpression/',
#       useDingbats = F, device = 'pdf', height = 4, width = 4)
```


```{r}

save(fetal_downsamp_fetal_consCoexp_df, file = '/home/werner/projects/meta_qc_organoid/data/coexpression/adult_coexpression/fetal_downsamp_fetal_consCoexp_df.Rdata')

```








do a leave one out fetal analysis too

```{r}

ranked_go_coexp_paths = c( '../data/coexpression/fetal/8_17_23_GO_coexp/ranked_fan_fu_coexp.Rdata',
                 '../data/coexpression/fetal/8_17_23_GO_coexp/ranked_poliodakis_batch1_coexp.Rdata',
                 '../data/coexpression/fetal/8_17_23_GO_coexp/ranked_poliodakis_batch2_coexp.Rdata',
                 '../data/coexpression/fetal/8_17_23_GO_coexp/ranked_arealization_GW14_coexp.Rdata',
                 '../data/coexpression/fetal/8_17_23_GO_coexp/ranked_arealization_GW18_2_coexp.Rdata',
                 '../data/coexpression/fetal/8_17_23_GO_coexp/ranked_arealization_GW19_coexp.Rdata',
                 '../data/coexpression/fetal/8_17_23_GO_coexp/ranked_arealization_GW19_2_coexp.Rdata',
                 '../data/coexpression/fetal/8_17_23_GO_coexp/ranked_arealization_GW20_coexp.Rdata',
                 '../data/coexpression/fetal/8_17_23_GO_coexp/ranked_arealization_GW20_31_coexp.Rdata',
                 '../data/coexpression/fetal/8_17_23_GO_coexp/ranked_arealization_GW20_34_coexp.Rdata',
                 '../data/coexpression/fetal/8_17_23_GO_coexp/ranked_arealization_GW25_coexp.Rdata', 
                 '../data/coexpression/fetal/8_17_23_GO_coexp/ranked_linnarsson_GW5_coexp.Rdata',
                 '../data/coexpression/fetal/8_17_23_GO_coexp/ranked_linnarsson_GW5-5_coexp.Rdata',
                 '../data/coexpression/fetal/8_17_23_GO_coexp/ranked_linnarsson_GW6_1_coexp.Rdata',
                 '../data/coexpression/fetal/8_17_23_GO_coexp/ranked_linnarsson_GW6_2_coexp.Rdata',
                 '../data/coexpression/fetal/8_17_23_GO_coexp/ranked_linnarsson_GW6-6_1_coexp.Rdata',
                 '../data/coexpression/fetal/8_17_23_GO_coexp/ranked_linnarsson_GW6-6_2_coexp.Rdata',
                 '../data/coexpression/fetal/8_17_23_GO_coexp/ranked_linnarsson_GW6-7_coexp.Rdata',
                 '../data/coexpression/fetal/8_17_23_GO_coexp/ranked_linnarsson_GW6-9_1_coexp.Rdata',
                 '../data/coexpression/fetal/8_17_23_GO_coexp/ranked_linnarsson_GW6-9_2_coexp.Rdata',
                 '../data/coexpression/fetal/8_17_23_GO_coexp/ranked_linnarsson_GW6-9_3_coexp.Rdata',
                 '../data/coexpression/fetal/8_17_23_GO_coexp/ranked_linnarsson_GW6-9_4_coexp.Rdata',
                 '../data/coexpression/fetal/8_17_23_GO_coexp/ranked_linnarsson_GW7_coexp.Rdata',
                 '../data/coexpression/fetal/8_17_23_GO_coexp/ranked_linnarsson_GW7-5_coexp.Rdata',
                 '../data/coexpression/fetal/8_17_23_GO_coexp/ranked_linnarsson_GW8_1_coexp.Rdata',
                 '../data/coexpression/fetal/8_17_23_GO_coexp/ranked_linnarsson_GW8_2_coexp.Rdata',
                 '../data/coexpression/fetal/8_17_23_GO_coexp/ranked_linnarsson_GW8_3_coexp.Rdata',
                 '../data/coexpression/fetal/8_17_23_GO_coexp/ranked_linnarsson_GW8-1_coexp.Rdata',
                 '../data/coexpression/fetal/8_17_23_GO_coexp/ranked_linnarsson_GW8-5_1_coexp.Rdata',
                 '../data/coexpression/fetal/8_17_23_GO_coexp/ranked_linnarsson_GW8-5_2_coexp.Rdata',
                 '../data/coexpression/fetal/8_17_23_GO_coexp/ranked_linnarsson_GW9-2_coexp.Rdata',
                 '../data/coexpression/fetal/8_17_23_GO_coexp/ranked_linnarsson_GW9-5_coexp.Rdata',
                 '../data/coexpression/fetal/8_17_23_GO_coexp/ranked_linnarsson_GW10_coexp.Rdata',
                 '../data/coexpression/fetal/8_17_23_GO_coexp/ranked_linnarsson_GW11-5_coexp.Rdata',
                 '../data/coexpression/fetal/8_17_23_GO_coexp/ranked_linnarsson_GW12_coexp.Rdata',
                 '../data/coexpression/fetal/8_17_23_GO_coexp/ranked_linnarsson_GW13_coexp.Rdata',
                 '../data/coexpression/fetal/8_17_23_GO_coexp/ranked_linnarsson_GW14_coexp.Rdata')


dataset_names = c('GW7-28','GW17-18_1', 'GW17-18_2','GW14_2','GW18','GW19_1','GW19_2','GW20_1','GW20_2','GW20_3','GW25',
                  'GW5','GW5-5','GW6_1','GW6_2','GW6-6_1','GW6-6_2','GW6-7','GW6-9_1','GW6-9_2','GW6-9_3','GW6-9_4','GW7','GW7-5','GW8_1','GW8_2','GW8_3','GW8-1','GW8-5_1',
                  'GW8-5_2','GW9-2','GW9-5','GW10','GW11-5','GW12','GW13','GW14_1')

```


```{r}

mean_fetal_consCoexp_scores = vector(mode = 'numeric', length = length(ranked_go_coexp_paths))

for(j in 1:length(ranked_go_coexp_paths)){
  start = Sys.time()
  #Aggregate all the fetal data
  current_dataset = dataset_names[j]
  temp_fetal_ranked_coexp = ranked_go_coexp_paths[dataset_names != current_dataset]
  
  #Aggregate leaving out that dataset
  sum_rank_corr_genes = matrix(rep(0, length(go_genes)^2), nrow = length(go_genes), ncol = length(go_genes), 
                               dimnames = list(go_genes, go_genes))
  
  for(network in 1:length(temp_fetal_ranked_coexp)){
    x = load(temp_fetal_ranked_coexp[network])
    rank_mat = get(x)
    rm(x)
    
    #Reorder to match the same gene order
    index = match(go_genes, rownames(rank_mat))
    rank_mat = rank_mat[index, index]
  
    sum_rank_corr_genes = sum_rank_corr_genes + rank_mat
    
  }

  #Divide sums by the number of datasets to get averages
  agg_rank_mat = sum_rank_corr_genes / (length(dataset_names) - 1)
  
  #Rank the coexpression data
  rank_corr = frank(c(agg_rank_mat), ties.method = 'min') - 1
  #Divide by max rank squeezes everything between 0 and 1
  norm_rank_corr = rank_corr / max(rank_corr)
  #Put back into a matrix
  agg_rank_mat = matrix(norm_rank_corr, nrow = nrow(agg_rank_mat), ncol = ncol(agg_rank_mat), 
                            dimnames = list(rownames(agg_rank_mat), colnames(agg_rank_mat)))
 
  diag(agg_rank_mat) = 1
  print('Done aggregating networks...')
  
  #Load in the test fetal network
  x = load(ranked_go_coexp_paths[j])
  rank_mat = get(x)
  rm(x)
  
  
  #Get all the conserved co-expression scores and then save the mean
  all_consCoexp_scores_fetal = unlist(mclapply(1:length(go_genes), 
                                function(i) get_conserved_coexp(agg_rank_mat, rank_mat, go_genes[i],top_genes ),mc.cores = 10))
  mean_fetal_consCoexp_scores[j] = mean(all_consCoexp_scores_fetal)
  
  end = Sys.time()
  print(sprintf('Done with %i fetal datasets', j))
  print(end - start)

}

```

```{r}
pdf(file = 'graphs/coexpression/conserved_coexpression/mean_org_and_adult_and_fetal_consCoexp_hist_v5.pdf', 
    height = 4, width = 4, useDingbats = F)

par(pty = 's')
hist(all_sample_meta$genome_cons_coexp_score, breaks = seq(.5,1,.01), xlim = c(.5,1), 
     xlab = 'Mean Fetal Preserved Co-expression AUROC', main = '')
legend("topright", c("Organoid", 'Fetal', "Adult"), col=c("grey",rgb(0,0,1,.85), "red"), lwd=10, cex = .75)
abline(v = mean_adult_to_fetal, col = 'red', lwd = 3)
hist(mean_fetal_consCoexp_scores, breaks = seq(.5,1,.01), add = T, col = rgb(0,0,1,.85))

dev.off()

par(pty = 's')
hist(all_sample_meta$genome_cons_coexp_score, breaks = seq(.5,1,.01), xlim = c(.5,1), 
     xlab = 'Mean Fetal Preserved Co-expression AUROC', main = '')
legend("topright", c("Organoid", 'Fetal', "Adult"), col=c("grey",rgb(0,0,1,.85), "red"), lwd=10, cex = .75)
abline(v = mean_adult_to_fetal, col = 'red', lwd = 3)
hist(mean_fetal_consCoexp_scores, breaks = seq(.5,1,.01), add = T, col = rgb(0,0,1,.85))

```

```{r}
save(mean_fetal_consCoexp_scores, file = '../data/coexpression/mean_global_consCoexp_org_and_fetal_v3.Rdata' )

```



Save the organoid metadata too aswell

```{r}

write.csv(all_sample_meta, file = '../data/seurat_objs/all_data_just_meta_seurat_with_addl_features_v3.csv', row.names = F)
```





#############################
Pasca organoid transplantation dataset
compare non-transplanted and transplanted organoids with the conservation of adult co-expression


##############################



```{r}


pasca_temp = filter(all_sample_meta, grepl('revah', sample_ids)) %>% select(sample_ids, Condition, adult_gaba_cons_coexp, adult_glut_cons_coexp, adult_astro_oligo_cons_coexp )
pasca_temp = reshape2::melt(pasca_temp)
pasca_temp

ggplot(pasca_temp, aes(x = Condition, y = value)) + geom_point(size = 2, alpha = .5) +
  facet_wrap(~variable) + ylim(.5, 1) + ylab('Preserved Co-expression AUROC') + ggtitle('Preservation of Adult Co-expression')

#ggsave(filename = 'pasca_transplant_consCoexp_adult_celltypes_v2.pdf', path = 'graphs/coexpression/conserved_coexpression/pasca_transplant/', 
#       device = 'pdf', 
#       height = 4, width = 6, useDingbats = F)
```






```{r}

pasca_temp %>% filter(Condition == 'Normal') %>% group_by(variable) %>% summarize(mean = mean(value), sd = sd(value))
pasca_temp %>% filter(Condition == 'Transplanted') %>% group_by(variable) %>% summarize(mean = mean(value), sd = sd(value))
```


normalize to the average of the non-transplanted and plot log2 FC

```{r}
#Get the means of the non tranplanted data
adult_glut_nonT_mean = mean(filter(pasca_temp,Condition == 'Normal' & variable == 'adult_glut_cons_coexp')$value)
adult_gaba_nonT_mean = mean(filter(pasca_temp,Condition == 'Normal' & variable == 'adult_gaba_cons_coexp')$value)
adult_nonN_nonT_mean = mean(filter(pasca_temp,Condition == 'Normal' & variable == 'adult_astro_oligo_cons_coexp')$value)


mean_norm_vec = vector(mode = 'numeric', length = nrow(pasca_temp))

glut_index = pasca_temp$variable == 'adult_glut_cons_coexp'
mean_norm_vec[glut_index] = pasca_temp$value[glut_index] / adult_glut_nonT_mean
gaba_index = pasca_temp$variable == 'adult_gaba_cons_coexp'
mean_norm_vec[gaba_index] = pasca_temp$value[gaba_index] / adult_gaba_nonT_mean
nonN_index = pasca_temp$variable == 'adult_astro_oligo_cons_coexp'
mean_norm_vec[nonN_index] = pasca_temp$value[nonN_index] / adult_nonN_nonT_mean

mean_norm_vec = log(mean_norm_vec, base = 2)
pasca_temp$nonT_FC = mean_norm_vec

pasca_temp

ggplot(pasca_temp, aes(x = Condition, y = nonT_FC)) + geom_point(size = 2, alpha = .5) + facet_wrap(~variable) +
  ylab('log2(Non-transplanted fold change)') + ggtitle('Preservation of Adult Co-expression')
#ggsave(filename = 'pasca_transplant_consCoexp_adult_celltypes_log2FC_v2.pdf', path = 'graphs/coexpression/conserved_coexpression/pasca_transplant/', 
#       device = 'pdf', 
#       height = 4, width = 6, useDingbats = F)
```

```{r}
pasca_adult_df = pasca_temp
save(pasca_adult_df, file = '/home/werner/projects/meta_qc_organoid/data/coexpression/adult_coexpression/pasca_adult_df.Rdata' )

```

