
Exploring the preservation of co-expression for cell-type specific gene markers, at a higher level of granularity than the broad cell-types used in the rest of the paper



Redo the analysis getting preserved co-expression scores for each gene in each dataset, this time save all the data, not just the mean value

```{r, setup, include=FALSE}
knitr::opts_knit$set(root.dir = '/home/werner/projects/meta_qc_organoid/code')
```


```{r}
library(EGAD)
library(dplyr)
library(ggplot2)
library(parallel)
library(Seurat)
library(MetaMarkers)
library(data.table)

```

```{r}
load('../data/go_annotations/go_annotations.Rdata')
length(go_genes)
```


```{r}
compute_auc = function(scores, label){
  label = as.logical(label)
  n1 = as.numeric(sum(label))
  n2 = as.numeric(sum(!label))
  R1 = sum(rank(scores)[label])
  U1 = R1 - n1 * (n1 + 1)/2
  auc = U1/(n1 * n2)
  return(auc)
}


get_conserved_coexp = function(network1, network2, test_gene, num_top = 10){

  #Get the top 10 coexpressed genes
  test_gene_coexp = network1[test_gene, ]
  test_gene_coexp = test_gene_coexp[names(test_gene_coexp) != test_gene]
  top_coexp = names(test_gene_coexp[order(test_gene_coexp, decreasing = T)][1:num_top])
  
  #Calculate the AUC using the organoid data
  scores = network2[test_gene, ]
  scores = scores[names(scores) != test_gene]
  labels = names(scores) %in% top_coexp
  
  #Get the auc
  auc = compute_auc(scores, labels)
  return(auc)
}

```


aagregated fetal co-expression network
```{r}

load('../data/coexpression/fetal/ranked_agg_fetal_coexp_network_8_17_23.Rdata')
dim(fetal_ranked_aggregate)

```


organoid metadata

```{r}

all_sample_meta = read.csv('../data/seurat_objs/all_data_just_meta_seurat_with_addl_features_v2.csv')
all_sample_meta
```




the mean gene dataframes
```{r}
load(file = "../data/coexpression/all_gene_presCoexp/all_gene_presCoexp_mats_v3.Rdata")

```


```{r}

fetal_all_gene_presCoexp_df

org_all_gene_presCoexp_df
```


organoid preserved co-expression for all genes

```{r}
top_genes = 10
all_gene_org_pres_fetal_mat = matrix(nrow = length(go_genes), ncol = length(all_sample_meta$sample_ids), dimnames = list(go_genes, all_sample_meta$sample_ids))

for(j in 1:length(all_sample_meta$sample_ids)){
  load(sprintf('../data/coexpression/organoid/6_26_24_GO_coexp/ranked_%s_coexp.Rdata', all_sample_meta$sample_ids[j]))
  
  start = Sys.time()
  test_all_genes = unlist(mclapply(1:length(go_genes), function(i) get_conserved_coexp(fetal_ranked_aggregate, rank_corr_matrix, go_genes[i],top_genes ), mc.cores = 10 ))
  end = Sys.time()
  print(end - start)
  
  all_gene_org_pres_fetal_mat[ ,j] = test_all_genes

}
```


```{r}

all_gene_org_pres_fetal_mat[1:10, 1:10]
```




And now do the fetal datasets


```{r}
ranked_go_coexp_paths =list( 
                 fan_fu = '../data/coexpression/fetal/8_17_23_GO_coexp/ranked_fan_fu_coexp.Rdata',
                 poliodakis_1 = '../data/coexpression/fetal/8_17_23_GO_coexp/ranked_poliodakis_batch1_coexp.Rdata',
                 poliodakis_2 = '../data/coexpression/fetal/8_17_23_GO_coexp/ranked_poliodakis_batch2_coexp.Rdata',
                 GW14 = '../data/coexpression/fetal/8_17_23_GO_coexp/ranked_arealization_GW14_coexp.Rdata',
                 GW18_2 = '../data/coexpression/fetal/8_17_23_GO_coexp/ranked_arealization_GW18_2_coexp.Rdata',
                 GW19 = '../data/coexpression/fetal/8_17_23_GO_coexp/ranked_arealization_GW19_coexp.Rdata',
                 GW19_2 = '../data/coexpression/fetal/8_17_23_GO_coexp/ranked_arealization_GW19_2_coexp.Rdata',
                 GW20 = '../data/coexpression/fetal/8_17_23_GO_coexp/ranked_arealization_GW20_coexp.Rdata',
                 GW20_31 = '../data/coexpression/fetal/8_17_23_GO_coexp/ranked_arealization_GW20_31_coexp.Rdata',
                 GW20_34 = '../data/coexpression/fetal/8_17_23_GO_coexp/ranked_arealization_GW20_34_coexp.Rdata',
                 GW25 = '../data/coexpression/fetal/8_17_23_GO_coexp/ranked_arealization_GW25_coexp.Rdata',
                 linnarsson_GW5 = '../data/coexpression/fetal/8_17_23_GO_coexp/ranked_linnarsson_GW5_coexp.Rdata',
                 linnarsson_GW5_point_5 = '../data/coexpression/fetal/8_17_23_GO_coexp/ranked_linnarsson_GW5-5_coexp.Rdata',
                 linnarsson_GW6_1 ='../data/coexpression/fetal/8_17_23_GO_coexp/ranked_linnarsson_GW6_1_coexp.Rdata',
                 linnarsson_GW6_2 = '../data/coexpression/fetal/8_17_23_GO_coexp/ranked_linnarsson_GW6_2_coexp.Rdata',
                 linnarsson_GW6_point_6_1 = '../data/coexpression/fetal/8_17_23_GO_coexp/ranked_linnarsson_GW6-6_1_coexp.Rdata',
                 linnarsson_GW6_point_6_2 ='../data/coexpression/fetal/8_17_23_GO_coexp/ranked_linnarsson_GW6-6_2_coexp.Rdata',
                 linnarsson_GW6_point_7 ='../data/coexpression/fetal/8_17_23_GO_coexp/ranked_linnarsson_GW6-7_coexp.Rdata',
                 linnarsson_GW6_point_9_1 = '../data/coexpression/fetal/8_17_23_GO_coexp/ranked_linnarsson_GW6-9_1_coexp.Rdata',
                 linnarsson_GW6_point_9_2 ='../data/coexpression/fetal/8_17_23_GO_coexp/ranked_linnarsson_GW6-9_2_coexp.Rdata',
                 linnarsson_GW6_point_9_3 ='../data/coexpression/fetal/8_17_23_GO_coexp/ranked_linnarsson_GW6-9_3_coexp.Rdata',
                 linnarsson_GW6_point_9_4 = '../data/coexpression/fetal/8_17_23_GO_coexp/ranked_linnarsson_GW6-9_4_coexp.Rdata',
                 linnarsson_GW7 = '../data/coexpression/fetal/8_17_23_GO_coexp/ranked_linnarsson_GW7_coexp.Rdata',
                 linnarsson_GW7_point_5 = '../data/coexpression/fetal/8_17_23_GO_coexp/ranked_linnarsson_GW7-5_coexp.Rdata',
                 linnarsson_GW8_1 = '../data/coexpression/fetal/8_17_23_GO_coexp/ranked_linnarsson_GW8_1_coexp.Rdata',
                 linnarsson_GW8_2 = '../data/coexpression/fetal/8_17_23_GO_coexp/ranked_linnarsson_GW8_2_coexp.Rdata',
                 linnarsson_GW8_3 = '../data/coexpression/fetal/8_17_23_GO_coexp/ranked_linnarsson_GW8_3_coexp.Rdata',
                 linnarsson_GW8_point_1 =  '../data/coexpression/fetal/8_17_23_GO_coexp/ranked_linnarsson_GW8-1_coexp.Rdata',
                 linnarsson_GW8_point_5_1 ='../data/coexpression/fetal/8_17_23_GO_coexp/ranked_linnarsson_GW8-5_1_coexp.Rdata',
                 linnarsson_GW8_point_5_2 = '../data/coexpression/fetal/8_17_23_GO_coexp/ranked_linnarsson_GW8-5_2_coexp.Rdata',
                 linnarsson_GW9_point_2 =  '../data/coexpression/fetal/8_17_23_GO_coexp/ranked_linnarsson_GW9-2_coexp.Rdata',
                 linnarsson_GW9_point_5 ='../data/coexpression/fetal/8_17_23_GO_coexp/ranked_linnarsson_GW9-5_coexp.Rdata',
                 linnarsson_GW10 = '../data/coexpression/fetal/8_17_23_GO_coexp/ranked_linnarsson_GW10_coexp.Rdata',
                 linnarsson_GW11_point_5 ='../data/coexpression/fetal/8_17_23_GO_coexp/ranked_linnarsson_GW11-5_coexp.Rdata',
                 linnarsson_GW12 ='../data/coexpression/fetal/8_17_23_GO_coexp/ranked_linnarsson_GW12_coexp.Rdata',
                 linnarsson_GW13 ='../data/coexpression/fetal/8_17_23_GO_coexp/ranked_linnarsson_GW13_coexp.Rdata',
                 linnarsson_GW14 ='../data/coexpression/fetal/8_17_23_GO_coexp/ranked_linnarsson_GW14_coexp.Rdata')

annot_marker_names = c('fan_fu','poliodakis_1','poliodakis_2','GW14','GW18_2','GW19','GW19_2','GW20','GW20_31','GW20_34','GW25',
                        'linnarsson_GW5', 'linnarsson_GW5_point_5', 'linnarsson_GW6_1', 'linnarsson_GW6_2', 'linnarsson_GW6_point_6_1', 'linnarsson_GW6_point_6_2', 'linnarsson_GW6_point_7',
                  'linnarsson_GW6_point_9_1', 'linnarsson_GW6_point_9_2', 'linnarsson_GW6_point_9_3', 'linnarsson_GW6_point_9_4', 'linnarsson_GW7', 'linnarsson_GW7_point_5', 
                  'linnarsson_GW8_1', 'linnarsson_GW8_2','linnarsson_GW8_3', 'linnarsson_GW8_point_1', 'linnarsson_GW8_point_5_1', 'linnarsson_GW8_point_5_2', 'linnarsson_GW9_point_2',
                  'linnarsson_GW9_point_5','linnarsson_GW10', 'linnarsson_GW11_point_5', 'linnarsson_GW12', 'linnarsson_GW13', 'linnarsson_GW14')


```



```{r}


top_genes = 10
all_gene_fetal_pres_fetal_mat = matrix(nrow = length(go_genes), ncol = length(annot_marker_names), dimnames = list(go_genes, annot_marker_names))

for(j in 1:length(annot_marker_names)){

  start = Sys.time()
  current_dataset = annot_marker_names[j]
  #Leave one out and get the new aggregated fetal network
  sum_rank_corr_genes = matrix(rep(0, length(go_genes)^2), nrow = length(go_genes), ncol = length(go_genes), dimnames = list(go_genes, go_genes))
  current_ranked_go_coexp_paths = ranked_go_coexp_paths[names(ranked_go_coexp_paths) != current_dataset]
  
  for(i in 1:length(current_ranked_go_coexp_paths)){
    x = load(current_ranked_go_coexp_paths[[i]])
    rank_mat = get(x)
    rm(x)
    #Reorder to match the same gene order
    index = match(go_genes, rownames(rank_mat))
    rank_mat = rank_mat[index, index]
    sum_rank_corr_genes = sum_rank_corr_genes + rank_mat
  }
  #Divide sums by the number of datasets to get averages
  agg_rank_mat = sum_rank_corr_genes / length(current_ranked_go_coexp_paths)

  #Rank
  rank_corr = frank(c(agg_rank_mat), ties.method = 'min') - 1
  #Divide by max rank squeezes everything between 0 and 1
  norm_rank_corr = rank_corr / max(rank_corr)
  #Put back into a matrix
  rank_corr_matrix = matrix(norm_rank_corr, nrow = nrow(agg_rank_mat), ncol = ncol(agg_rank_mat), 
                            dimnames = list(rownames(agg_rank_mat), colnames(agg_rank_mat)))
  agg_rank_mat = rank_corr_matrix
  diag(agg_rank_mat) = 1
  print('Done aggregating networks...')
  
  #Get the preserved coexpression aurocs for all genes
  x = load(ranked_go_coexp_paths[[j]])
  rank_mat = get(x)
  rm(x)
  
  test_all_genes = unlist(mclapply(1:length(go_genes), function(k) get_conserved_coexp(agg_rank_mat, rank_mat, go_genes[k],top_genes ), mc.cores = 10 ))
  all_gene_fetal_pres_fetal_mat[ ,j] = test_all_genes
  
  end = Sys.time()
  print(sprintf('Done with %i datasets...', j))
  print(end - start)
  
}

```


```{r}

all_gene_fetal_pres_fetal_mat[1:10, 1:10]
```


```{r}
save(all_gene_org_pres_fetal_mat, all_gene_fetal_pres_fetal_mat, file = "../data/coexpression/all_gene_presCoexp/all_gene_and_dataset_presCoexp_mats_v3.Rdata")

```

```{r}
load(file = "../data/coexpression/all_gene_presCoexp/all_gene_and_dataset_presCoexp_mats_v3.Rdata")

```



explore the cell-type markers
Starting with the organoid morphogen screen preprint from the Pasca lab

load in the supplemental DE tables and grab the top DE genes as markers per cluster

Data is ordered by adjusted p-value, but there are 0 p-values. Add a Rank column for each cluster for easy access
```{r}

pasca_cluster_de = read.csv('../data/pasca_morpho_screen/pasca_morpho_supp_tables/pasca_morpho_screen_org_cluster_DE.csv')
#Add rank column
pasca_cluster_de = pasca_cluster_de %>% group_by(cluster) %>% mutate(rank = 1:length(gene) )

pasca_cluster_de

#Example to grab the top 10 markers of a cluster
filter(pasca_cluster_de, cluster == 0 & rank <= 10)$gene

```

```{r}


pasca_cluster_annotations = read.csv('../data/pasca_morpho_screen/pasca_morpho_supp_tables/pasca_morpho_cluster_annotations.csv')
pasca_cluster_annotations
```





Compare the preserved co-expression of cell-type markers in organoid datasets to the preserved co-expression MetaMarker score for that cell type



```{r}

n_datasets = nrow(all_sample_meta)

all_cluster_nums = c()
all_cluster_annotations = c()
all_dataset_ids = c()
all_dataset_protocols = c()
all_cluster_presCoexp = c()

for(i in 0:max(pasca_cluster_de$cluster)){

  cluster_annotation = pasca_cluster_annotations$name_fig2[pasca_cluster_annotations$cluster == i]
  #Grab markers for that cluster
  cluster_markers = filter(pasca_cluster_de, cluster == i & rank <= 10)$gene
  
  #Index for the average marker expression and the average preserved co-exp aurocs
  gene_index = rownames(all_gene_org_pres_fetal_mat) %in% cluster_markers
  num_go = sum(gene_index)
  avg_presCoexp_vec = colMeans(all_gene_org_pres_fetal_mat[gene_index, ], na.rm = T)
  
  all_cluster_nums = c(all_cluster_nums, rep(i, length = n_datasets))
  all_cluster_annotations = c(all_cluster_annotations, rep(cluster_annotation, length = n_datasets))
  all_dataset_ids = c(all_dataset_ids, all_sample_meta$sample_ids)
  all_dataset_protocols = c(all_dataset_protocols, all_sample_meta$Protocol.classification)
  all_cluster_presCoexp = c(all_cluster_presCoexp, unname(avg_presCoexp_vec))


}
org_celltype_presCoexp_df = data.frame(cluster = all_cluster_nums, annotation = all_cluster_annotations, org_dataset = all_dataset_ids,
                                       org_protocol = all_dataset_protocols, presCoexp = all_cluster_presCoexp)


org_celltype_presCoexp_df

```

```{r}

all_protocols = names(table(org_celltype_presCoexp_df$org_protocol))

plot_celltype_presCoexp = function(org_celltype_df, protocol_type, useDots = F){
  
  order_df = org_celltype_presCoexp_df %>% filter(org_protocol == protocol_type & annotation != 'NA') %>% group_by(annotation) %>% 
  summarise(mean_presCoexp = mean(presCoexp)) %>% arrange(desc(mean_presCoexp))
  temp_df = filter(org_celltype_df, org_protocol == protocol_type & annotation != 'NA')
  temp_df$annotation = factor(temp_df$annotation, levels = order_df$annotation)
  
  if(useDots){
    p1 = ggplot(temp_df, aes(x = annotation, y = presCoexp)) + geom_point() + ylim(.4, 1) + geom_boxplot() + #geom_violin(scale = 'width') +
      theme(axis.text.x = element_text(angle = 90)) + ggtitle(protocol_type)
  }else{
    p1 = ggplot(temp_df, aes(x = annotation, y = presCoexp)) + ylim(.4, 1) + geom_boxplot() + #geom_violin(scale = 'width') +
      theme(axis.text.x = element_text(angle = 90)) + ggtitle(protocol_type)    
  }
  
  return(p1)
}

celltypes_to_plot = c('CB granule','Dop. MB','Forebrain MEIS2+ GABA','Forebrain NKX2-1+ GABA','GABA CB','vGlut-1/2 Telencephalon','vGlut-2 MB','vGlut-2 Thal', 
                      'Astroglia', 'Floor plate', 'Choroid/Ependymal', 'Thal CRH')

p1 = plot_celltype_presCoexp(filter(org_celltype_presCoexp_df, annotation %in% celltypes_to_plot), all_protocols[1])
p1
p1 = plot_celltype_presCoexp(filter(org_celltype_presCoexp_df, annotation %in% celltypes_to_plot), all_protocols[2])
p1
p1 = plot_celltype_presCoexp(filter(org_celltype_presCoexp_df, annotation %in% celltypes_to_plot), all_protocols[3])
p1
p1 = plot_celltype_presCoexp(filter(org_celltype_presCoexp_df, annotation %in% celltypes_to_plot), all_protocols[4])
p1
p1 = plot_celltype_presCoexp(filter(org_celltype_presCoexp_df, annotation %in% celltypes_to_plot), all_protocols[5])
p1
p1 = plot_celltype_presCoexp(filter(org_celltype_presCoexp_df, annotation %in% celltypes_to_plot), all_protocols[6])
p1
p1 = plot_celltype_presCoexp(filter(org_celltype_presCoexp_df, annotation %in% celltypes_to_plot), all_protocols[7])
p1
p1 = plot_celltype_presCoexp(filter(org_celltype_presCoexp_df, annotation %in% celltypes_to_plot),all_protocols[8])
p1
p1 = plot_celltype_presCoexp(filter(org_celltype_presCoexp_df, annotation %in% celltypes_to_plot), all_protocols[9])
p1
p1 = plot_celltype_presCoexp(filter(org_celltype_presCoexp_df, annotation %in% celltypes_to_plot), all_protocols[10])
p1
p1 = plot_celltype_presCoexp(filter(org_celltype_presCoexp_df, annotation %in% celltypes_to_plot),all_protocols[11])
p1
p1 = plot_celltype_presCoexp(filter(org_celltype_presCoexp_df, annotation %in% celltypes_to_plot), all_protocols[12])
p1
p1 = plot_celltype_presCoexp(filter(org_celltype_presCoexp_df, annotation %in% celltypes_to_plot), all_protocols[13])
p1
p1 = plot_celltype_presCoexp(filter(org_celltype_presCoexp_df, annotation %in% celltypes_to_plot),all_protocols[14])
p1


temp_df = filter(org_celltype_presCoexp_df, annotation %in% c('vGlut-1/2 Telencephalon', 'Dop. MB', 'vGlut-2 MB', 'vGlut-2 Thal') & 
                   org_protocol %in% c('directed dorsal forebrain','directed ventral midbrain') )
temp_df$org_protocol = factor(temp_df$org_protocol, levels = c('directed dorsal forebrain','directed ventral midbrain'))
ggplot(temp_df, aes(x = org_protocol, y = presCoexp, fill = annotation )) + geom_boxplot() + ylim(.4, 1)

#ggsave(filename = 'dorsal_ventral_neuron_boxplots_v2.pdf', device = 'pdf', useDingbats = F, path = 'graphs/cell_type_specific_presCoexp/', height = 3, width = 6)


```




Compare the preserved coexpression of cell-type specific markers to the top 100 metaMarker preserved co-expression



```{r}

table(pasca_cluster_annotations$name_fig2)

```
```{r}
get_celltype_ranked_markers = function(metamarkers, celltype, num_markers, metric = 'rank'){
  
  if(metric != 'rank'){
    celltype_data = filter(metamarkers, cell_type == celltype) %>% arrange(desc(!!as.name(metric)))
    return(celltype_data[1:num_markers, ])}
  else{
    return(filter(metamarkers, cell_type == celltype & rank <= num_markers))
  }
}
```


```{r}

fetal_meta_markers = read_meta_markers("../data/metamarkers/fetal_meta_markers_v2.csv.gz")
fetal_meta_markers


gaba_markers = get_celltype_ranked_markers(fetal_meta_markers, 'GABAergic',100, 'rank')
glut_markers = get_celltype_ranked_markers(fetal_meta_markers, 'Glutamatergic',100, 'rank')
nProg_markers = get_celltype_ranked_markers(fetal_meta_markers, 'Neural_Progenitor',100, 'rank')
nonN_markers = get_celltype_ranked_markers(fetal_meta_markers, 'Non-neuronal',100, 'rank')
intProg_markers = get_celltype_ranked_markers(fetal_meta_markers, 'Intermediate_Progenitor',100, 'rank')
divProg_markers = get_celltype_ranked_markers(fetal_meta_markers, 'Dividing_Progenitor',100, 'rank')
micro_markers = get_celltype_ranked_markers(fetal_meta_markers, 'Microglia/Immune',100, 'rank')

#Ignore duplicates
all_markers = c( intProg_markers$gene, gaba_markers$gene, glut_markers$gene, nProg_markers$gene, divProg_markers$gene, nonN_markers$gene, micro_markers$gene )
dups = all_markers[duplicated(all_markers)]

intProg_markers = intProg_markers %>% filter(!gene %in% dups)
nProg_markers = nProg_markers %>% filter(!gene %in% dups)
divProg_markers = divProg_markers %>% filter(!gene %in% dups)
nonN_markers = nonN_markers %>% filter(!gene %in% dups)
gaba_markers = gaba_markers %>% filter(!gene %in% dups)
glut_markers = glut_markers %>% filter(!gene %in% dups)
micro_markers = micro_markers %>% filter(!gene %in% dups)

dim(intProg_markers)
dim(nProg_markers)
dim(divProg_markers)
dim(nonN_markers)
dim(gaba_markers)
dim(glut_markers)
dim(micro_markers)
```



```{r}
num_markers = 20
cell_type = 'vGlut-1/2 Telencephalon'
cell_type_clusters = filter(pasca_cluster_annotations, name_fig2 == cell_type)$cluster

temp_pres_df = matrix(nrow = nrow(all_sample_meta), ncol = length(cell_type_clusters), dimnames = list(all_sample_meta$sample_ids, cell_type_clusters))

for(i in 1:length(cell_type_clusters)){

  #Grab markers for that cluster
  cluster_markers = filter(pasca_cluster_de, cluster == cell_type_clusters[i] & rank <= num_markers)$gene
  cluster_markers = cluster_markers[!cluster_markers %in% glut_markers$gene]
  if(length(cluster_markers) >10){cluster_markers = cluster_markers[1:10]}
  #Index for the average marker expression and the average preserved co-exp aurocs
  gene_index = rownames(all_gene_org_pres_fetal_mat) %in% cluster_markers
  num_go = sum(gene_index)
  print(sprintf('%s: %i markers',cell_type, num_go))
  avg_presCoexp_vec = colMeans(all_gene_org_pres_fetal_mat[gene_index, ], na.rm = T)

  temp_pres_df[ ,i] = avg_presCoexp_vec
  
}
y = rowMeans(temp_pres_df)
x = all_sample_meta$glut_cons_coexp_metric
correlation = cor(x, y, method = 'spearman')
par(pty = 's')
plot(x, y, xlim = c(0,1), ylim = c(0,1), ylab = cell_type, xlab = 'Glutamatergic MetaMarkers', main = correlation )
abline(a = 0, b = 1, col = 'red')

pdf(file = 'graphs/cell_type_specific_presCoexp/vGlut_telencephalon_scatter_v2.pdf', useDingbats = F, height = 4, width = 4)
par(pty = 's')
plot(x, y, xlim = c(0,1), ylim = c(0,1), ylab = cell_type, xlab = 'Glutamatergic MetaMarkers', main = correlation )
abline(a = 0, b = 1, col = 'red')
dev.off()



cell_type = 'vGlut-2 MB'
cell_type_clusters = filter(pasca_cluster_annotations, name_fig2 == cell_type)$cluster

temp_pres_df = matrix(nrow = nrow(all_sample_meta), ncol = length(cell_type_clusters), dimnames = list(all_sample_meta$sample_ids, cell_type_clusters))

for(i in 1:length(cell_type_clusters)){

  #Grab markers for that cluster
  cluster_markers = filter(pasca_cluster_de, cluster == cell_type_clusters[i] & rank <= num_markers)$gene
  cluster_markers = cluster_markers[!cluster_markers %in% glut_markers$gene]
  if(length(cluster_markers) >10){cluster_markers = cluster_markers[1:10]}
  #Index for the average marker expression and the average preserved co-exp aurocs
  gene_index = rownames(all_gene_org_pres_fetal_mat) %in% cluster_markers
  num_go = sum(gene_index)
  print(sprintf('%s: %i markers',cell_type, num_go))
  avg_presCoexp_vec = colMeans(all_gene_org_pres_fetal_mat[gene_index, ], na.rm = T)

  temp_pres_df[ ,i] = avg_presCoexp_vec
  
}
y = rowMeans(temp_pres_df)
x = all_sample_meta$glut_cons_coexp_metric
correlation = cor(x, y, method = 'spearman')
par(pty = 's')
plot(x, y, xlim = c(0,1), ylim = c(0,1), ylab = cell_type, xlab = 'Glutamatergic MetaMarkers', main = correlation )
abline(a = 0, b = 1, col = 'red')

pdf(file = 'graphs/cell_type_specific_presCoexp/vGlut_midbrain_scatter_v2.pdf', useDingbats = F, height = 4, width = 4)
par(pty = 's')
plot(x, y, xlim = c(0,1), ylim = c(0,1), ylab = cell_type, xlab = 'Glutamatergic MetaMarkers', main = correlation )
abline(a = 0, b = 1, col = 'red')
dev.off()



cell_type = 'vGlut-2 Thal'
cell_type_clusters = filter(pasca_cluster_annotations, name_fig2 == cell_type)$cluster

temp_pres_df = matrix(nrow = nrow(all_sample_meta), ncol = length(cell_type_clusters), dimnames = list(all_sample_meta$sample_ids, cell_type_clusters))

for(i in 1:length(cell_type_clusters)){

  #Grab markers for that cluster
  cluster_markers = filter(pasca_cluster_de, cluster == cell_type_clusters[i] & rank <= num_markers)$gene
  cluster_markers = cluster_markers[!cluster_markers %in% glut_markers$gene]
  if(length(cluster_markers) >10){cluster_markers = cluster_markers[1:10]}
  #Index for the average marker expression and the average preserved co-exp aurocs
  gene_index = rownames(all_gene_org_pres_fetal_mat) %in% cluster_markers
  num_go = sum(gene_index)
  print(sprintf('%s: %i markers',cell_type, num_go))
  avg_presCoexp_vec = colMeans(all_gene_org_pres_fetal_mat[gene_index, ], na.rm = T)

  temp_pres_df[ ,i] = avg_presCoexp_vec
  
}
y = temp_pres_df[ ,1]
x = all_sample_meta$glut_cons_coexp_metric
correlation = cor(x, y, method = 'spearman')
par(pty = 's')
plot(x, y, xlim = c(0,1), ylim = c(0,1), ylab = cell_type, xlab = 'Glutamatergic MetaMarkers', main = correlation )
abline(a = 0, b = 1, col = 'red')

pdf(file = 'graphs/cell_type_specific_presCoexp/vGlut_thalamus_scatter_v2.pdf', useDingbats = F, height = 4, width = 4)
par(pty = 's')
plot(x, y, xlim = c(0,1), ylim = c(0,1), ylab = cell_type, xlab = 'Glutamatergic MetaMarkers', main = correlation )
abline(a = 0, b = 1, col = 'red')
dev.off()


cell_type = 'Forebrain MEIS2+ GABA'
cell_type_clusters = filter(pasca_cluster_annotations, name_fig2 == cell_type)$cluster

temp_pres_df = matrix(nrow = nrow(all_sample_meta), ncol = length(cell_type_clusters), dimnames = list(all_sample_meta$sample_ids, cell_type_clusters))

for(i in 1:length(cell_type_clusters)){

  #Grab markers for that cluster
  cluster_markers = filter(pasca_cluster_de, cluster == cell_type_clusters[i] & rank <= num_markers)$gene
  cluster_markers = cluster_markers[!cluster_markers %in% gaba_markers$gene]
  if(length(cluster_markers) >10){cluster_markers = cluster_markers[1:10]}
  #Index for the average marker expression and the average preserved co-exp aurocs
  gene_index = rownames(all_gene_org_pres_fetal_mat) %in% cluster_markers
  num_go = sum(gene_index)
  print(sprintf('%s: %i markers',cell_type, num_go))
  avg_presCoexp_vec = colMeans(all_gene_org_pres_fetal_mat[gene_index, ], na.rm = T)

  temp_pres_df[ ,i] = avg_presCoexp_vec
  
}
y = temp_pres_df[ ,1]
x = all_sample_meta$gaba_cons_coexp_metric
correlation = cor(x, y, method = 'spearman')
par(pty = 's')
plot(x, y, xlim = c(0,1), ylim = c(0,1), ylab = cell_type, xlab = 'GABAergic MetaMarkers', main = correlation )
abline(a = 0, b = 1, col = 'red')

pdf(file = 'graphs/cell_type_specific_presCoexp/GABA_MEIS2_scatter_v2.pdf', useDingbats = F, height = 4, width = 4)
par(pty = 's')
plot(x, y, xlim = c(0,1), ylim = c(0,1), ylab = cell_type, xlab = 'GABAergic MetaMarkers', main = correlation )
abline(a = 0, b = 1, col = 'red')
dev.off()


cell_type = 'Forebrain NKX2-1+ GABA'
cell_type_clusters = filter(pasca_cluster_annotations, name_fig2 == cell_type)$cluster

temp_pres_df = matrix(nrow = nrow(all_sample_meta), ncol = length(cell_type_clusters), dimnames = list(all_sample_meta$sample_ids, cell_type_clusters))

for(i in 1:length(cell_type_clusters)){

  #Grab markers for that cluster
  cluster_markers = filter(pasca_cluster_de, cluster == cell_type_clusters[i] & rank <= num_markers)$gene
  cluster_markers = cluster_markers[!cluster_markers %in% gaba_markers$gene]
  if(length(cluster_markers) >10){cluster_markers = cluster_markers[1:10]}
  #Index for the average marker expression and the average preserved co-exp aurocs
  gene_index = rownames(all_gene_org_pres_fetal_mat) %in% cluster_markers
  num_go = sum(gene_index)
  print(sprintf('%s: %i markers',cell_type, num_go))
  avg_presCoexp_vec = colMeans(all_gene_org_pres_fetal_mat[gene_index, ], na.rm = T)

  temp_pres_df[ ,i] = avg_presCoexp_vec
  
}
y = rowMeans(temp_pres_df)
x = all_sample_meta$gaba_cons_coexp_metric
correlation = cor(x, y, method = 'spearman')
par(pty = 's')
plot(x, y, xlim = c(0,1), ylim = c(0,1), ylab = cell_type, xlab = 'GABAergic MetaMarkers', main = correlation )
abline(a = 0, b = 1, col = 'red')

pdf(file = 'graphs/cell_type_specific_presCoexp/GABA_NKX2_scatter_v2.pdf', useDingbats = F, height = 4, width = 4)
par(pty = 's')
plot(x, y, xlim = c(0,1), ylim = c(0,1), ylab = cell_type, xlab = 'GABAergic MetaMarkers', main = correlation )
abline(a = 0, b = 1, col = 'red')
dev.off()

```
















