

downsampling fetal data for global fetal preservation of fetal co-expression

Use the annotated fetal dataset with the largest number of cells GW18_2

Downsample 100, 1000, 10000, 100000
Make a co-expression network
Get global conservation of co-expression with the rest of the aggregate fetal co-expression network

```{r}
library(Seurat)
library(reticulate)
library(knitr)
library(matrixStats)
library(data.table)
library(ggplot2)
library(parallel)

Sys.setenv(RETICULATE_PYTHON = "/usr/bin/python3.6")
library(reticulate)
np <- reticulate::import("numpy")

```





```{r}

load('../data/go_annotations/go_annotations.Rdata')
```
grab the aggregate annotated fetal network

```{r}
load('../data/coexpression/fetal/ranked_agg_fetal_coexp_network_8_17_23.Rdata')
dim(fetal_ranked_aggregate)

```


```{r}
source_python('python_coexpression.py')
```


function to make a ranked co-expression matrix from an expression matrix

```{r}

get_ranked_coexp = function(exp_matrix, go_genes){
  
  dataset_genes = rownames(exp_matrix)
  #Filter for the GO gene universe
  dataset_genes_in_go = dataset_genes[dataset_genes %in% go_genes]
  dataset_genes_missing_from_go = go_genes[! go_genes %in% dataset_genes]
  #Subset the gene expression for the genes present in GO
  dataset_gene_index = dataset_genes %in% dataset_genes_in_go
  exp_data = as.matrix(exp_matrix[dataset_gene_index, ])
  #Add zero counts for the missing GO genes
  go_missing = matrix(0, nrow = length(dataset_genes_missing_from_go), ncol = ncol(exp_data), 
                      dimnames = list(dataset_genes_missing_from_go, colnames(exp_data)))
  exp_data = rbind(exp_data, go_missing)
  r_genes = rownames(exp_data)
  
  
  #Get coexpression data
  test_py_exp = np$array(exp_data) #Turn R matrix into numpy array
  rm(exp_data)
  gc()
  python_script_test = get_coexpression(test_py_exp, r_genes) #Python method for spearman correlations, coexpression
  corr_matrix = as.matrix(python_script_test) #Turn to matrix
  #Convert Nans to 0, these are genes with 0 expression in any cells
  corr_matrix[!is.finite(corr_matrix)] = 0
  
  
  #Rank the coexpression data
  #Rank everything, I want the minimum rank to be 1, so ties are minimum. Subtracting 1 sets the minimum rank to 0
  rank_corr = frank(c(corr_matrix), ties.method = 'min') - 1
  #Divide by max rank squeezes everything between 0 and 1
  norm_rank_corr = rank_corr / max(rank_corr)
  #Put back into a matrix
  rank_corr_matrix = matrix(norm_rank_corr, nrow = nrow(corr_matrix), ncol = ncol(corr_matrix), 
                            dimnames = list(rownames(corr_matrix), colnames(corr_matrix)))
  diag(rank_corr_matrix) = 1
  rm(rank_corr,norm_rank_corr, corr_matrix )
  gc()
  return(rank_corr_matrix)
  
}


```



preserved co-expression functions


```{r}

compute_auc = function(scores, label){
  label = as.logical(label)
  n1 = as.numeric(sum(label))
  n2 = as.numeric(sum(!label))
  R1 = sum(rank(scores)[label])
  U1 = R1 - n1 * (n1 + 1)/2
  auc = U1/(n1 * n2)
  return(auc)
}

get_conserved_coexp = function(network1, network2, test_gene, num_top = 10){

  #Get the top 10 coexpressed genes
  test_gene_coexp = network1[test_gene, ]
  test_gene_coexp = test_gene_coexp[names(test_gene_coexp) != test_gene]
  top_coexp = names(test_gene_coexp[order(test_gene_coexp, decreasing = T)][1:num_top])
  
  #Calculate the AUC using the organoid data
  scores = network2[test_gene, ]
  scores = scores[names(scores) != test_gene]
  labels = names(scores) %in% top_coexp
  
  #Get the auc
  auc = compute_auc(scores, labels)
  return(auc)
}


```



Load up the GW22T fetal dataset

```{r}

load('../data/seurat_objs/individual/areal_kreigstein_dataset/merged_with_metadata/GW22T.Rdata' )
full_dataset_exp_mat = GW22T_merged_dataset@assays$RNA@data
dim(full_dataset_exp_mat )
rm(GW22T_merged_dataset)
gc()
```

Randomly sample cells, build a co-expression network, and get the global conserved co-expression score from the aggregate
```{r}
top_genes = 10
num_cells = ncol(full_dataset_exp_mat)
cell_sample_vec = c(100,500,1000,5000,10000,50000,100000)
num_samples = 10

downsample_mat = matrix(nrow = num_samples, ncol = length(cell_sample_vec))
rownames(downsample_mat) = paste('sample', 1:num_samples, sep = ' ')
colnames(downsample_mat) = paste('cells', cell_sample_vec, sep = '_')

for(j in 1:length(cell_sample_vec)){

  for(sample in 1:num_samples){
  
    start = Sys.time()
    #Random sample cells and get co-expression network
    cell_sample = sample(1:num_cells, cell_sample_vec[j])
    sampled_exp = full_dataset_exp_mat[ , cell_sample]
    rank_mat = get_ranked_coexp(sampled_exp, go_genes)
    
    downsample_mat[sample, j ] = mean(unlist(mclapply(1:length(go_genes), 
                                    function(i) get_conserved_coexp(fetal_ranked_aggregate, rank_mat, go_genes[i],top_genes ), mc.cores = 10)))
    
    rm(sampled_exp, rank_mat)
    gc()
    end = Sys.time()
    print(end - start)
  }
    
}


```




```{r}

dim(downsample_mat)

```

```{r}
save(downsample_mat, file = '../data/coexpression/fetal_downsample_consCoexp_fetal_matrix_v2.Rdata')



```


```{r}
load('../data/coexpression/fetal_downsample_consCoexp_fetal_matrix_v2.Rdata')
downsample_mat
```


















