

```{r, setup, include=FALSE}
knitr::opts_knit$set(root.dir = '/home/werner/projects/meta_qc_organoid/code')
```

Starting to explore organoid datasets

```{r}
library(Seurat)
library(data.table)
library(ggplot2)
library(MetaMarkers)
library(dplyr)
library(MetBrewer)
library(tidyr)
library(tibble)
library(parallel)
library(viridis)
library(gridExtra)
library(grid)
library(ComplexHeatmap)
```



organoid dataset metadata
```{r}

all_sample_meta = read.csv('../data/seurat_objs/all_data_just_meta_seurat_with_addl_features_v2.csv')
all_sample_meta
```

Only work with the orghanoid datasets that had raw data available. All of those are CPM. There are a handful of datasets with only log(CPM), log(CPM +1), or various other normalizations available

I double checked all the 'not stated' normalizations were raw counts
```{r}
all_sample_meta
table(all_sample_meta$Normalization.for.this.study)
org_cpm_index = which(all_sample_meta$Normalization.for.this.study == 'CPM')
```

Get a nice bargraph of the organoid datasets with colors

```{r}
#Get palette for the author described protocols
table(all_sample_meta$Author.described.protocol)

sample_region_author_palette = met.brewer('Renoir', n = 19)
names(sample_region_author_palette) = names(table(all_sample_meta$Author.described.protocol))

#Get palette for the grouped described protocols
table(all_sample_meta$Protocol.classification.region)

sample_region_grouped_palette = met.brewer('Renoir', n = 14)
names(sample_region_grouped_palette) = names(table(all_sample_meta$Protocol.classification.region))



```


```{r}
org_samp_df = all_sample_meta %>% count(Author.described.protocol) %>% arrange(desc(n))
org_samp_df$Author.described.protocol = factor(org_samp_df$Author.described.protocol, levels = c(org_samp_df$Author.described.protocol) )
ggplot(org_samp_df, aes(x = Author.described.protocol, y = n, fill = Author.described.protocol)) + geom_bar( stat = 'identity') +
  scale_fill_manual(values =sample_region_author_palette, name = 'Organoid type') +
  ylab('# of datasets') + xlab('Organoid type: Author described protocol') + ggtitle('Human neural organoid scRNA-seq datasets') +
  theme(axis.text.x = element_blank(), axis.ticks.x = element_blank(), 
        legend.key.size = unit(.25, 'cm'))

ggsave(filename = 'organoid_author_proto_num_dataset_barplot_v2.pdf', path = '/home/werner/projects/meta_qc_organoid/code/graphs/', device = 'pdf', useDingbats = F,
       height = 2.5 , width = 4.25 )

org_samp_df = all_sample_meta %>% count(Protocol.classification.region) %>% arrange(desc(n))
org_samp_df$Protocol.classification= factor(org_samp_df$Protocol.classification, levels = c(org_samp_df$Protocol.classification) )
ggplot(org_samp_df, aes(x = Protocol.classification, y = n, fill = Protocol.classification)) + geom_bar( stat = 'identity') +
  scale_fill_manual(values =sample_region_grouped_palette, name = 'Organoid type') +
  ylab('# of datasets') + xlab('Organoid type') + ggtitle('Human neural organoid scRNA-seq datasets') +
  theme(axis.text.x = element_blank(), axis.ticks.x = element_blank(), 
        legend.key.size = unit(.25, 'cm'))


ggsave(filename = 'organoid_grouped_proto_num_dataset_barplot_v2.pdf', path = '/home/werner/projects/meta_qc_organoid/code/graphs/', device = 'pdf', useDingbats = F,
       height = 2.5 , width = 4.25 )
```





Fetal markers

```{r}
get_celltype_ranked_markers = function(metamarkers, celltype, num_markers, metric = 'rank'){
  if(metric != 'rank'){
    celltype_data = filter(metamarkers, cell_type == celltype) %>% arrange(desc(!!as.name(metric)))
    return(celltype_data[1:num_markers, ])}
  else{
    return(filter(metamarkers, cell_type == celltype & rank <= num_markers))
  }
}
```

```{r}

fetal_meta_markers = read_meta_markers("../data/metamarkers/fetal_meta_markers_v2.csv.gz")
fetal_meta_markers


gaba_markers = get_celltype_ranked_markers(fetal_meta_markers, 'GABAergic',100, 'rank')
glut_markers = get_celltype_ranked_markers(fetal_meta_markers, 'Glutamatergic',100, 'rank')
nProg_markers = get_celltype_ranked_markers(fetal_meta_markers, 'Neural_Progenitor',100, 'rank')
nonN_markers = get_celltype_ranked_markers(fetal_meta_markers, 'Non-neuronal',100, 'rank')
intProg_markers = get_celltype_ranked_markers(fetal_meta_markers, 'Intermediate_Progenitor',100, 'rank')
divProg_markers = get_celltype_ranked_markers(fetal_meta_markers, 'Dividing_Progenitor',100, 'rank')
micro_markers = get_celltype_ranked_markers(fetal_meta_markers, 'Microglia/Immune',100, 'rank')

#Ignore duplicates
all_markers = c( intProg_markers$gene, gaba_markers$gene, glut_markers$gene, nProg_markers$gene, divProg_markers$gene, nonN_markers$gene, micro_markers$gene )
dups = all_markers[duplicated(all_markers)]

intProg_markers = intProg_markers %>% filter(!gene %in% dups)
nProg_markers = nProg_markers %>% filter(!gene %in% dups)
divProg_markers = divProg_markers %>% filter(!gene %in% dups)
nonN_markers = nonN_markers %>% filter(!gene %in% dups)
gaba_markers = gaba_markers %>% filter(!gene %in% dups)
glut_markers = glut_markers %>% filter(!gene %in% dups)
micro_markers = micro_markers %>% filter(!gene %in% dups)

dim(intProg_markers)
dim(nProg_markers)
dim(divProg_markers)
dim(nonN_markers)
dim(gaba_markers)
dim(glut_markers)
dim(micro_markers)
```


Seems that two major axes of variation in organoid data is neuron vs progenitor/non neuronal and then maybe non neuronal vs progenitor/neuron

check out the genes associated with the first 2 PCs for the datasets



```{r}
dataset_num =6
load(all_sample_meta$processed_seurat_path[dataset_num])
pca_org = DimPlot(dataset, reduction = 'pca')
pca_org

```

Get some aesthetic values, colors, alphas
Get the top X fetal markers
```{r}
#Get colors
class_palette = met.brewer("Archambault", 20)
class_marker_palette = c('Fetal Non-neuronal marker' = class_palette[20], 'Fetal GABAergic marker' = class_palette[1], 
                         'Fetal Glutamatergic marker' = class_palette[14], 
                         'Fetal Neural Progenitor marker' = class_palette[4],
                         'Fetal Intermediate Progenitor marker' = class_palette[16],
                         'Fetal Dividing Progenitor marker' = class_palette[10],
                         'Fetal Microglia/Immune marker' = class_palette[18],
                         'non-marker' = 'white')

alpha_vector = c('non-marker' = .25,'Fetal GABAergic marker' = 1,'Fetal Glutamatergic marker' = 1,'Fetal Neural Progenitor marker' = 1,'Fetal Non-neuronal marker' = 1, 
                 'Fetal Dividing Progenitor marker' = 1, 'Fetal Intermediate Progenitor marker' = 1, 'Fetal Microglia/Immune marker' = 1)

#plotting the eigenvectors for the first 4 PCs, coloring each gene whether they are a fetal metamarker or not
#the eigenvectors are the weight each gene is given for that principle component, visually exploring whether fetal marker sets are enriched for
#heavily weighted genes in the first few PCs
plot_pc_graphs = function(pc_dataframe, pc_to_plot){
  
  pc_dataframe$genes = factor(pc_dataframe$genes, levels = pc_dataframe$genes[order(pc_dataframe[ ,pc_to_plot])])
  
  g1 = ggplot(pc_dataframe, aes(x = genes, y = !!as.name(pc_to_plot), fill = marker_label, color = marker_label)) + geom_point(aes(alpha = marker_label)) +
    scale_fill_manual(values = class_marker_palette) + scale_color_manual(values = class_marker_palette) + scale_alpha_manual(values = alpha_vector) +
    theme(axis.text.x = element_blank(), axis.ticks.x = element_blank(), legend.position = 'none') + coord_cartesian(clip = 'off')
  
  g2 = ggplot(pc_dataframe, aes(x = marker_label, y = !!as.name(pc_to_plot), fill = marker_label)) + geom_violin(scale = 'width')+ scale_x_discrete(guide = guide_axis(n.dodge = 2)) +
    scale_fill_manual(values = class_marker_palette) + geom_hline(yintercept = 0, color = 'red', linetype = 'dashed') +
    theme(legend.key.height = unit(.25,'cm'),legend.key.width = unit(.25, 'cm'), legend.text = element_text(size =8), axis.text.x = element_blank(), axis.ticks.x = element_blank() )
  
  pc_graph = arrangeGrob(g1, g2, ncol = 2, widths = c(1,2.5))
  return(pc_graph)
}


```




Save PCA plots and the eigenvector plots for all datasets

```{r}
Sys.time()
for(i in 1:nrow(all_sample_meta)){
  
  if(i %% 10 == 0){print(i); print(Sys.time())}
  
  dataset_num =i
  x = load(all_sample_meta$processed_seurat_path[dataset_num])
  dataset = get(x)
  
  pca_org = DimPlot(dataset, reduction = 'pca')
  #ggsave(plot = pca_org, filename = sprintf('org_dataset_%i_pca.pdf', dataset_num), path = '/home/werner/projects/meta_qc_organoid/code/graphs/organoid_pca/', 
  #       device = 'pdf', useDingbats = F, height = 4, width = 5)
  
  df_pc = as.data.frame(dataset@reductions$pca@feature.loadings[, 1:4])
  df_pc$genes = rownames(df_pc)
  
  marker_label = rep('non-marker', nrow(df_pc))
  marker_label[df_pc$genes %in% gaba_markers$gene] = 'Fetal GABAergic marker'
  marker_label[df_pc$genes %in% glut_markers$gene] = 'Fetal Glutamatergic marker'
  marker_label[df_pc$genes %in% nProg_markers$gene] = 'Fetal Neural Progenitor marker'
  marker_label[df_pc$genes %in% nonN_markers$gene] = 'Fetal Non-neuronal marker'
  marker_label[df_pc$genes %in% intProg_markers$gene] = 'Fetal Intermediate Progenitor marker'
  marker_label[df_pc$genes %in% divProg_markers$gene] = 'Fetal Dividing Progenitor marker'
  marker_label[df_pc$genes %in% micro_markers$gene] = 'Fetal Microglia/Immune marker'
  df_pc$marker_label = marker_label
  df_pc$marker_label = factor(df_pc$marker_label, levels = c('non-marker','Fetal GABAergic marker','Fetal Glutamatergic marker','Fetal Neural Progenitor marker','Fetal Non-neuronal marker',
                                                             'Fetal Intermediate Progenitor marker','Fetal Dividing Progenitor marker', 'Fetal Microglia/Immuns marker'))
  
  
  pc_1_graph = plot_pc_graphs(df_pc, 'PC_1')
  pc_2_graph = plot_pc_graphs(df_pc, 'PC_2')
  pc_3_graph = plot_pc_graphs(df_pc, 'PC_3')
  pc_4_graph = plot_pc_graphs(df_pc, 'PC_4')
  
  grid_1 = arrangeGrob(pc_1_graph, pc_2_graph, top = textGrob(sprintf("Organoid dataset %i", dataset_num),gp=gpar(fontsize=15)))
 # ggsave(plot = grid_1, filename = sprintf('pca_1_2_fetalMarker_weights_orgDataset_%i.pdf', dataset_num), path = '/home/werner/projects/meta_qc_organoid/code/graphs/organoid_pca/', 
  #       device = 'pdf', useDingbats = F, height = 5, width = 7)

  grid_2 = arrangeGrob(pc_3_graph, pc_4_graph, top = textGrob(sprintf("Organoid dataset %i", dataset_num),gp=gpar(fontsize=15)))
  #ggsave(plot = grid_2, filename = sprintf('pca_3_4_fetalMarker_weights_orgDataset_%i.pdf', dataset_num), path = '/home/werner/projects/meta_qc_organoid/code/graphs/organoid_pca/', 
  #       device = 'pdf', useDingbats = F, height = 5, width = 7)

}
```


```{r}


plot(grid_1)

```




Comparing HVG genes across organoid datasets,
Comparing eigenvectors across organoid datasets
Maybe thinking of a metaPCA approach
Learn a meta-eigen space from fetal and project the organoids, vice versa



Pilot experiment
gather all the HVG genes across some datasets
get a measure of the magnitude of the eigenvector value for each gene in PC1
abs(eigenvector), maybe rank them, or divide by max(abs(eigenvector))
repeat across datasets
genes by datasets, with value being the magnitude of PC1 weights. 
Idea is to see if genes are consistently heavily weighted across the datasets
Is there any structure across datasets, do some always weight the same genes
are the consistently heavily weighted genes enriched for fetal class markers?


```{r}
combined_hvg = c()

for(i in 1:nrow(all_sample_meta)){

  x = load(all_sample_meta$processed_seurat_path[i])
  current_dataset = get(x)
  suppressWarnings(rm(seurat_object, dataset, x))
  gc()
  
  combined_hvg = c(combined_hvg, current_dataset@assays$RNA@var.features)
}

combined_hvg = unique(combined_hvg)
#empty matrices to hold the weights of PC1 and PC2 for all the datasets
hvg_pc1_eigenvec_matrix = matrix(nrow = length(combined_hvg), ncol = nrow(all_sample_meta))
rownames(hvg_pc1_eigenvec_matrix) = combined_hvg
hvg_pc2_eigenvec_matrix = matrix(nrow = length(combined_hvg), ncol = nrow(all_sample_meta))
rownames(hvg_pc2_eigenvec_matrix) = combined_hvg
hvg_pc3_eigenvec_matrix = matrix(nrow = length(combined_hvg), ncol = nrow(all_sample_meta))
rownames(hvg_pc3_eigenvec_matrix) = combined_hvg
hvg_pc4_eigenvec_matrix = matrix(nrow = length(combined_hvg), ncol = nrow(all_sample_meta))
rownames(hvg_pc4_eigenvec_matrix) = combined_hvg
```

```{r}

dim(hvg_pc1_eigenvec_matrix)
dim(hvg_pc2_eigenvec_matrix)
dim(hvg_pc3_eigenvec_matrix)
dim(hvg_pc4_eigenvec_matrix)
```


```{r}

for(i in 1:nrow(all_sample_meta)){

  x = load(all_sample_meta$processed_seurat_path[i])
  current_dataset = get(x)
  suppressWarnings(rm(seurat_object, dataset, x))
  gc()

  #Get the eigenvector for the first PC
  pc1_eigenvector = current_dataset@reductions$pca@feature.loadings[ ,'PC_1' ]
  pc2_eigenvector = current_dataset@reductions$pca@feature.loadings[ ,'PC_2' ]
  pc3_eigenvector = current_dataset@reductions$pca@feature.loadings[ ,'PC_3' ]
  pc4_eigenvector = current_dataset@reductions$pca@feature.loadings[ ,'PC_4' ]
  #Normalize the values, getting the magnitudes of the weights
  norm_pc1_eigenvector = abs(pc1_eigenvector) / max(abs(pc1_eigenvector))
  norm_pc2_eigenvector = abs(pc2_eigenvector) / max(abs(pc2_eigenvector))
  norm_pc3_eigenvector = abs(pc3_eigenvector) / max(abs(pc3_eigenvector))
  norm_pc4_eigenvector = abs(pc4_eigenvector) / max(abs(pc4_eigenvector))
  #Grab the features, gene names
  features = rownames(current_dataset@reductions$pca@feature.loadings)
  #put it in the matrix
  index = match(features, rownames(hvg_pc1_eigenvec_matrix))
  hvg_pc1_eigenvec_matrix[index, i ] = norm_pc1_eigenvector
  hvg_pc2_eigenvec_matrix[index, i ] = norm_pc2_eigenvector
  hvg_pc3_eigenvec_matrix[index, i ] = norm_pc3_eigenvector
  hvg_pc4_eigenvec_matrix[index, i ] = norm_pc4_eigenvector
  print(i)
}
```


```{r}

table(rowSums(!is.na(hvg_pc1_eigenvec_matrix)) == 1)
table(rowSums(!is.na(hvg_pc2_eigenvec_matrix)) == 1)
```





```{r}
Heatmap(hvg_pc1_eigenvec_matrix, cluster_rows = F, cluster_columns = F, show_row_names = F)
Heatmap(hvg_pc2_eigenvec_matrix, cluster_rows = F, cluster_columns = F, show_row_names = F)
Heatmap(hvg_pc3_eigenvec_matrix, cluster_rows = F, cluster_columns = F, show_row_names = F)
Heatmap(hvg_pc4_eigenvec_matrix, cluster_rows = F, cluster_columns = F, show_row_names = F)
```


```{r}
save(hvg_pc1_eigenvec_matrix, hvg_pc2_eigenvec_matrix, hvg_pc3_eigenvec_matrix, hvg_pc4_eigenvec_matrix,file = '../data/organoid_pca/org_eigenvec_matrices_v3.Rdata' )
```

```{r}
load('../data/organoid_pca/org_eigenvec_matrices_v3.Rdata')

```


```{r}
library(circlize)

```

Color palette for the different brain regions across the datasets, mostly capturing protocol variance
```{r}

#sample_region_palette = met.brewer('Renoir', n = 12)
#sample_region_palette = c('brainstem' = sample_region_palette[1],'cerebellum' = sample_region_palette[2],'cerebral' = sample_region_palette[3],'cortical' = sample_region_palette[4],
#                          'cortical_and_neural_retina' = sample_region_palette[5],'dorsal_patterned_forebrain' = sample_region_palette[6],'hypothalamic_arcuate' = sample_region_palette[7],
#                          'MGE_and_cortical' = sample_region_palette[8], 'neural_induced_blood_vessel' = sample_region_palette[9],'thalamic' = sample_region_palette[10],
#                          'vascularized_cortical' = sample_region_palette[11],'ventral_midbrain' = sample_region_palette[12])

sample_region_grouped_palette
```


```{r}
#color for the weights
col_fun = colorRamp2(c(0, .5, 1), c(inferno(3)[1], inferno(3)[2], inferno(3)[3]))

#Turn NAs to 0s and filter out genes that are not present in at least ten dataset
temp_hvg_pc_eigenvec_matrix = hvg_pc1_eigenvec_matrix
temp_hvg_pc_eigenvec_matrix[is.na(temp_hvg_pc_eigenvec_matrix)] = 0
temp_hvg_pc_eigenvec_matrix = temp_hvg_pc_eigenvec_matrix[rowSums(temp_hvg_pc_eigenvec_matrix != 0) >= 10 , ]


#filter out the non markers, just making the point that fetal markers are consistently weighted in the first PC across organoid datasets
keep_index = rownames(temp_hvg_pc_eigenvec_matrix) %in% c(gaba_markers$gene, glut_markers$gene, nProg_markers$gene, nonN_markers$gene, intProg_markers$gene, divProg_markers$gene)
temp_hvg_pc_eigenvec_matrix = temp_hvg_pc_eigenvec_matrix[keep_index, ]

#Get marker annotations
fetal_marker_annot_vec = rep('non-marker', length = nrow(temp_hvg_pc_eigenvec_matrix))
fetal_marker_annot_vec[rownames(temp_hvg_pc_eigenvec_matrix) %in% intProg_markers$gene] = 'Fetal Intermediate Progenitor marker' 
fetal_marker_annot_vec[rownames(temp_hvg_pc_eigenvec_matrix) %in% gaba_markers$gene] = 'Fetal GABAergic marker' 
fetal_marker_annot_vec[rownames(temp_hvg_pc_eigenvec_matrix) %in% glut_markers$gene] = 'Fetal Glutamatergic marker' 
fetal_marker_annot_vec[rownames(temp_hvg_pc_eigenvec_matrix) %in% nProg_markers$gene] = 'Fetal Neural Progenitor marker' 
fetal_marker_annot_vec[rownames(temp_hvg_pc_eigenvec_matrix) %in% nonN_markers$gene] = 'Fetal Non-neuronal marker'
fetal_marker_annot_vec[rownames(temp_hvg_pc_eigenvec_matrix) %in% divProg_markers$gene] = 'Fetal Dividing Progenitor marker' 

ha = rowAnnotation(fetal_marker = fetal_marker_annot_vec, 
                       col = list(fetal_marker = class_marker_palette))
dataset_annot = HeatmapAnnotation(organoid = all_sample_meta$Protocol.classification, 
                                  col = list(organoid = sample_region_grouped_palette))

Heatmap(temp_hvg_pc_eigenvec_matrix, cluster_rows = T, cluster_columns = T, show_row_names = F, col = col_fun, name = 'PC1 weights', 
        clustering_method_rows = 'ward.D2', clustering_method_columns = 'ward.D2', left_annotation = ha, top_annotation = dataset_annot)


pdf(file = 'graphs/organoid_pca/pc1_weight_heatmap_v4.pdf', useDingbats = F, height = 6, width = 10)
Heatmap(temp_hvg_pc_eigenvec_matrix, cluster_rows = T, cluster_columns = T, show_row_names = F, col = col_fun, name = 'PC1 weights', 
        clustering_method_rows = 'ward.D2', clustering_method_columns = 'ward.D2', left_annotation = ha, top_annotation = dataset_annot, use_raster = T)
dev.off()


```



violin plots of PC1 weights for each class

```{r}
temp_hvg_pc_eigenvec_matrix = hvg_pc1_eigenvec_matrix
temp_hvg_pc_eigenvec_matrix[is.na(temp_hvg_pc_eigenvec_matrix)] = 0
temp_hvg_pc_eigenvec_matrix = temp_hvg_pc_eigenvec_matrix[rowSums(temp_hvg_pc_eigenvec_matrix != 0) >= 10 , ]

test_df = reshape2::melt(temp_hvg_pc_eigenvec_matrix)
test_df$marker_class = rep('non-marker', nrow(test_df))
test_df$marker_class[test_df$Var1 %in% intProg_markers$gene] = 'Fetal Intermediate Progenitor marker'
test_df$marker_class[test_df$Var1 %in% gaba_markers$gene] = 'Fetal GABAergic marker'
test_df$marker_class[test_df$Var1 %in% glut_markers$gene] = 'Fetal Glutamatergic marker'
test_df$marker_class[test_df$Var1 %in% nProg_markers$gene] = 'Fetal Neural Progenitor marker'
test_df$marker_class[test_df$Var1 %in% nonN_markers$gene] = 'Fetal Non-neuronal marker'
test_df$marker_class[test_df$Var1 %in% divProg_markers$gene] = 'Fetal Dividing Progenitor marker'

test_df$marker_class = factor(test_df$marker_class, levels = c('Fetal Dividing Progenitor marker', 'Fetal Neural Progenitor marker', 'Fetal Glutamatergic marker',
                                                              'Fetal Intermediate Progenitor marker','Fetal Non-neuronal marker',
                                                               'Fetal GABAergic marker','non-marker' ))

length(unique(filter(test_df, marker_class == 'Fetal GABAergic marker')$Var1))
length(unique(filter(test_df, marker_class == 'Fetal Glutamatergic marker')$Var1))
length(unique(filter(test_df, marker_class == 'Fetal Neural Progenitor marker')$Var1))
length(unique(filter(test_df, marker_class == 'Fetal Non-neuronal marker')$Var1))
length(unique(filter(test_df, marker_class == 'Fetal Intermediate Progenitor marker')$Var1))
length(unique(filter(test_df, marker_class == 'Fetal Dividing Progenitor marker')$Var1))

ggplot(test_df, aes(x = marker_class, y = value, fill = marker_class)) + geom_violin(scale = 'width', alpha = .5) + geom_boxplot(outlier.shape = NA, width = .2) +
  scale_x_discrete(guide = guide_axis(n.dodge = 2)) + scale_fill_manual(values = class_marker_palette) +
  ylab('Organoid PC1 weights') + ggtitle('Top fetal brain markers define PC1 in organoid data')


ggsave(filename = 'organoid_pc1_condensed_boxplots_v4.pdf', path = 'graphs/organoid_pca/', 
       device = 'pdf', useDingbats = F, height = 4, width = 6)


```


Now look at the fraction of datasets a gene is determined as 'heavily weighted' in PC1, with varying thresholds for what is 'heavily weighted'
Compare across the metamarker genes and the non-marker genes


```{r}

#Turn NAs to 0s and filter out genes that are not present in at least ten dataset
temp_hvg_pc_eigenvec_matrix = hvg_pc1_eigenvec_matrix
temp_hvg_pc_eigenvec_matrix[is.na(temp_hvg_pc_eigenvec_matrix)] = 0
temp_hvg_pc_eigenvec_matrix = temp_hvg_pc_eigenvec_matrix[rowSums(temp_hvg_pc_eigenvec_matrix != 0) >= 10 , ]

test_df = reshape2::melt(temp_hvg_pc_eigenvec_matrix)
test_df$marker_class = rep('non-marker', nrow(test_df))
test_df$marker_class[test_df$Var1 %in% intProg_markers$gene] = 'Fetal Intermediate Progenitor marker'
test_df$marker_class[test_df$Var1 %in% gaba_markers$gene] = 'Fetal GABAergic marker'
test_df$marker_class[test_df$Var1 %in% glut_markers$gene] = 'Fetal Glutamatergic marker'
test_df$marker_class[test_df$Var1 %in% nProg_markers$gene] = 'Fetal Neural Progenitor marker'
test_df$marker_class[test_df$Var1 %in% nonN_markers$gene] = 'Fetal Non-neuronal marker'
test_df$marker_class[test_df$Var1 %in% divProg_markers$gene] = 'Fetal Dividing Progenitor marker'

thresh_func = function(pc_thresh, pc_weights){
 return(sum(pc_weights >= pc_thresh)/length(pc_weights))
}
thresholds = seq(0,1,.05)

test_genes = test_df %>% filter(Var2 ==1) %>% pull(Var1)
test_gene_marker_labels = test_df %>% filter(Var2 ==1) %>% pull(marker_class)


all_gene_pc_df = data.frame(gene = c(), weight_threshold = c(), weighted_frac = c(), marker_label = c())
for(i in 1:length(test_genes)){

  gene_pc_weights = test_df %>% filter(Var1 == test_genes[i]) %>% pull(value)
  weighted_frac = sapply(thresholds, thresh_func, pc_weights = gene_pc_weights)
  
  
  gene_pc_df = data.frame(gene = rep(test_genes[i], length = length(thresholds)), weight_threshold = thresholds, weighted_frac = weighted_frac,
                          marker_label = rep(test_gene_marker_labels[i], length = length(thresholds)))
  
  all_gene_pc_df = rbind(all_gene_pc_df, gene_pc_df)

}

all_gene_pc_df


```


```{r}


all_gene_pc_df_plotting = all_gene_pc_df %>% group_by(weight_threshold, marker_label) %>% summarize(mean_frac = mean(weighted_frac), sd_frac = sd(weighted_frac))

save(all_gene_pc_df_plotting, file = '/home/werner/projects/meta_qc_organoid/data/organoid_pca/all_gene_pc_df.Rdata')

ggplot(all_gene_pc_df_plotting, aes(x = weight_threshold, y = mean_frac, group = marker_label, color = marker_label)) +
  geom_line() +
  geom_point() +
  scale_color_manual(values = class_marker_palette) +
  ylab('Fraction of neural organoid datasets gene is heavily weighted in PC1') + xlab('PC1 weight threshold to determine heavily weighted genes')
ggsave(filename = 'dataset_fraction_gene_pc1_weight.pdf', path = 'graphs/organoid_pca/', 
       device = 'pdf', useDingbats = F, height = 4, width = 6)




```






















