

Testing the fetal conservation of fetal data
First look at the unannotated datasets
Then do the annotated, leave one out and reaggregate the coexpression, then test



```{r}
library(Seurat)
library(data.table)
library(ggplot2)
library(MetaMarkers)
library(dplyr)
library(MetBrewer)
library(ComplexHeatmap)
library(circlize)
library(EGAD)
library(parallel)

```



Fetal data
```{r}


human_fetal_dataset_names = c( 'fan_fu','poliodakis_batch1', 'poliodakis_batch2','areal_GW14','areal_GW18_2','areal_GW19','areal_GW19_2', 
                               'areal_GW20','areal_GW20_31','areal_GW20_34','areal_GW25')

human_fetal_dataset_paths = list('../data/seurat_objs/individual/fan_fu_seurat_processed.Rdata',
                                 '../data/seurat_objs/individual/poliodakis_geschwind_batch1_seurat_processed.Rdata',
                                 '../data/seurat_objs/individual/poliodakis_geschwind_batch2_seurat_processed.Rdata',
                                 '../data/seurat_objs/individual/areal_kreigstein_dataset/merged_with_metadata/GW14.Rdata', 
                                 '../data/seurat_objs/individual/areal_kreigstein_dataset/merged_with_metadata/GW18_2.Rdata', 
                                 '../data/seurat_objs/individual/areal_kreigstein_dataset/merged_with_metadata/GW19.Rdata', 
                                 '../data/seurat_objs/individual/areal_kreigstein_dataset/merged_with_metadata/GW19_2.Rdata', 
                                 '../data/seurat_objs/individual/areal_kreigstein_dataset/merged_with_metadata/GW20.Rdata', 
                                 '../data/seurat_objs/individual/areal_kreigstein_dataset/merged_with_metadata/GW20_31.Rdata', 
                                 '../data/seurat_objs/individual/areal_kreigstein_dataset/merged_with_metadata/GW20_34.Rdata',
                                 '../data/seurat_objs/individual/areal_kreigstein_dataset/merged_with_metadata/GW25.Rdata')

names(human_fetal_dataset_paths)= human_fetal_dataset_names 
human_fetal_dataset_paths


unannot_human_fetal_dataset_names = c('areal_GW16','areal_GW18', 'areal_GW22','areal_GW22T','shi_wang','zhou_lu',
                                      'yu_sun_1','yu_sun_2','yu_sun_3','yu_sun_4',
                                      'trevino_greenleaf_1','trevino_greenleaf_2','trevino_greenleaf_3','trevino_greenleaf_4')

unannot_human_fetal_dataset_paths = list(
                                 '../data/seurat_objs/individual/areal_kreigstein_dataset/merged_with_metadata/GW16.Rdata',
                                 '../data/seurat_objs/individual/areal_kreigstein_dataset/merged_with_metadata/GW18.Rdata', 
                                 '../data/seurat_objs/individual/areal_kreigstein_dataset/merged_with_metadata/GW22.Rdata',
                                 '../data/seurat_objs/individual/areal_kreigstein_dataset/merged_with_metadata/GW22T.Rdata',
                                 '../data/seurat_objs/individual/shi_wang_seurat_processed.Rdata', 
                                 '../data/seurat_objs/individual/zhou_lu_seurat_processed.Rdata',
                                 '../data/seurat_objs/individual/yu_sun_GSM5032680_seurat_processed.Rdata',
                                 '../data/seurat_objs/individual/yu_sun_GSM5032681_seurat_processed.Rdata',
                                 '../data/seurat_objs/individual/yu_sun_GSM5032682_seurat_processed.Rdata',
                                 '../data/seurat_objs/individual/yu_sun_GSM5032683_seurat_processed.Rdata',
                                 '../data/seurat_objs/individual/trevino_greenleaf_batch1_seurat_processed.Rdata',
                                 '../data/seurat_objs/individual/trevino_greenleaf_batch2_seurat_processed.Rdata',
                                 '../data/seurat_objs/individual/trevino_greenleaf_batch3_seurat_processed.Rdata',
                                 '../data/seurat_objs/individual/trevino_greenleaf_batch4_seurat_processed.Rdata')

names(unannot_human_fetal_dataset_paths)= unannot_human_fetal_dataset_names 
unannot_human_fetal_dataset_paths

```



```{r}

#Get colors
class_palette = met.brewer("Archambault", 20)
class_marker_palette = c('Fetal Non-neuronal marker' = class_palette[20], 'Fetal GABAergic marker' = class_palette[1], 
                         'Fetal Glutamatergic marker' = class_palette[14], 
                         'Fetal Neural Progenitor marker' = class_palette[4],
                         'Fetal Intermediate Progenitor marker' = class_palette[16],
                         'Fetal Dividing Progenitor marker' = class_palette[10],
                         'non-marker' = 'white')

class_celltype_palette = c('Non-neuronal' = class_palette[20], 'GABAergic' = class_palette[1], 
                         'Glutamatergic' = class_palette[14], 
                         'Neural_Progenitor' = class_palette[4],
                         'Intermediate_Progenitor' = class_palette[16],
                         'Dividing_Progenitor' = class_palette[10],
                         'other' = 'white')

```


```{r}
get_celltype_ranked_markers = function(metamarkers, celltype, num_markers, metric = 'rank'){
  
  if(metric != 'rank'){
    celltype_data = filter(metamarkers, cell_type == celltype) %>% arrange(desc(!!as.name(metric)))
    return(celltype_data[1:num_markers, ])}
  else{
    return(filter(metamarkers, cell_type == celltype & rank <= num_markers))
  }
}
```


```{r}

test_markers = list(
    fan_fu = read_markers("6_26_24_annot_metaMarkers/fan_fu_fetal_class_markers.csv.gz"), 
    poliodakis_1 = read_markers("6_26_24_annot_metaMarkers/poliodakis_batch1_fetal_class_markers.csv.gz"),
    poliodakis_2 = read_markers("6_26_24_annot_metaMarkers/poliodakis_batch2_fetal_class_markers.csv.gz"),
    GW14 = read_markers("6_26_24_annot_metaMarkers/areal_GW14_fetal_class_markers.csv.gz"),
    GW18_2 = read_markers("6_26_24_annot_metaMarkers/areal_GW18_2_fetal_class_markers.csv.gz"),
    GW19 = read_markers("6_26_24_annot_metaMarkers/areal_GW19_fetal_class_markers.csv.gz"),
    GW19_2 = read_markers("6_26_24_annot_metaMarkers/areal_GW19_2_fetal_class_markers.csv.gz"),
    GW20 = read_markers("6_26_24_annot_metaMarkers/areal_GW20_fetal_class_markers.csv.gz"),
    GW20_31 = read_markers("6_26_24_annot_metaMarkers/areal_GW20_31_fetal_class_markers.csv.gz"),
    GW20_34 = read_markers("6_26_24_annot_metaMarkers/areal_GW20_34_fetal_class_markers.csv.gz"),
    GW25 = read_markers("6_26_24_annot_metaMarkers/areal_GW25_fetal_class_markers.csv.gz"),
    linnarsson_GW5 = read_markers("6_26_24_linnarsson_metaMarkers/linnarsson_GW5_fetal_class_markers_v4.csv.gz"),
    linnarsson_GW5_point_5 = read_markers("6_26_24_linnarsson_metaMarkers/linnarsson_GW5-5_fetal_class_markers_v4.csv.gz"),
    linnarsson_GW6_1 = read_markers("6_26_24_linnarsson_metaMarkers/linnarsson_GW6_1_fetal_class_markers_v4.csv.gz"),
    linnarsson_GW6_2 = read_markers("6_26_24_linnarsson_metaMarkers/linnarsson_GW6_2_fetal_class_markers_v4.csv.gz"),
    linnarsson_GW6_point_6_1 = read_markers("6_26_24_linnarsson_metaMarkers/linnarsson_GW6-6_1_fetal_class_markers_v4.csv.gz"),
    linnarsson_GW6_point_6_2 = read_markers("6_26_24_linnarsson_metaMarkers/linnarsson_GW6-6_2_fetal_class_markers_v4.csv.gz"),
    linnarsson_GW6_point_7 = read_markers("6_26_24_linnarsson_metaMarkers/linnarsson_GW6-7_fetal_class_markers_v4.csv.gz"),
    linnarsson_GW6_point_9_1 = read_markers("6_26_24_linnarsson_metaMarkers/linnarsson_GW6-9_1_fetal_class_markers_v4.csv.gz"),
    linnarsson_GW6_point_9_2 = read_markers("6_26_24_linnarsson_metaMarkers/linnarsson_GW6-9_2_fetal_class_markers_v4.csv.gz"),
    linnarsson_GW6_point_9_3 = read_markers("6_26_24_linnarsson_metaMarkers/linnarsson_GW6-9_3_fetal_class_markers_v4.csv.gz"),
    linnarsson_GW6_point_9_4 = read_markers("6_26_24_linnarsson_metaMarkers/linnarsson_GW6-9_4_fetal_class_markers_v4.csv.gz"), 
    linnarsson_GW7 = read_markers("6_26_24_linnarsson_metaMarkers/linnarsson_GW7_fetal_class_markers_v4.csv.gz"),
    linnarsson_GW7_point_5 = read_markers("6_26_24_linnarsson_metaMarkers/linnarsson_GW7-5_fetal_class_markers_v4.csv.gz"),
    linnarsson_GW8_1 = read_markers("6_26_24_linnarsson_metaMarkers/linnarsson_GW8_1_fetal_class_markers_v4.csv.gz"),
    linnarsson_GW8_2 = read_markers("6_26_24_linnarsson_metaMarkers/linnarsson_GW8_2_fetal_class_markers_v4.csv.gz"),
    linnarsson_GW8_3 = read_markers("6_26_24_linnarsson_metaMarkers/linnarsson_GW8_3_fetal_class_markers_v4.csv.gz"),
    linnarsson_GW8_point_1 = read_markers("6_26_24_linnarsson_metaMarkers/linnarsson_GW8-1_fetal_class_markers_v4.csv.gz"),
    linnarsson_GW8_point_5_1 = read_markers("6_26_24_linnarsson_metaMarkers/linnarsson_GW8-5_1_fetal_class_markers_v4.csv.gz"),
    linnarsson_GW8_point_5_2 = read_markers("6_26_24_linnarsson_metaMarkers/linnarsson_GW8-5_2_fetal_class_markers_v4.csv.gz"),
    linnarsson_GW9_point_2 = read_markers("6_26_24_linnarsson_metaMarkers/linnarsson_GW9-2_fetal_class_markers_v4.csv.gz"),
    linnarsson_GW9_point_5 = read_markers("6_26_24_linnarsson_metaMarkers/linnarsson_GW9-5_fetal_class_markers_v4.csv.gz"),
    linnarsson_GW10 = read_markers("6_26_24_linnarsson_metaMarkers/linnarsson_GW10_fetal_class_markers_v4.csv.gz"),
    linnarsson_GW11_point_5 = read_markers("6_26_24_linnarsson_metaMarkers/linnarsson_GW11-5_fetal_class_markers_v4.csv.gz"),
    linnarsson_GW12 = read_markers("6_26_24_linnarsson_metaMarkers/linnarsson_GW12_fetal_class_markers_v4.csv.gz"),
    linnarsson_GW13 = read_markers("6_26_24_linnarsson_metaMarkers/linnarsson_GW13_fetal_class_markers_v4.csv.gz"),
    linnarsson_GW14 = read_markers("6_26_24_linnarsson_metaMarkers/linnarsson_GW14_fetal_class_markers_v4.csv.gz")
    )

fetal_meta_markers = make_meta_markers(test_markers, detailed_stats = TRUE)

gaba_markers = get_celltype_ranked_markers(fetal_meta_markers, 'GABAergic',100, 'rank')
glut_markers = get_celltype_ranked_markers(fetal_meta_markers, 'Glutamatergic',100, 'rank')
nProg_markers = get_celltype_ranked_markers(fetal_meta_markers, 'Neural_Progenitor',100, 'rank')
nonN_markers = get_celltype_ranked_markers(fetal_meta_markers, 'Non-neuronal',100, 'rank')
intProg_markers = get_celltype_ranked_markers(fetal_meta_markers, 'Intermediate_Progenitor',100, 'rank')
divProg_markers = get_celltype_ranked_markers(fetal_meta_markers, 'Dividing_Progenitor',100, 'rank')
micro_markers = get_celltype_ranked_markers(fetal_meta_markers, 'Microglia/Immune',100, 'rank')

#Ignore duplicates
all_markers = c( intProg_markers$gene, gaba_markers$gene, glut_markers$gene, nProg_markers$gene, divProg_markers$gene, nonN_markers$gene, micro_markers$gene )
dups = all_markers[duplicated(all_markers)]

intProg_markers = intProg_markers %>% filter(!gene %in% dups)
nProg_markers = nProg_markers %>% filter(!gene %in% dups)
divProg_markers = divProg_markers %>% filter(!gene %in% dups)
nonN_markers = nonN_markers %>% filter(!gene %in% dups)
gaba_markers = gaba_markers %>% filter(!gene %in% dups)
glut_markers = glut_markers %>% filter(!gene %in% dups)
micro_markers = micro_markers %>% filter(!gene %in% dups)

```


```{r}

load('../data/coexpression/fetal/ranked_agg_fetal_coexp_network_8_17_23.Rdata')
dim(fetal_ranked_aggregate)

load(file = '../data/coexpression/fetal/ranked_agg_unannot_fetal_coexp_network_8_17_23.Rdata')
dim(unannot_fetal_ranked_aggregate)
```


```{r}
library(org.Hs.eg.db)
library(AnnotationDbi)

#GO database from GOSOURCEDATE: 2023-01-01, it's tied to the version of R. current version of R is 4.3.1
org.Hs.eg.db #running this will print out the GO database version information
go_table <- as.data.frame(org.Hs.egGO2ALLEGS) #gene mapping to the direct GO term and that term's children
go_table$gene_symbol = mapIds(org.Hs.eg.db, go_table$gene_id, "SYMBOL","ENTREZID") #Add gene symbol 

go_genes = unique(go_table$gene_symbol)
length(go_genes)
go_terms = unique(go_table$go_id)
length(go_terms)
```


```{r}
#library(org.Hs.eg.db)
library(EGAD)
#library(AnnotationDbi)
load('../data/go_annotations/go_annotations.Rdata')
```


```{r}
compute_auc = function(scores, label){
  label = as.logical(label)
  n1 = as.numeric(sum(label))
  n2 = as.numeric(sum(!label))
  R1 = sum(rank(scores)[label])
  U1 = R1 - n1 * (n1 + 1)/2
  auc = U1/(n1 * n2)
  return(auc)
}


#Compute ROC curve for plotting
#Input is a list of true/false labels (celltype identity) ranked by some metric (gene expression). 
roc_curve = function(ranked_labels){

  #Get total number of positives and negatives
  num_pos = sum(ranked_labels)
  num_neg = sum(!ranked_labels)
  num_total = length(ranked_labels)
  #Cumulative sum of positives and negatives
  tp = cumsum(ranked_labels)
  fp = cumsum(!ranked_labels)
  
  fpr = fp/num_neg
  tpr = tp/num_pos
  
  return(list('fpr' = fpr, 'tpr' = tpr))
}

get_conserved_coexp = function(network1, network2, test_gene, num_top = 10){

  #Get the top 10 coexpressed genes
  test_gene_coexp = network1[test_gene, ]
  test_gene_coexp = test_gene_coexp[names(test_gene_coexp) != test_gene]
  top_coexp = names(test_gene_coexp[order(test_gene_coexp, decreasing = T)][1:num_top])
  
  #Calculate the AUC using the organoid data
  scores = network2[test_gene, ]
  scores = scores[names(scores) != test_gene]
  labels = names(scores) %in% top_coexp
  
  #Get the auc
  auc = compute_auc(scores, labels)
  return(auc)
}

```


the cell type marker sets filtered for just the GO genes
```{r}

go_nProg = nProg_markers$gene[nProg_markers$gene %in% go_genes]
go_nonN = nonN_markers$gene[nonN_markers$gene %in% go_genes]
go_gaba = gaba_markers$gene[gaba_markers$gene %in% go_genes]
go_glut = glut_markers$gene[glut_markers$gene %in% go_genes]
go_intProg = intProg_markers$gene[intProg_markers$gene %in% go_genes & (!intProg_markers$gene %in% divProg_markers$gene)]
go_divProg = divProg_markers$gene[divProg_markers$gene %in% go_genes]
go_micro = micro_markers$gene[micro_markers$gene %in% go_genes]

length(go_nProg)
length(go_nonN)
length(go_gaba)
length(go_glut)
length(go_divProg)
length(go_intProg)
length(go_micro)


```

```{r}

#fetal coexpression network paths

unannot_ranked_coexp_paths = c('../data/coexpression/fetal/8_17_23_GO_coexp/ranked_shi_wang_coexp.Rdata',
'../data/coexpression/fetal/8_17_23_GO_coexp/ranked_yu_sun_1_coexp.Rdata',
'../data/coexpression/fetal/8_17_23_GO_coexp/ranked_yu_sun_2_coexp.Rdata',
'../data/coexpression/fetal/8_17_23_GO_coexp/ranked_yu_sun_3_coexp.Rdata',
'../data/coexpression/fetal/8_17_23_GO_coexp/ranked_yu_sun_4_coexp.Rdata',
'../data/coexpression/fetal/8_17_23_GO_coexp/ranked_trevino_greenleaf_1_coexp.Rdata',
'../data/coexpression/fetal/8_17_23_GO_coexp/ranked_trevino_greenleaf_2_coexp.Rdata',
'../data/coexpression/fetal/8_17_23_GO_coexp/ranked_trevino_greenleaf_3_coexp.Rdata',
'../data/coexpression/fetal/8_17_23_GO_coexp/ranked_trevino_greenleaf_4_coexp.Rdata',
'../data/coexpression/fetal/8_17_23_GO_coexp/ranked_zhou_lu_coexp.Rdata',
'../data/coexpression/fetal/8_17_23_GO_coexp/ranked_areal_GW16_coexp.Rdata',
'../data/coexpression/fetal/8_17_23_GO_coexp/ranked_areal_GW18_coexp.Rdata',
'../data/coexpression/fetal/8_17_23_GO_coexp/ranked_areal_GW22_coexp.Rdata',
'../data/coexpression/fetal/8_17_23_GO_coexp/ranked_areal_GW22T_coexp.Rdata')


unannot_human_fetal_dataset_names = c('shi_wang', 'yu_sun_1','yu_sun_2','yu_sun_3','yu_sun_4',
                                      'trevino_greenleaf_1','trevino_greenleaf_2','trevino_greenleaf_3','trevino_greenleaf_4',
                                      'zhou_lu','areal_GW16','areal_GW18', 'areal_GW22','areal_GW22T')
```


```{r}

top_genes = 10

nProg_cons_matrix_1 = matrix(nrow = length(go_nProg), ncol = length(unannot_ranked_coexp_paths), dimnames = list(go_nProg, unannot_human_fetal_dataset_names))
intProg_cons_matrix_1 = matrix(nrow = length(go_intProg), ncol = length(unannot_ranked_coexp_paths), dimnames = list(go_intProg, unannot_human_fetal_dataset_names))
divProg_cons_matrix_1 = matrix(nrow = length(go_divProg), ncol = length(unannot_ranked_coexp_paths), dimnames = list(go_divProg, unannot_human_fetal_dataset_names))
glut_cons_matrix_1 = matrix(nrow = length(go_glut), ncol = length(unannot_ranked_coexp_paths), dimnames = list(go_glut, unannot_human_fetal_dataset_names))
gaba_cons_matrix_1 = matrix(nrow = length(go_gaba), ncol = length(unannot_ranked_coexp_paths), dimnames = list(go_gaba, unannot_human_fetal_dataset_names))
nonN_cons_matrix_1 = matrix(nrow = length(go_nonN), ncol = length(unannot_ranked_coexp_paths), dimnames = list(go_nonN, unannot_human_fetal_dataset_names))
micro_cons_matrix_1 = matrix(nrow = length(go_micro), ncol = length(unannot_ranked_coexp_paths), dimnames = list(go_micro, unannot_human_fetal_dataset_names))


for(i in 1:length(unannot_ranked_coexp_paths)){

  load(unannot_ranked_coexp_paths[[i]])
  
  fetal_nProg_cons_coexp = unlist(mclapply(1:length(go_nProg), function(i) get_conserved_coexp(fetal_ranked_aggregate, rank_corr_matrix, go_nProg[i],top_genes ), mc.cores = 10 ))
  nProg_cons_matrix_1[ ,i] = fetal_nProg_cons_coexp 
  
  fetal_intProg_cons_coexp = unlist(mclapply(1:length(go_intProg), function(i) get_conserved_coexp(fetal_ranked_aggregate, rank_corr_matrix, go_intProg[i],top_genes ), mc.cores = 10  ))
  intProg_cons_matrix_1[ ,i] = fetal_intProg_cons_coexp
  
  fetal_divProg_cons_coexp = unlist(mclapply(1:length(go_divProg), function(i) get_conserved_coexp(fetal_ranked_aggregate, rank_corr_matrix, go_divProg[i],top_genes ), mc.cores = 10  ))
  divProg_cons_matrix_1[ ,i] = fetal_divProg_cons_coexp
  
  fetal_glut_cons_coexp = unlist(mclapply(1:length(go_glut), function(i) get_conserved_coexp(fetal_ranked_aggregate, rank_corr_matrix, go_glut[i],top_genes ), mc.cores = 10  ))
  glut_cons_matrix_1[ ,i] = fetal_glut_cons_coexp 
  
  fetal_gaba_cons_coexp = unlist(mclapply(1:length(go_gaba), function(i) get_conserved_coexp(fetal_ranked_aggregate, rank_corr_matrix, go_gaba[i],top_genes ), mc.cores = 10  ))
  gaba_cons_matrix_1[ ,i] = fetal_gaba_cons_coexp 
  
  fetal_nonN_cons_coexp = unlist(mclapply(1:length(go_nonN), function(i) get_conserved_coexp(fetal_ranked_aggregate, rank_corr_matrix, go_nonN[i],top_genes ), mc.cores = 10  ))
  nonN_cons_matrix_1[ ,i] = fetal_nonN_cons_coexp 
  
  fetal_micro_cons_coexp = unlist(mclapply(1:length(go_micro), function(i) get_conserved_coexp(fetal_ranked_aggregate, rank_corr_matrix, go_micro[i],top_genes ), mc.cores = 10  ))
  micro_cons_matrix_1[ ,i] = fetal_micro_cons_coexp 
  
  print(i)

}


```


```{r}
#Order genes based on mean conservation value
nProg_cons_matrix = reshape2::melt(t(nProg_cons_matrix_1))
mean_order_df = nProg_cons_matrix %>% group_by(Var2) %>% summarize(median = median(value)) %>% arrange(desc(median))
nProg_cons_matrix$Var2  = factor(nProg_cons_matrix$Var2, levels = c(as.character(mean_order_df$Var2)))
ggplot(nProg_cons_matrix, aes(x = Var2, y = value)) + geom_boxplot(outlier.size = .5, outlier.alpha = .5) +
  ylab('Conservation of Fetal coexpression (AUC)') + xlab('Top 100 fetal Neural Progenitor markers') + ylim(0,1) +
  theme(axis.text.x = element_blank()) 

intProg_cons_matrix = reshape2::melt(t(intProg_cons_matrix_1))
mean_order_df = intProg_cons_matrix %>% group_by(Var2) %>% summarize(median = median(value)) %>% arrange(desc(median))
intProg_cons_matrix$Var2  = factor(intProg_cons_matrix$Var2, levels = c(as.character(mean_order_df$Var2)))
ggplot(intProg_cons_matrix, aes(x = Var2, y = value)) + geom_boxplot(outlier.size = .5, outlier.alpha = .5) +
  ylab('Conservation of Fetal coexpression (AUC)') + xlab('Top 100 fetal Intermediate Progenitor markers') + ylim(0,1) +
  theme(axis.text.x = element_blank()) 

divProg_cons_matrix = reshape2::melt(t(divProg_cons_matrix_1))
mean_order_df = divProg_cons_matrix %>% group_by(Var2) %>% summarize(median = median(value)) %>% arrange(desc(median))
divProg_cons_matrix$Var2  = factor(divProg_cons_matrix$Var2, levels = c(as.character(mean_order_df$Var2)))
ggplot(divProg_cons_matrix, aes(x = Var2, y = value)) + geom_boxplot(outlier.size = .5, outlier.alpha = .5) +
  ylab('Conservation of Fetal coexpression (AUC)') + xlab('Top 100 fetal Dividing Progenitor markers') + ylim(0,1) +
  theme(axis.text.x = element_blank()) 

glut_cons_matrix = reshape2::melt(t(glut_cons_matrix_1))
mean_order_df = glut_cons_matrix %>% group_by(Var2) %>% summarize(median = median(value)) %>% arrange(desc(median))
glut_cons_matrix$Var2  = factor(glut_cons_matrix$Var2, levels = c(as.character(mean_order_df$Var2)))
ggplot(glut_cons_matrix, aes(x = Var2, y = value)) + geom_boxplot(outlier.size = .5, outlier.alpha = .5) +
  ylab('Conservation of Fetal coexpression (AUC)') + xlab('Top 100 fetal Glutamatergic markers') + ylim(0,1) +
  theme(axis.text.x = element_blank()) 

gaba_cons_matrix = reshape2::melt(t(gaba_cons_matrix_1))
mean_order_df = gaba_cons_matrix %>% group_by(Var2) %>% summarize(median = median(value)) %>% arrange(desc(median))
gaba_cons_matrix$Var2  = factor(gaba_cons_matrix$Var2, levels = c(as.character(mean_order_df$Var2)))
ggplot(gaba_cons_matrix, aes(x = Var2, y = value)) + geom_boxplot(outlier.size = .5, outlier.alpha = .5) +
  ylab('Conservation of Fetal coexpression (AUC)') + xlab('Top 100 fetal GABAergic markers') + ylim(0,1) +
  theme(axis.text.x = element_blank()) 

nonN_cons_matrix = reshape2::melt(t(nonN_cons_matrix_1))
mean_order_df = nonN_cons_matrix %>% group_by(Var2) %>% summarize(median = median(value)) %>% arrange(desc(median))
nonN_cons_matrix$Var2  = factor(nonN_cons_matrix$Var2, levels = c(as.character(mean_order_df$Var2)))
ggplot(nonN_cons_matrix, aes(x = Var2, y = value)) + geom_boxplot(outlier.size = .5, outlier.alpha = .5) +
  ylab('Conservation of Fetal coexpression (AUC)') + xlab('Top 100 fetal Non-neuronal markers') + ylim(0,1) +
  theme(axis.text.x = element_blank()) 

micro_cons_matrix = reshape2::melt(t(micro_cons_matrix_1))
mean_order_df = micro_cons_matrix %>% group_by(Var2) %>% summarize(median = median(value)) %>% arrange(desc(median))
micro_cons_matrix$Var2  = factor(micro_cons_matrix$Var2, levels = c(as.character(mean_order_df$Var2)))
ggplot(micro_cons_matrix, aes(x = Var2, y = value)) + geom_boxplot(outlier.size = .5, outlier.alpha = .5) +
  ylab('Conservation of Fetal coexpression (AUC)') + xlab('Top 100 fetal Microglia/Immune markers') + ylim(0,1) +
  theme(axis.text.x = element_blank()) 

```
```{r}
nonN_temp = nonN_cons_matrix %>% group_by(Var1) %>% summarize(sum = sum(value)/length(go_nonN)) %>% 
  mutate(gene_celltype_label = rep('fetal non-neuronal marker', 14))

micro_temp = micro_cons_matrix %>% group_by(Var1) %>% summarize(sum = sum(value)/length(go_micro)) %>% 
  mutate(gene_celltype_label = rep('fetal Microglia/Immune marker', 14))

gaba_temp = gaba_cons_matrix %>% group_by(Var1) %>% summarize(sum = sum(value)/length(go_gaba)) %>% 
  mutate(gene_celltype_label = rep('fetal GABAergic marker',14))

glut_temp = glut_cons_matrix %>% group_by(Var1) %>% summarize(sum = sum(value)/length(go_glut)) %>% 
  mutate(gene_celltype_label = rep('fetal glutamatergic marker', 14))

nProg_temp = nProg_cons_matrix %>% group_by(Var1) %>% summarize(sum = sum(value)/length(go_nProg)) %>% 
  mutate(gene_celltype_label = rep('fetal Neural progenitor marker',14))

intProg_temp = intProg_cons_matrix %>% group_by(Var1) %>% summarize(sum = sum(value)/length(go_intProg)) %>% 
  mutate(gene_celltype_label = rep('fetal Intermediate progenitor marker', 14))

divProg_temp = divProg_cons_matrix %>% group_by(Var1) %>% summarize(sum = sum(value)/length(go_divProg)) %>% 
  mutate(gene_celltype_label = rep('fetal Dividing progenitor marker', 14))

unannot_sum_cons_coexp_df = do.call(rbind, list(nonN_temp, micro_temp, gaba_temp, glut_temp, nProg_temp, intProg_temp, divProg_temp))

ggplot(unannot_sum_cons_coexp_df, aes(x = gene_celltype_label, y = sum)) + geom_boxplot() + 
  ylab('Strength of conserved celltype coexp (sum(AUC))') + 
  ggtitle('Unannotated fetal datasets') +
  scale_x_discrete(guide = guide_axis(n.dodge=2))

unannot_sum_cons_coexp_df
```


Now do with the annotated fetal datasets
Leave one out, redo metamarkers, get top metamarkers, reaggregate the individual networks, test conservation of coexpression for the left out network

```{r}
ranked_go_coexp_paths =list( 
                 fan_fu = '../data/coexpression/fetal/8_17_23_GO_coexp/ranked_fan_fu_coexp.Rdata',
                 poliodakis_1 = '../data/coexpression/fetal/8_17_23_GO_coexp/ranked_poliodakis_batch1_coexp.Rdata',
                 poliodakis_2 = '../data/coexpression/fetal/8_17_23_GO_coexp/ranked_poliodakis_batch2_coexp.Rdata',
                 GW14 = '../data/coexpression/fetal/8_17_23_GO_coexp/ranked_arealization_GW14_coexp.Rdata',
                 GW18_2 = '../data/coexpression/fetal/8_17_23_GO_coexp/ranked_arealization_GW18_2_coexp.Rdata',
                 GW19 = '../data/coexpression/fetal/8_17_23_GO_coexp/ranked_arealization_GW19_coexp.Rdata',
                 GW19_2 = '../data/coexpression/fetal/8_17_23_GO_coexp/ranked_arealization_GW19_2_coexp.Rdata',
                 GW20 = '../data/coexpression/fetal/8_17_23_GO_coexp/ranked_arealization_GW20_coexp.Rdata',
                 GW20_31 = '../data/coexpression/fetal/8_17_23_GO_coexp/ranked_arealization_GW20_31_coexp.Rdata',
                 GW20_34 = '../data/coexpression/fetal/8_17_23_GO_coexp/ranked_arealization_GW20_34_coexp.Rdata',
                 GW25 = '../data/coexpression/fetal/8_17_23_GO_coexp/ranked_arealization_GW25_coexp.Rdata',
                 linnarsson_GW5 = '../data/coexpression/fetal/8_17_23_GO_coexp/ranked_linnarsson_GW5_coexp.Rdata',
                 linnarsson_GW5_point_5 = '../data/coexpression/fetal/8_17_23_GO_coexp/ranked_linnarsson_GW5-5_coexp.Rdata',
                 linnarsson_GW6_1 ='../data/coexpression/fetal/8_17_23_GO_coexp/ranked_linnarsson_GW6_1_coexp.Rdata',
                 linnarsson_GW6_2 = '../data/coexpression/fetal/8_17_23_GO_coexp/ranked_linnarsson_GW6_2_coexp.Rdata',
                 linnarsson_GW6_point_6_1 = '../data/coexpression/fetal/8_17_23_GO_coexp/ranked_linnarsson_GW6-6_1_coexp.Rdata',
                 linnarsson_GW6_point_6_2 ='../data/coexpression/fetal/8_17_23_GO_coexp/ranked_linnarsson_GW6-6_2_coexp.Rdata',
                 linnarsson_GW6_point_7 ='../data/coexpression/fetal/8_17_23_GO_coexp/ranked_linnarsson_GW6-7_coexp.Rdata',
                 linnarsson_GW6_point_9_1 = '../data/coexpression/fetal/8_17_23_GO_coexp/ranked_linnarsson_GW6-9_1_coexp.Rdata',
                 linnarsson_GW6_point_9_2 ='../data/coexpression/fetal/8_17_23_GO_coexp/ranked_linnarsson_GW6-9_2_coexp.Rdata',
                 linnarsson_GW6_point_9_3 ='../data/coexpression/fetal/8_17_23_GO_coexp/ranked_linnarsson_GW6-9_3_coexp.Rdata',
                 linnarsson_GW6_point_9_4 = '../data/coexpression/fetal/8_17_23_GO_coexp/ranked_linnarsson_GW6-9_4_coexp.Rdata',
                 linnarsson_GW7 = '../data/coexpression/fetal/8_17_23_GO_coexp/ranked_linnarsson_GW7_coexp.Rdata',
                 linnarsson_GW7_point_5 = '../data/coexpression/fetal/8_17_23_GO_coexp/ranked_linnarsson_GW7-5_coexp.Rdata',
                 linnarsson_GW8_1 = '../data/coexpression/fetal/8_17_23_GO_coexp/ranked_linnarsson_GW8_1_coexp.Rdata',
                 linnarsson_GW8_2 = '../data/coexpression/fetal/8_17_23_GO_coexp/ranked_linnarsson_GW8_2_coexp.Rdata',
                 linnarsson_GW8_3 = '../data/coexpression/fetal/8_17_23_GO_coexp/ranked_linnarsson_GW8_3_coexp.Rdata',
                 linnarsson_GW8_point_1 =  '../data/coexpression/fetal/8_17_23_GO_coexp/ranked_linnarsson_GW8-1_coexp.Rdata',
                 linnarsson_GW8_point_5_1 ='../data/coexpression/fetal/8_17_23_GO_coexp/ranked_linnarsson_GW8-5_1_coexp.Rdata',
                 linnarsson_GW8_point_5_2 = '../data/coexpression/fetal/8_17_23_GO_coexp/ranked_linnarsson_GW8-5_2_coexp.Rdata',
                 linnarsson_GW9_point_2 =  '../data/coexpression/fetal/8_17_23_GO_coexp/ranked_linnarsson_GW9-2_coexp.Rdata',
                 linnarsson_GW9_point_5 ='../data/coexpression/fetal/8_17_23_GO_coexp/ranked_linnarsson_GW9-5_coexp.Rdata',
                 linnarsson_GW10 = '../data/coexpression/fetal/8_17_23_GO_coexp/ranked_linnarsson_GW10_coexp.Rdata',
                 linnarsson_GW11_point_5 ='../data/coexpression/fetal/8_17_23_GO_coexp/ranked_linnarsson_GW11-5_coexp.Rdata',
                 linnarsson_GW12 ='../data/coexpression/fetal/8_17_23_GO_coexp/ranked_linnarsson_GW12_coexp.Rdata',
                 linnarsson_GW13 ='../data/coexpression/fetal/8_17_23_GO_coexp/ranked_linnarsson_GW13_coexp.Rdata',
                 linnarsson_GW14 ='../data/coexpression/fetal/8_17_23_GO_coexp/ranked_linnarsson_GW14_coexp.Rdata')

annot_marker_names = c('fan_fu','poliodakis_1','poliodakis_2','GW14','GW18_2','GW19','GW19_2','GW20','GW20_31','GW20_34','GW25',
                        'linnarsson_GW5', 'linnarsson_GW5_point_5', 'linnarsson_GW6_1', 'linnarsson_GW6_2', 'linnarsson_GW6_point_6_1', 'linnarsson_GW6_point_6_2', 'linnarsson_GW6_point_7',
                  'linnarsson_GW6_point_9_1', 'linnarsson_GW6_point_9_2', 'linnarsson_GW6_point_9_3', 'linnarsson_GW6_point_9_4', 'linnarsson_GW7', 'linnarsson_GW7_point_5', 
                  'linnarsson_GW8_1', 'linnarsson_GW8_2','linnarsson_GW8_3', 'linnarsson_GW8_point_1', 'linnarsson_GW8_point_5_1', 'linnarsson_GW8_point_5_2', 'linnarsson_GW9_point_2',
                  'linnarsson_GW9_point_5','linnarsson_GW10', 'linnarsson_GW11_point_5', 'linnarsson_GW12', 'linnarsson_GW13', 'linnarsson_GW14')


```


```{r}
#Dataframe to hold all the results
#The marker sets are going to be slightly different lengths because I'm redoing the metamarkers each time
cross_valid_fetal_cons_coexp_df = data.frame(Var1 = c(), Var2 = c(), value = c(), gene_celltype_label = c(), num_go_markers = c())

for(j in 1:length(annot_marker_names)){

  start = Sys.time()
  current_dataset = annot_marker_names[j]
  
  #Redo the metamarkers and get the top 100 cell type markers
  fetal_meta_markers = make_meta_markers(test_markers[names(test_markers) != current_dataset], detailed_stats = TRUE)
  
  gaba_markers = get_celltype_ranked_markers(fetal_meta_markers, 'GABAergic',100, 'rank')
  glut_markers = get_celltype_ranked_markers(fetal_meta_markers, 'Glutamatergic',100, 'rank')
  nProg_markers = get_celltype_ranked_markers(fetal_meta_markers, 'Neural_Progenitor',100, 'rank')
  nonN_markers = get_celltype_ranked_markers(fetal_meta_markers, 'Non-neuronal',100, 'rank')
  intProg_markers = get_celltype_ranked_markers(fetal_meta_markers, 'Intermediate_Progenitor',100, 'rank')
  divProg_markers = get_celltype_ranked_markers(fetal_meta_markers, 'Dividing_Progenitor',100, 'rank')
  micro_markers = get_celltype_ranked_markers(fetal_meta_markers, 'Microglia/Immune',100, 'rank')
  
  #Ignore duplicates
  all_markers = c( intProg_markers$gene, gaba_markers$gene, glut_markers$gene, nProg_markers$gene, divProg_markers$gene, nonN_markers$gene, micro_markers$gene )
  dups = all_markers[duplicated(all_markers)]
  
  intProg_markers = intProg_markers %>% filter(!gene %in% dups)
  nProg_markers = nProg_markers %>% filter(!gene %in% dups)
  divProg_markers = divProg_markers %>% filter(!gene %in% dups)
  nonN_markers = nonN_markers %>% filter(!gene %in% dups)
  gaba_markers = gaba_markers %>% filter(!gene %in% dups)
  glut_markers = glut_markers %>% filter(!gene %in% dups)
  micro_markers = micro_markers %>% filter(!gene %in% dups)
  
  go_nProg = nProg_markers$gene[nProg_markers$gene %in% go_genes]
  go_nonN = nonN_markers$gene[nonN_markers$gene %in% go_genes]
  go_gaba = gaba_markers$gene[gaba_markers$gene %in% go_genes]
  go_glut = glut_markers$gene[glut_markers$gene %in% go_genes]
  go_intProg = intProg_markers$gene[intProg_markers$gene %in% go_genes]
  go_divProg = divProg_markers$gene[divProg_markers$gene %in% go_genes]
  go_micro = micro_markers$gene[micro_markers$gene %in% go_genes]

  #Leave one out and get the new aggregated fetal network
  sum_rank_corr_genes = matrix(rep(0, length(go_genes)^2), nrow = length(go_genes), ncol = length(go_genes), dimnames = list(go_genes, go_genes))
  current_ranked_go_coexp_paths = ranked_go_coexp_paths[names(ranked_go_coexp_paths) != current_dataset]
  
  for(i in 1:length(current_ranked_go_coexp_paths)){
    x = load(current_ranked_go_coexp_paths[[i]])
    rank_mat = get(x)
    rm(x)
    #Reorder to match the same gene order
    index = match(go_genes, rownames(rank_mat))
    rank_mat = rank_mat[index, index]
    sum_rank_corr_genes = sum_rank_corr_genes + rank_mat
  }
  #Divide sums by the number of datasets to get averages
  agg_rank_mat = sum_rank_corr_genes / length(current_ranked_go_coexp_paths)
  
  #Rank
  rank_corr = frank(c(agg_rank_mat), ties.method = 'min') - 1
  #Divide by max rank squeezes everything between 0 and 1
  norm_rank_corr = rank_corr / max(rank_corr)
  #Put back into a matrix
  rank_corr_matrix = matrix(norm_rank_corr, nrow = nrow(agg_rank_mat), ncol = ncol(agg_rank_mat), 
                            dimnames = list(rownames(agg_rank_mat), colnames(agg_rank_mat)))
  agg_rank_mat = rank_corr_matrix
  diag(agg_rank_mat) = 1
  print('Done aggregating networks...')


  #Load in the coexpression network of the leftout dataset and get the conservation of coexpression
  top_genes = 10
  load(ranked_go_coexp_paths[names(ranked_go_coexp_paths) == current_dataset][[1]])
  #Conservation of coexpression
  fetal_nProg_cons_coexp = unlist(mclapply(1:length(go_nProg), function(i) get_conserved_coexp(agg_rank_mat, rank_corr_matrix, go_nProg[i],top_genes ), mc.cores = 10 ))
  fetal_intProg_cons_coexp = unlist(mclapply(1:length(go_intProg), function(i) get_conserved_coexp(agg_rank_mat, rank_corr_matrix, go_intProg[i],top_genes ), mc.cores = 10  ))
  fetal_divProg_cons_coexp = unlist(mclapply(1:length(go_divProg), function(i) get_conserved_coexp(agg_rank_mat, rank_corr_matrix, go_divProg[i],top_genes ), mc.cores = 10  ))
  fetal_glut_cons_coexp = unlist(mclapply(1:length(go_glut), function(i) get_conserved_coexp(agg_rank_mat, rank_corr_matrix, go_glut[i],top_genes ), mc.cores = 10  ))
  fetal_gaba_cons_coexp = unlist(mclapply(1:length(go_gaba), function(i) get_conserved_coexp(agg_rank_mat, rank_corr_matrix, go_gaba[i],top_genes ), mc.cores = 10  ))
  fetal_nonN_cons_coexp = unlist(mclapply(1:length(go_nonN), function(i) get_conserved_coexp(agg_rank_mat, rank_corr_matrix, go_nonN[i],top_genes ), mc.cores = 10  ))
  fetal_micro_cons_coexp = unlist(mclapply(1:length(go_micro), function(i) get_conserved_coexp(agg_rank_mat, rank_corr_matrix, go_micro[i],top_genes ), mc.cores = 10  ))
  
  #Vectors for dataframe
  genes = c(go_nProg, go_intProg, go_divProg, go_glut, go_gaba, go_nonN, go_micro)
  gene_celltype_labels = c(rep('fetal Neural progenitor marker', length = length(go_nProg)),
                           rep('fetal Intermediate progenitor marker', length = length(go_intProg)),
                           rep('fetal Dividing progenitor marker', length = length(go_divProg)),
                           rep('fetal glutamatergic marker', length = length(go_glut)),
                           rep('fetal GABAergic marker', length = length(go_gaba)),
                           rep('fetal non-neuronal marker', length = length(go_nonN)),
                           rep('fetal Microglia/Immune marker', length = length(go_micro)))
  
  num_markers = c(rep(length(go_nProg), length = length(go_nProg)),
                           rep(length(go_intProg), length = length(go_intProg)),
                           rep(length(go_divProg), length = length(go_divProg)),
                           rep(length(go_glut), length = length(go_glut)),
                           rep(length(go_gaba), length = length(go_gaba)),
                           rep(length(go_nonN), length = length(go_nonN)),
                           rep(length(go_micro), length = length(go_micro))) 
  
  conserved_coexp_vec = c(fetal_nProg_cons_coexp ,fetal_intProg_cons_coexp,fetal_divProg_cons_coexp,
                          fetal_glut_cons_coexp , fetal_gaba_cons_coexp , fetal_nonN_cons_coexp, fetal_micro_cons_coexp )
  dataset_label = rep(current_dataset, length = length(genes))
  
  
  #save results
  temp_df = data.frame(Var1 = dataset_label, Var2 = genes, value = conserved_coexp_vec, gene_celltype_label = gene_celltype_labels, num_go_markers = num_markers)
  
  cross_valid_fetal_cons_coexp_df = rbind(cross_valid_fetal_cons_coexp_df, temp_df)
  end = Sys.time()
  print(end - start)
  print(sprintf('Done with %i datasets...', j))

}
```



```{r}
cross_valid_fetal_cons_coexp_df

```



save fetal results
```{r}

save(cross_valid_fetal_cons_coexp_df, unannot_sum_cons_coexp_df, file = '../data/coexpression/fetal_preserve_fetal_coexp_dfs.Rdata' )
```

```{r}

load(file = '../data/coexpression/fetal_preserve_fetal_coexp_dfs.Rdata' )

```





sum for each dataset

```{r}
#Normalizing the sum by the number of markers, can compare across the datasets

sum_cons_coexp_df = cross_valid_fetal_cons_coexp_df %>% group_by(Var1, gene_celltype_label) %>% summarize(sum = sum(value)/num_go_markers[1])

ggplot(sum_cons_coexp_df, aes(x = gene_celltype_label, y = sum)) + geom_boxplot(scale = 'width') + ylab('Strength of conserved celltype coexp (sum(AUC))') +
  scale_x_discrete(guide = guide_axis(n.dodge=2))

```



load in the organoid results and compare
```{r}

all_sample_meta = read.csv('../data/seurat_objs/all_data_just_meta_seurat_with_addl_features_v2.csv')
all_sample_meta
```

```{r}

sum_vec = c(all_sample_meta$nProg_cons_coexp_metric, all_sample_meta$intProg_cons_coexp_metric, all_sample_meta$dividing_cons_coexp_metric,
            all_sample_meta$glut_cons_coexp_metric, all_sample_meta$gaba_cons_coexp_metric, all_sample_meta$nonN_cons_coexp_metric,
            all_sample_meta$micro_cons_coexp_metric)

label_vec = c(rep('fetal Neural progenitor marker', nrow(all_sample_meta)), rep('fetal Intermediate progenitor marker', nrow(all_sample_meta)),
              rep('fetal Dividing progenitor marker', nrow(all_sample_meta)),
              rep('fetal glutamatergic marker', nrow(all_sample_meta)), 
              rep('fetal GABAergic marker', nrow(all_sample_meta)), rep('fetal non-neuronal marker', nrow(all_sample_meta)),
              rep('fetal Microglia/Immune marker', nrow(all_sample_meta)))

var1_vec = rep(paste('organoid dataset', as.character(1:nrow(all_sample_meta))), 7)
dataset_label = rep('organoid', length = length(var1_vec))


organoid_sum_cons_coexp_df = data.frame(Var1 = var1_vec, sum = sum_vec, gene_celltype_label = label_vec, dataset_type = dataset_label)

organoid_sum_cons_coexp_df
```




```{r}

unannot_sum_cons = unannot_sum_cons_coexp_df %>% mutate(dataset_type = rep('unannotated fetal', length = nrow(unannot_sum_cons_coexp_df)))
unannot_sum_cons
#Being weird, mutate wasn't working for some reason
sum_cons = sum_cons_coexp_df
sum_cons$dataset_type = rep('annotated fetal', length = nrow(sum_cons_coexp_df))
sum_cons
```


```{r}

comb_cons_coexp_df = rbind(sum_cons, unannot_sum_cons, organoid_sum_cons_coexp_df)
comb_cons_coexp_df$dataset_type = factor(comb_cons_coexp_df$dataset_type, levels = c('annotated fetal','unannotated fetal','organoid'))
comb_cons_coexp_df$gene_celltype_label = factor(comb_cons_coexp_df$gene_celltype_label, levels = c('fetal Dividing progenitor marker', 'fetal Neural progenitor marker',
                                                                                                   'fetal glutamatergic marker','fetal non-neuronal marker',
                                                                                                   'fetal Intermediate progenitor marker',
                                                                                                   'fetal GABAergic marker','fetal Microglia/Immune marker'))

color_vec = c('annotated fetal' = 'darkorange', 'unannotated fetal' = 'firebrick', 'organoid'= 'darkorchid')

ggplot(filter(comb_cons_coexp_df, gene_celltype_label != 'fetal Microglia/Immune marker'), aes(x = gene_celltype_label, y = sum, fill = dataset_type)) + geom_boxplot(alpha = .75) +
  geom_hline(yintercept = .5, col = 'red', linetype = 'dashed') +
  scale_x_discrete(guide = guide_axis(n.dodge=2)) + ylab('Strength of conserved coexpression') +
  scale_fill_manual(values = color_vec) +
  theme(axis.title.x = element_blank())
ggsave(filename = 'combined_strength_cons_coexpression_v5.pdf', path = 'graphs/coexpression/conserved_coexpression/', 
       device = 'pdf', useDingbats = F, height = 4, width = 8.5)

ggplot(comb_cons_coexp_df, aes(x = gene_celltype_label, y = sum, fill = dataset_type)) + geom_boxplot(alpha = .75) +
  geom_hline(yintercept = .5, col = 'red', linetype = 'dashed') +
  scale_x_discrete(guide = guide_axis(n.dodge=2)) + ylab('Strength of conserved coexpression') +
  scale_fill_manual(values = color_vec) +
  theme(axis.title.x = element_blank())
ggsave(filename = 'combined_strength_cons_coexpression_with_micro_v5.pdf', path = 'graphs/coexpression/conserved_coexpression/', 
       device = 'pdf', useDingbats = F, height = 4, width = 8.5)

```


```{r}

save(comb_cons_coexp_df, file = '../data/coexpression/combined_consCoexp_celltype_markers_v5.Rdata')

```

```{r}

load( file = '../data/coexpression/combined_consCoexp_celltype_markers_v5.Rdata')
comb_cons_coexp_df
```
get the mean across the cell types for the fetal datasets, then mean across the datasets

```{r}

comb_cons_coexp_df %>% group_by(Var1) %>% summarize(mean_celltype = mean(sum), dataset_type = dataset_type[1]) %>% 
  group_by(dataset_type) %>% 
  summarize(mean_type = mean(mean_celltype), sd_type = sd(mean_celltype))
```

get the scores for the vertical/orbital shaker organoid study

datasets 128 - 130 are orbital
datasets 131-133 are vertical
```{r}
all_sample_meta
comb_cons_coexp_df %>% filter(Var1 %in% c('organoid dataset 128','organoid dataset 129','organoid dataset 130')) %>% group_by(gene_celltype_label) %>% summarize(mean = mean(sum), sd =  sd(sum))
comb_cons_coexp_df %>% filter(Var1 %in% c('organoid dataset 131','organoid dataset 132','organoid dataset 133')) %>% group_by(gene_celltype_label) %>% summarize(mean = mean(sum), sd =  sd(sum))

```




Find the organoid datasets that have high conservation of coexpression

Trying to put a number to the organoid datasets that produce siimilar coexpression to fetal data. Most of the organoid protocols are not geared toward producing no neuronal or gabaergic cells, so a bit unfair to include those categories.
Also the worst performing fetal dataset is zhou_lu which is a different normalization from the others, so makes me uncomfortable to use that as the fetal reference.
An cut off of .9 seems decent, slightly more range than that of the fetal (lowest fetal excluding zhou_lu seems like 9.1) but also a really high score (roughly 90% of the markers had conserved coexp AUROCs close to 1)



2 organoid datasets have an average greater than .90

```{r}

comb_cons_coexp_df %>% filter(!gene_celltype_label %in% c('fetal non-neuronal marker','fetal GABAergic marker')) %>% group_by(Var1) %>% 
  summarize(avg_cons_coexp = mean(sum)) %>% arrange(desc(avg_cons_coexp))


organoid_sum_cons_coexp_df %>% filter(!gene_celltype_label %in% c('fetal non-neuronal marker','fetal GABAergic marker')) %>% group_by(Var1) %>% 
  summarize(avg_cons_coexp = mean(sum)) %>% arrange(desc(avg_cons_coexp))


```




Check out what protocols they come from

```{r}

temp_org = organoid_sum_cons_coexp_df %>% filter(!gene_celltype_label %in% c('fetal non-neuronal marker','fetal GABAergic marker')) %>% group_by(Var1) %>% 
  summarize(avg_cons_coexp = mean(sum)) %>% arrange(desc(avg_cons_coexp))


dataset_nums = filter(temp_org, avg_cons_coexp > .9)$Var1
dataset_nums = as.numeric(sapply(strsplit(dataset_nums, ' '),'[[',3))


table(all_sample_meta[dataset_nums, 'Author.described.protocol' ])


```



######################################
Violin plots placing an example organoid dataset in reference to the fetal datasets
And then the organoid datasets
Put the percentiles in text above the violin plots
For showcasing the future use of this analysis for benchmarking new data

#####################################


```{r}
#compute percentile from ecdf
ecdf_fun <- function(x,perc) ecdf(x)(perc)


example_org_df = filter(comb_cons_coexp_df, Var1 == 'organoid dataset 16' & gene_celltype_label != 'fetal Microglia/Immune marker')
temp_comb_cons_coexp_df = filter(comb_cons_coexp_df, Var1 != 'organoid dataset 16'& gene_celltype_label != 'fetal Microglia/Immune marker') 

fetal_label = vector(mode = 'character', length = nrow(temp_comb_cons_coexp_df))
fetal_label[grepl('fetal', temp_comb_cons_coexp_df$dataset_type)] = 'Fetal'
fetal_label[grepl('organoid', temp_comb_cons_coexp_df$dataset_type)] = 'Organoid'
temp_comb_cons_coexp_df$fetal_label = fetal_label

#Get percentiles of the example organoid dataset using the empirical CDFs, round to nearest percentile
org_fetal_percentile_vec = sapply(1:nrow(example_org_df), function(i) paste(as.character(round(ecdf_fun(filter(temp_comb_cons_coexp_df, 
                                                                              fetal_label == 'Fetal' & gene_celltype_label == example_org_df$gene_celltype_label[i])$sum,
                                                                       filter(example_org_df, gene_celltype_label == example_org_df$gene_celltype_label[i])$sum)*100)), 'percentile'))

org_org_percentile_vec = sapply(1:nrow(example_org_df), function(i) paste(as.character(round(ecdf_fun(filter(temp_comb_cons_coexp_df, 
                                                                              fetal_label == 'Organoid' & gene_celltype_label == example_org_df$gene_celltype_label[i])$sum,
                                                                       filter(example_org_df, gene_celltype_label == example_org_df$gene_celltype_label[i])$sum)*100)), 'percentile'))

example_org_df$org_fetal_percentile = org_fetal_percentile_vec 
example_org_df$org_org_percentile = org_org_percentile_vec 

ggplot(filter(temp_comb_cons_coexp_df, fetal_label == 'Fetal'), aes(x = gene_celltype_label, y = sum)) + geom_violin(scale = 'width') +
  ylim(0,1) + ylab('Preserved Co-expression score') + ggtitle('Fetal datasets') +
  stat_summary(data = example_org_df, aes(x = gene_celltype_label, y = sum), fun = 'sum', color = 'red', geom = 'crossbar', width = .5) +
  geom_text(data = example_org_df, aes(label = org_fetal_percentile), y = .25) +
  scale_x_discrete(guide = guide_axis(n.dodge=2))
ggsave(filename = 'example_org_fetal_percentiles_violins_v2.pdf', path = 'graphs/coexpression/conserved_coexpression/', useDingbats = F, device = 'pdf', height = 4, width = 7)
ggsave(filename = 'example_org_fetal_percentiles_violins_v2.png', path = 'graphs/coexpression/conserved_coexpression/', device = 'png', height = 4, width = 7)

ggplot(filter(temp_comb_cons_coexp_df, fetal_label == 'Organoid'), aes(x = gene_celltype_label, y = sum)) + geom_violin(scale = 'width') +
  ylim(0,1) + ylab('Preserved Co-expression score') + ggtitle('Organoid datasets') +
  stat_summary(data = example_org_df, aes(x = gene_celltype_label, y = sum), fun = 'sum', color = 'red', geom = 'crossbar', width = .5) +
  geom_text(data = example_org_df, aes(label = org_org_percentile), y = .25) +
  scale_x_discrete(guide = guide_axis(n.dodge=2))
ggsave(filename = 'example_org_org_percentiles_violins_v2.pdf', path = 'graphs/coexpression/conserved_coexpression/', useDingbats = F, device = 'pdf', height = 4, width = 7)
ggsave(filename = 'example_org_org_percentiles_violins_v2.png', path = 'graphs/coexpression/conserved_coexpression/', device = 'png', height = 4, width = 7)
```
save a version of the comb_df for use in the Rpackage


```{r}

meta_presCoexp_df = comb_cons_coexp_df
fetal_label = vector(mode = 'character', length = nrow(meta_presCoexp_df))
fetal_label[grepl('fetal', meta_presCoexp_df$dataset_type)] = 'Fetal'
fetal_label[grepl('organoid', meta_presCoexp_df$dataset_type)] = 'Organoid'
meta_presCoexp_df$dataset_type = fetal_label
meta_presCoexp_df


saveRDS(meta_presCoexp_df, file = 'r_package_data/meta_presCoexp_df_v2.rds')

```




##########################################
Boxplots of the cell-type consCoexpression scores across the organoid protocols
Order by cell type
##########################################

```{r}
all_sample_meta


```







```{r}
org_egad_marker_palette = c('Fetal Non-neuronal markers' = class_palette[20], 'Fetal GABAergic markers' = class_palette[1], 
                         'Fetal Glutamatergic markers' = class_palette[14], 
                         'Fetal Neural Progenitor markers' = class_palette[4],
                         'Fetal Intermediate Progenitor markers' = class_palette[16],
                         'Fetal Dividing Progenitor markers' = class_palette[10],
                         'Fetal Microglia/Immune markers' = class_palette[18])


#Order the boxplots per celltype
label_order_df = all_sample_meta %>% group_by(Author.described.protocol) %>% summarize(median = median(intProg_cons_coexp_metric)) %>% arrange(desc(median))
all_sample_meta$Author.described.protocol = factor(all_sample_meta$Author.described.protocol, levels = label_order_df$Author.described.protocol)
ggplot(all_sample_meta, aes(x = Author.described.protocol, y = intProg_cons_coexp_metric)) + geom_boxplot(fill = org_egad_marker_palette['Fetal Intermediate Progenitor markers'] ) +
  ylab('Organoid Preserved Co-expression scores') + xlab('Author described organoid protocol') + ggtitle('Intermediate progenitors') + ylim(.4, 1) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = .5))
ggsave(filename = 'organoid_intProg_consCoexpScores_by_region_v5.pdf', path = 'graphs/coexpression/conserved_coexpression',
       device = 'pdf', useDingbats = F,
       height = 5, width = 6)


label_order_df = all_sample_meta %>% group_by(Author.described.protocol) %>% summarize(median = median(nProg_cons_coexp_metric)) %>% arrange(desc(median))
all_sample_meta$Author.described.protocol = factor(all_sample_meta$Author.described.protocol, levels = label_order_df$Author.described.protocol)
ggplot(all_sample_meta, aes(x = Author.described.protocol, y = nProg_cons_coexp_metric)) + geom_boxplot(fill = org_egad_marker_palette['Fetal Neural Progenitor markers'] ) +
  ylab('Organoid Preserved Co-expression scores') + xlab('Author described organoid protocol') + ggtitle('Neural progenitors') + ylim(.4, 1) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = .5))
ggsave(filename = 'organoid_nProg_consCoexpScores_by_region_v5.pdf', path = 'graphs/coexpression/conserved_coexpression',
       device = 'pdf', useDingbats = F,
       height = 5, width = 6)


label_order_df = all_sample_meta %>% group_by(Author.described.protocol) %>% summarize(median = median(dividing_cons_coexp_metric)) %>% arrange(desc(median))
all_sample_meta$Author.described.protocol = factor(all_sample_meta$Author.described.protocol, levels = label_order_df$Author.described.protocol)
ggplot(all_sample_meta, aes(x = Author.described.protocol, y = dividing_cons_coexp_metric)) + geom_boxplot(fill = org_egad_marker_palette['Fetal Dividing Progenitor markers'] ) +
  ylab('Organoid Preserved Co-expression scores') + xlab('Author described organoid protocol') + ggtitle('Dividing progenitors') + ylim(.4, 1) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = .5))
ggsave(filename = 'organoid_divProg_consCoexpScores_by_region_v5.pdf', path = 'graphs/coexpression/conserved_coexpression',
       device = 'pdf', useDingbats = F,
       height = 5, width = 6)


label_order_df = all_sample_meta %>% group_by(Author.described.protocol) %>% summarize(median = median(glut_cons_coexp_metric)) %>% arrange(desc(median))
all_sample_meta$Author.described.protocol = factor(all_sample_meta$Author.described.protocol, levels = label_order_df$Author.described.protocol)
ggplot(all_sample_meta, aes(x = Author.described.protocol, y = glut_cons_coexp_metric)) + geom_boxplot(fill = org_egad_marker_palette['Fetal Glutamatergic markers'] ) +
  ylab('Organoid Preserved Co-expression scores') + xlab('Author described organoid protocol') + ggtitle('Glutamatergic') + ylim(.4, 1) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = .5))
ggsave(filename = 'organoid_glut_consCoexpScores_by_region_v5.pdf', path = 'graphs/coexpression/conserved_coexpression',
       device = 'pdf', useDingbats = F,
       height = 5, width = 6)


label_order_df = all_sample_meta %>% group_by(Author.described.protocol) %>% summarize(median = median(gaba_cons_coexp_metric)) %>% arrange(desc(median))
all_sample_meta$Author.described.protocol = factor(all_sample_meta$Author.described.protocol, levels = label_order_df$Author.described.protocol)
ggplot(all_sample_meta, aes(x = Author.described.protocol, y = gaba_cons_coexp_metric)) + geom_boxplot(fill = org_egad_marker_palette['Fetal GABAergic markers'] ) +
  ylab('Organoid Preserved Co-expression scores') + xlab('Author described organoid protocol') + ggtitle('GABAergic') + ylim(.4, 1) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = .5))
ggsave(filename = 'organoid_gaba_consCoexpScores_by_region_v5.pdf', path = 'graphs/coexpression/conserved_coexpression',
       device = 'pdf', useDingbats = F,
       height = 5, width = 6)


label_order_df = all_sample_meta %>% group_by(Author.described.protocol) %>% summarize(median = median(nonN_cons_coexp_metric)) %>% arrange(desc(median))
all_sample_meta$Author.described.protocol = factor(all_sample_meta$Author.described.protocol, levels = label_order_df$Author.described.protocol)
ggplot(all_sample_meta, aes(x = Author.described.protocol, y = nonN_cons_coexp_metric)) + geom_boxplot(fill = org_egad_marker_palette['Fetal Non-neuronal markers'] ) +
  ylab('Organoid Preserved Co-expression scores') + xlab('Author described organoid protocol') + ggtitle('Non-neuronal') + ylim(.4, 1) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = .5))
ggsave(filename = 'organoid_nonN_consCoexpScores_by_region_v5.pdf', path = 'graphs/coexpression/conserved_coexpression',
       device = 'pdf', useDingbats = F,
       height = 5, width = 6)


label_order_df = all_sample_meta %>% group_by(Author.described.protocol) %>% summarize(median = median(micro_cons_coexp_metric)) %>% arrange(desc(median))
all_sample_meta$Author.described.protocol = factor(all_sample_meta$Author.described.protocol, levels = label_order_df$Author.described.protocol)
ggplot(all_sample_meta, aes(x = Author.described.protocol, y = micro_cons_coexp_metric)) + geom_boxplot(fill = org_egad_marker_palette['Fetal Microglia/Immune markers'] ) +
  ylab('Organoid Preserved Co-expression scores') + xlab('Author described organoid protocol') + ggtitle('Microglia/Immune') + ylim(.4, 1) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = .5))
ggsave(filename = 'organoid_micro_consCoexpScores_by_region_v5.pdf', path = 'graphs/coexpression/conserved_coexpression',
       device = 'pdf', useDingbats = F,
       height = 5, width = 6)
```
And now with the grouped protocol classification

```{r}

#Order the boxplots per celltype
label_order_df = all_sample_meta %>% group_by(Protocol.classification.region) %>% summarize(median = median(intProg_cons_coexp_metric)) %>% arrange(desc(median))
all_sample_meta$Protocol.classification.region = factor(all_sample_meta$Protocol.classification.region, levels = label_order_df$Protocol.classification.region)
ggplot(all_sample_meta, aes(x = Protocol.classification.region, y = intProg_cons_coexp_metric)) + 
  geom_boxplot(fill = org_egad_marker_palette['Fetal Intermediate Progenitor markers'] ) +
  ylab('Organoid Preserved Co-expression scores') + xlab('Organoid protocol') + ggtitle('Intermediate progenitors') + ylim(.4, 1) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = .5))
ggsave(filename = 'organoid_intProg_consCoexpScores_by_grouped_proto_v5.pdf', path = 'graphs/coexpression/conserved_coexpression',
       device = 'pdf', useDingbats = F,
       height = 5, width = 6)


label_order_df = all_sample_meta %>% group_by(Protocol.classification.region) %>% summarize(median = median(nProg_cons_coexp_metric)) %>% arrange(desc(median))
all_sample_meta$Protocol.classification.region = factor(all_sample_meta$Protocol.classification.region, levels = label_order_df$Protocol.classification.region)
ggplot(all_sample_meta, aes(x = Protocol.classification.region, y = nProg_cons_coexp_metric)) + geom_boxplot(fill = org_egad_marker_palette['Fetal Neural Progenitor markers'] ) +
  ylab('Organoid Preserved Co-expression scores') + xlab('Organoid protocol') + ggtitle('Neural progenitors') + ylim(.4, 1) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = .5))
ggsave(filename = 'organoid_nProg_consCoexpScores_by_grouped_proto_v5.pdf', path = 'graphs/coexpression/conserved_coexpression',
       device = 'pdf', useDingbats = F,
       height = 5, width = 6)


label_order_df = all_sample_meta %>% group_by(Protocol.classification.region) %>% summarize(median = median(dividing_cons_coexp_metric)) %>% arrange(desc(median))
all_sample_meta$Protocol.classification.region = factor(all_sample_meta$Protocol.classification.region, levels = label_order_df$Protocol.classification.region)
ggplot(all_sample_meta, aes(x = Protocol.classification.region, y = dividing_cons_coexp_metric)) + geom_boxplot(fill = org_egad_marker_palette['Fetal Dividing Progenitor markers'] ) +
  ylab('Organoid Preserved Co-expression scores') + xlab('Organoid protocol') + ggtitle('Dividing progenitors') + ylim(.4, 1) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = .5))
ggsave(filename = 'organoid_divProg_consCoexpScores_by_grouped_proto_v5.pdf', path = 'graphs/coexpression/conserved_coexpression',
       device = 'pdf', useDingbats = F,
       height = 5, width = 6)


label_order_df = all_sample_meta %>% group_by(Protocol.classification.region) %>% summarize(median = median(glut_cons_coexp_metric)) %>% arrange(desc(median))
all_sample_meta$Protocol.classification.region = factor(all_sample_meta$Protocol.classification.region, levels = label_order_df$Protocol.classification.region)
ggplot(all_sample_meta, aes(x = Protocol.classification.region, y = glut_cons_coexp_metric)) + geom_boxplot(fill = org_egad_marker_palette['Fetal Glutamatergic markers'] ) +
  ylab('Organoid Preserved Co-expression scores') + xlab('Organoid protocol') + ggtitle('Glutamatergic') + ylim(.4, 1) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = .5))
ggsave(filename = 'organoid_glut_consCoexpScores_by_grouped_proto_v5.pdf', path = 'graphs/coexpression/conserved_coexpression',
       device = 'pdf', useDingbats = F,
       height = 5, width = 6)


label_order_df = all_sample_meta %>% group_by(Protocol.classification.region) %>% summarize(median = median(gaba_cons_coexp_metric)) %>% arrange(desc(median))
all_sample_meta$Protocol.classification.region = factor(all_sample_meta$Protocol.classification.region, levels = label_order_df$Protocol.classification.region)
ggplot(all_sample_meta, aes(x = Protocol.classification.region, y = gaba_cons_coexp_metric)) + geom_boxplot(fill = org_egad_marker_palette['Fetal GABAergic markers'] ) +
  ylab('Organoid Preserved Co-expression scores') + xlab('Organoid protocol') + ggtitle('GABAergic') + ylim(.4, 1) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = .5))
ggsave(filename = 'organoid_gaba_consCoexpScores_by_grouped_proto_v5.pdf', path = 'graphs/coexpression/conserved_coexpression',
       device = 'pdf', useDingbats = F,
       height = 5, width = 6)


label_order_df = all_sample_meta %>% group_by(Protocol.classification.region) %>% summarize(median = median(nonN_cons_coexp_metric)) %>% arrange(desc(median))
all_sample_meta$Protocol.classification.region = factor(all_sample_meta$Protocol.classification.region, levels = label_order_df$Protocol.classification.region)
ggplot(all_sample_meta, aes(x = Protocol.classification.region, y = nonN_cons_coexp_metric)) + geom_boxplot(fill = org_egad_marker_palette['Fetal Non-neuronal markers'] ) +
  ylab('Organoid Preserved Co-expression scores') + xlab('Organoid protocol') + ggtitle('Non-neuronal') + ylim(.4, 1) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = .5))
ggsave(filename = 'organoid_nonN_consCoexpScores_by_grouped_proto_v5.pdf', path = 'graphs/coexpression/conserved_coexpression',
       device = 'pdf', useDingbats = F,
       height = 5, width = 6)

label_order_df = all_sample_meta %>% group_by(Protocol.classification.region) %>% summarize(median = median(micro_cons_coexp_metric)) %>% arrange(desc(median))
all_sample_meta$Protocol.classification.region = factor(all_sample_meta$Protocol.classification.region, levels = label_order_df$Protocol.classification.region)
ggplot(all_sample_meta, aes(x = Protocol.classification.region, y = micro_cons_coexp_metric)) + geom_boxplot(fill = org_egad_marker_palette['Fetal Microglia/Immune markers'] ) +
  ylab('Organoid Preserved Co-expression scores') + xlab('Organoid protocol') + ggtitle('Microglia/Immune') + ylim(.4, 1) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = .5))
ggsave(filename = 'organoid_micro_consCoexpScores_by_grouped_proto_v5.pdf', path = 'graphs/coexpression/conserved_coexpression',
       device = 'pdf', useDingbats = F,
       height = 5, width = 6)
```







##########################################

For all genes, get a matrix of fetal preserved coexpression and then organoid preserved co-expression
Looking for genes that are consistently preserved across all fetal datasets and organoids do rather poorly
Or genes that organoids consistently do well with

Fetal will be leave one out, get a new aggregate network, score preserved co-expression across all genes
organoids will use the annotated fetal network as reference
Just use the annotated fetal data for this


#######################################



Start with the organoids just because leave one out fetal aggregation will take awhile
Check out the variance in scores for genes across the organoid datasets, any with consistent performance that are not cell-cycle related?


```{r}
top_genes = 10
all_gene_org_pres_fetal_mat = matrix(nrow = length(go_genes), ncol = length(all_sample_meta$sample_ids), dimnames = list(go_genes, all_sample_meta$sample_ids))

for(j in 1:length(all_sample_meta$sample_ids)){
  load(sprintf('../data/coexpression/organoid/6_26_24_GO_coexp/ranked_%s_coexp.Rdata', all_sample_meta$sample_ids[j]))
  
  start = Sys.time()
  test_all_genes = unlist(mclapply(1:length(go_genes), function(i) get_conserved_coexp(fetal_ranked_aggregate, rank_corr_matrix, go_genes[i],top_genes ), mc.cores = 10 ))
  end = Sys.time()
  print(end - start)
  
  all_gene_org_pres_fetal_mat[ ,j] = test_all_genes

}
```



```{r}


all_gene_org_pres_fetal_mat[1:10, 1:10]


```

```{r}
gene_means = rowMeans(all_gene_org_pres_fetal_mat)
gene_sds =apply(all_gene_org_pres_fetal_mat, MARGIN = 1, FUN = sd)
num_dataset_gene_zero_exp = rowSums(all_gene_org_pres_fetal_mat == .5)
table(num_dataset_gene_zero_exp <= 5)

hist(num_dataset_gene_zero_exp, breaks = seq(0,175, 1))

exp_genes_index = which(num_dataset_gene_zero_exp <= 5)

plot(gene_means[exp_genes_index], gene_sds[exp_genes_index], cex = .5)
plot(gene_sds[exp_genes_index], gene_means[exp_genes_index], cex = .5)

```


```{r}


org_all_gene_presCoexp_df = data.frame(gene = go_genes, mean_org_preCoexp = gene_means, sd_org_preCoexp = gene_sds, num_dataset_gene_not_expressed = num_dataset_gene_zero_exp)

org_all_gene_presCoexp_df %>% filter(num_dataset_gene_not_expressed <= 10) %>% arrange(desc(mean_org_preCoexp))
org_all_gene_presCoexp_df %>% filter(num_dataset_gene_not_expressed <= 10) %>% arrange(mean_org_preCoexp)

```

And now do a leave one out approach with the annotated fetal datasets
Leave one out and aggregate the rest
Get preserved coexpression scores for all genes



```{r}
ranked_go_coexp_paths =list( 
                 fan_fu = '../data/coexpression/fetal/8_17_23_GO_coexp/ranked_fan_fu_coexp.Rdata',
                 poliodakis_1 = '../data/coexpression/fetal/8_17_23_GO_coexp/ranked_poliodakis_batch1_coexp.Rdata',
                 poliodakis_2 = '../data/coexpression/fetal/8_17_23_GO_coexp/ranked_poliodakis_batch2_coexp.Rdata',
                 GW14 = '../data/coexpression/fetal/8_17_23_GO_coexp/ranked_arealization_GW14_coexp.Rdata',
                 GW18_2 = '../data/coexpression/fetal/8_17_23_GO_coexp/ranked_arealization_GW18_2_coexp.Rdata',
                 GW19 = '../data/coexpression/fetal/8_17_23_GO_coexp/ranked_arealization_GW19_coexp.Rdata',
                 GW19_2 = '../data/coexpression/fetal/8_17_23_GO_coexp/ranked_arealization_GW19_2_coexp.Rdata',
                 GW20 = '../data/coexpression/fetal/8_17_23_GO_coexp/ranked_arealization_GW20_coexp.Rdata',
                 GW20_31 = '../data/coexpression/fetal/8_17_23_GO_coexp/ranked_arealization_GW20_31_coexp.Rdata',
                 GW20_34 = '../data/coexpression/fetal/8_17_23_GO_coexp/ranked_arealization_GW20_34_coexp.Rdata',
                 GW25 = '../data/coexpression/fetal/8_17_23_GO_coexp/ranked_arealization_GW25_coexp.Rdata',
                 linnarsson_GW5 = '../data/coexpression/fetal/8_17_23_GO_coexp/ranked_linnarsson_GW5_coexp.Rdata',
                 linnarsson_GW5_point_5 = '../data/coexpression/fetal/8_17_23_GO_coexp/ranked_linnarsson_GW5-5_coexp.Rdata',
                 linnarsson_GW6_1 ='../data/coexpression/fetal/8_17_23_GO_coexp/ranked_linnarsson_GW6_1_coexp.Rdata',
                 linnarsson_GW6_2 = '../data/coexpression/fetal/8_17_23_GO_coexp/ranked_linnarsson_GW6_2_coexp.Rdata',
                 linnarsson_GW6_point_6_1 = '../data/coexpression/fetal/8_17_23_GO_coexp/ranked_linnarsson_GW6-6_1_coexp.Rdata',
                 linnarsson_GW6_point_6_2 ='../data/coexpression/fetal/8_17_23_GO_coexp/ranked_linnarsson_GW6-6_2_coexp.Rdata',
                 linnarsson_GW6_point_7 ='../data/coexpression/fetal/8_17_23_GO_coexp/ranked_linnarsson_GW6-7_coexp.Rdata',
                 linnarsson_GW6_point_9_1 = '../data/coexpression/fetal/8_17_23_GO_coexp/ranked_linnarsson_GW6-9_1_coexp.Rdata',
                 linnarsson_GW6_point_9_2 ='../data/coexpression/fetal/8_17_23_GO_coexp/ranked_linnarsson_GW6-9_2_coexp.Rdata',
                 linnarsson_GW6_point_9_3 ='../data/coexpression/fetal/8_17_23_GO_coexp/ranked_linnarsson_GW6-9_3_coexp.Rdata',
                 linnarsson_GW6_point_9_4 = '../data/coexpression/fetal/8_17_23_GO_coexp/ranked_linnarsson_GW6-9_4_coexp.Rdata',
                 linnarsson_GW7 = '../data/coexpression/fetal/8_17_23_GO_coexp/ranked_linnarsson_GW7_coexp.Rdata',
                 linnarsson_GW7_point_5 = '../data/coexpression/fetal/8_17_23_GO_coexp/ranked_linnarsson_GW7-5_coexp.Rdata',
                 linnarsson_GW8_1 = '../data/coexpression/fetal/8_17_23_GO_coexp/ranked_linnarsson_GW8_1_coexp.Rdata',
                 linnarsson_GW8_2 = '../data/coexpression/fetal/8_17_23_GO_coexp/ranked_linnarsson_GW8_2_coexp.Rdata',
                 linnarsson_GW8_3 = '../data/coexpression/fetal/8_17_23_GO_coexp/ranked_linnarsson_GW8_3_coexp.Rdata',
                 linnarsson_GW8_point_1 =  '../data/coexpression/fetal/8_17_23_GO_coexp/ranked_linnarsson_GW8-1_coexp.Rdata',
                 linnarsson_GW8_point_5_1 ='../data/coexpression/fetal/8_17_23_GO_coexp/ranked_linnarsson_GW8-5_1_coexp.Rdata',
                 linnarsson_GW8_point_5_2 = '../data/coexpression/fetal/8_17_23_GO_coexp/ranked_linnarsson_GW8-5_2_coexp.Rdata',
                 linnarsson_GW9_point_2 =  '../data/coexpression/fetal/8_17_23_GO_coexp/ranked_linnarsson_GW9-2_coexp.Rdata',
                 linnarsson_GW9_point_5 ='../data/coexpression/fetal/8_17_23_GO_coexp/ranked_linnarsson_GW9-5_coexp.Rdata',
                 linnarsson_GW10 = '../data/coexpression/fetal/8_17_23_GO_coexp/ranked_linnarsson_GW10_coexp.Rdata',
                 linnarsson_GW11_point_5 ='../data/coexpression/fetal/8_17_23_GO_coexp/ranked_linnarsson_GW11-5_coexp.Rdata',
                 linnarsson_GW12 ='../data/coexpression/fetal/8_17_23_GO_coexp/ranked_linnarsson_GW12_coexp.Rdata',
                 linnarsson_GW13 ='../data/coexpression/fetal/8_17_23_GO_coexp/ranked_linnarsson_GW13_coexp.Rdata',
                 linnarsson_GW14 ='../data/coexpression/fetal/8_17_23_GO_coexp/ranked_linnarsson_GW14_coexp.Rdata')

annot_marker_names = c('fan_fu','poliodakis_1','poliodakis_2','GW14','GW18_2','GW19','GW19_2','GW20','GW20_31','GW20_34','GW25',
                        'linnarsson_GW5', 'linnarsson_GW5_point_5', 'linnarsson_GW6_1', 'linnarsson_GW6_2', 'linnarsson_GW6_point_6_1', 'linnarsson_GW6_point_6_2', 'linnarsson_GW6_point_7',
                  'linnarsson_GW6_point_9_1', 'linnarsson_GW6_point_9_2', 'linnarsson_GW6_point_9_3', 'linnarsson_GW6_point_9_4', 'linnarsson_GW7', 'linnarsson_GW7_point_5', 
                  'linnarsson_GW8_1', 'linnarsson_GW8_2','linnarsson_GW8_3', 'linnarsson_GW8_point_1', 'linnarsson_GW8_point_5_1', 'linnarsson_GW8_point_5_2', 'linnarsson_GW9_point_2',
                  'linnarsson_GW9_point_5','linnarsson_GW10', 'linnarsson_GW11_point_5', 'linnarsson_GW12', 'linnarsson_GW13', 'linnarsson_GW14')


```


```{r}


top_genes = 10
all_gene_fetal_pres_fetal_mat = matrix(nrow = length(go_genes), ncol = length(annot_marker_names), dimnames = list(go_genes, annot_marker_names))

for(j in 1:length(annot_marker_names)){

  start = Sys.time()
  current_dataset = annot_marker_names[j]
  #Leave one out and get the new aggregated fetal network
  sum_rank_corr_genes = matrix(rep(0, length(go_genes)^2), nrow = length(go_genes), ncol = length(go_genes), dimnames = list(go_genes, go_genes))
  current_ranked_go_coexp_paths = ranked_go_coexp_paths[names(ranked_go_coexp_paths) != current_dataset]
  
  for(i in 1:length(current_ranked_go_coexp_paths)){
    x = load(current_ranked_go_coexp_paths[[i]])
    rank_mat = get(x)
    rm(x)
    #Reorder to match the same gene order
    index = match(go_genes, rownames(rank_mat))
    rank_mat = rank_mat[index, index]
    sum_rank_corr_genes = sum_rank_corr_genes + rank_mat
  }
  #Divide sums by the number of datasets to get averages
  agg_rank_mat = sum_rank_corr_genes / length(current_ranked_go_coexp_paths)
  
  #Rank
  rank_corr = frank(c(agg_rank_mat), ties.method = 'min') - 1
  #Divide by max rank squeezes everything between 0 and 1
  norm_rank_corr = rank_corr / max(rank_corr)
  #Put back into a matrix
  rank_corr_matrix = matrix(norm_rank_corr, nrow = nrow(agg_rank_mat), ncol = ncol(agg_rank_mat), 
                            dimnames = list(rownames(agg_rank_mat), colnames(agg_rank_mat)))
  agg_rank_mat = rank_corr_matrix
  diag(agg_rank_mat) = 1
  print('Done aggregating networks...')
  
  #Get the preserved coexpression aurocs for all genes
  x = load(ranked_go_coexp_paths[[j]])
  rank_mat = get(x)
  rm(x)
  
  test_all_genes = unlist(mclapply(1:length(go_genes), function(k) get_conserved_coexp(agg_rank_mat, rank_mat, go_genes[k],top_genes ), mc.cores = 10 ))
  all_gene_fetal_pres_fetal_mat[ ,j] = test_all_genes
  
  end = Sys.time()
  print(sprintf('Done with %i datasets...', j))
  print(end - start)
  
}

```




```{r}
gene_means = rowMeans(all_gene_fetal_pres_fetal_mat)
gene_sds =apply(all_gene_fetal_pres_fetal_mat, MARGIN = 1, FUN = sd)
num_dataset_gene_zero_exp = rowSums(all_gene_fetal_pres_fetal_mat == .5)
table(num_dataset_gene_zero_exp == 0)

exp_genes_index = which(num_dataset_gene_zero_exp == 0)

plot(gene_means[exp_genes_index], gene_sds[exp_genes_index], cex = .5)
plot(gene_sds[exp_genes_index], gene_means[exp_genes_index], cex = .5)

```




```{r}


fetal_all_gene_presCoexp_df = data.frame(gene = go_genes, mean_fetal_preCoexp = gene_means, sd_fetal_preCoexp = gene_sds, num_dataset_gene_not_expressed = num_dataset_gene_zero_exp)

fetal_all_gene_presCoexp_df %>% filter(num_dataset_gene_not_expressed == 0) %>% arrange(desc(mean_fetal_preCoexp))
fetal_all_gene_presCoexp_df %>% filter(num_dataset_gene_not_expressed == 0) %>% arrange(mean_fetal_preCoexp)

```


```{r}
save(org_all_gene_presCoexp_df, fetal_all_gene_presCoexp_df, file = "../data/coexpression/all_gene_presCoexp/all_gene_presCoexp_mats_v3.Rdata")

```



```{r}
load(file = "../../../data/coexpression/all_gene_presCoexp/all_gene_presCoexp_mats_v3.Rdata")

```


```{r}

org_exp_df = org_all_gene_presCoexp_df %>% filter(num_dataset_gene_not_expressed <= 5) %>% arrange(desc(mean_org_preCoexp))
fetal_exp_df = fetal_all_gene_presCoexp_df %>% filter(num_dataset_gene_not_expressed <= 5) %>% arrange(desc(mean_fetal_preCoexp))

shared_genes = intersect(fetal_exp_df$gene, org_exp_df$gene)

length(shared_genes)

org_exp_df = org_exp_df[org_exp_df$gene %in% shared_genes, ]
fetal_exp_df = fetal_exp_df[fetal_exp_df$gene %in% shared_genes, ]


#Reorder organoid to match fetal
index = match(fetal_exp_df$gene, org_exp_df$gene)
org_exp_df = org_exp_df[index, ]


table(org_exp_df$gene == fetal_exp_df$gene)

bad_org_index = org_exp_df$mean_org_preCoexp< .7 & fetal_exp_df$mean_fetal_preCoexp >= .99

par(pty = 's')
plot(fetal_exp_df$mean_fetal_preCoexp, org_exp_df$mean_org_preCoexp, cex = .5, xlim = c(0,1), ylim = c(0,1))
points(fetal_exp_df$mean_fetal_preCoexp[bad_org_index], org_exp_df$mean_org_preCoexp[bad_org_index], col ='red', cex = .75, pch = 16)
abline(a = 0, b = 1,col = 'red')
#abline(v = .99, col = 'blue', lty = 'dashed')
#abline(h = .7, col = 'blue', lty = 'dashed')

#pdf(file = 'graphs/organoid_ind_gene_presCoexp/ind_global_gene_mean_presCoexp_scatter_v3.pdf', height = 4, width = 4, useDingbats = F)
#par(pty = 's')
#plot(fetal_exp_df$mean_fetal_preCoexp, org_exp_df$mean_org_preCoexp, cex = .5, xlim = c(0,1), ylim = c(0,1))
#points(fetal_exp_df$mean_fetal_preCoexp[bad_org_index], org_exp_df$mean_org_preCoexp[bad_org_index], col ='red', cex = .75, pch = 16)
#abline(a = 0, b = 1,col = 'red')
#dev.off()

```


```{r}

org_fetal_exp_df = fetal_exp_df
org_fetal_exp_df$mean_org_preCoexp = org_exp_df$mean_org_preCoexp


bad_org_df = org_fetal_exp_df %>% filter(mean_fetal_preCoexp >= .99, mean_org_preCoexp < .7)
bad_org_df$gene
```

```{r}

source('/home/werner/tools/GO_enrichment_human/GO_enrichment.R')
load('/home/werner/tools/GO_enrichment_human/go_tables.Rdata')

```

Make a bar chart for the terms and pull out the specific ECM genes as examples

```{r}


go_enrich_results = gene_set_enrichment(bad_org_df$gene, BP_GO_mat)

go_enrich_results_filt = filter(go_enrich_results, N_univ <= 500 & N_univ >= 10) %>% arrange(adj_pvals)

top_10_go = go_enrich_results_filt[1:10 , ]
top_10_go = top_10_go %>% arrange(desc(adj_pvals))
top_10_go$description = factor(top_10_go$description, levels = c(top_10_go$description))
ggplot(top_10_go, aes(x = -log10(adj_pvals), y = description)) + geom_bar(stat = 'identity') +
  geom_vline(xintercept = -log10(.05), col = 'red', linetype = 'dashed')
#ggsave(filename = 'bad_org_go_terms_barplot_v3.pdf', path = 'graphs/organoid_ind_gene_presCoexp/', device = 'pdf', useDingbats = F, width = 6, height = 4)

top_10_go


save(go_enrich_results, file = '/home/werner/projects/meta_qc_organoid/data/coexpression/all_gene_presCoexp/full_data_GO_enrich_bad_org_genes.Rdata')


```

```{r}

go_enrich_results_filt

```




look at the genes in the ECM terms

```{r}

go_table %>% filter(gene_symbol %in% bad_org_df$gene & go_id %in% top_10_go$GO_term)

go_table %>% filter(gene_symbol %in% bad_org_df$gene & go_id == 'GO:0071711')
go_table %>% filter(gene_symbol %in% bad_org_df$gene & go_id == 'GO:0031589')
go_table %>% filter(gene_symbol %in% bad_org_df$gene & go_id == 'GO:0043062')
go_table %>% filter(gene_symbol %in% bad_org_df$gene & go_id == 'GO:0034446')
go_table %>% filter(gene_symbol %in% bad_org_df$gene & go_id == 'GO:0030198')
go_table %>% filter(gene_symbol %in% bad_org_df$gene & go_id == 'GO:0045229')

ecm_genes = unique(filter(go_table, gene_symbol %in% bad_org_df$gene & go_id %in% c('GO:0071711','GO:0031589', 'GO:0043062', 'GO:0034446','GO:0030198','GO:0045229' ))$gene_symbol)
ecm_genes
```




Run an experiment to see if this is specific to neurons

Make co-expression networks of the primary tissue data just using the Glut and GABA neurons, make an aggregate too.
Then do the leave one out preserved co-expression quantification using just these networks
Are the ECM genes still highly preserved? How do the organoids compare to the just neuron networks?


Load in data and filter for the GO genes
Use python to get coexpression matrix
```{r}
Sys.setenv(RETICULATE_PYTHON = "/usr/bin/python3.6")
library(reticulate)
np <- reticulate::import("numpy")

source_python('python_coexpression.py')
```


```{r}

human_fetal_dataset_paths = c('../data/seurat_objs/individual/fan_fu_seurat_processed.Rdata',
                             '../data/seurat_objs/individual/poliodakis_geschwind_batch1_seurat_processed.Rdata',
                             '../data/seurat_objs/individual/poliodakis_geschwind_batch2_seurat_processed.Rdata',
                             '../data/seurat_objs/individual/areal_kreigstein_dataset/merged_with_metadata/GW14.Rdata', 
                             '../data/seurat_objs/individual/areal_kreigstein_dataset/merged_with_metadata/GW18_2.Rdata', 
                             '../data/seurat_objs/individual/areal_kreigstein_dataset/merged_with_metadata/GW19.Rdata', 
                             '../data/seurat_objs/individual/areal_kreigstein_dataset/merged_with_metadata/GW19_2.Rdata', 
                             '../data/seurat_objs/individual/areal_kreigstein_dataset/merged_with_metadata/GW20.Rdata', 
                             '../data/seurat_objs/individual/areal_kreigstein_dataset/merged_with_metadata/GW20_31.Rdata', 
                             '../data/seurat_objs/individual/areal_kreigstein_dataset/merged_with_metadata/GW20_34.Rdata',
                             '../data/seurat_objs/individual/areal_kreigstein_dataset/merged_with_metadata/GW25.Rdata',
                              '../data/seurat_objs/individual/linnarsson_GW5_processed.Rdata',
                              '../data/seurat_objs/individual/linnarsson_GW5-5_processed.Rdata',
                              '../data/seurat_objs/individual/linnarsson_GW6_1_processed.Rdata',
                              '../data/seurat_objs/individual/linnarsson_GW6_2_processed.Rdata',
                              '../data/seurat_objs/individual/linnarsson_GW6-6_1_processed.Rdata',
                              '../data/seurat_objs/individual/linnarsson_GW6-6_2_processed.Rdata',
                              '../data/seurat_objs/individual/linnarsson_GW6-7_processed.Rdata',
                              '../data/seurat_objs/individual/linnarsson_GW6-9_1_processed.Rdata',
                              '../data/seurat_objs/individual/linnarsson_GW6-9_2_processed.Rdata',
                              '../data/seurat_objs/individual/linnarsson_GW6-9_3_processed.Rdata',
                              '../data/seurat_objs/individual/linnarsson_GW6-9_4_processed.Rdata',
                              '../data/seurat_objs/individual/linnarsson_GW7_processed.Rdata',
                              '../data/seurat_objs/individual/linnarsson_GW7-5_processed.Rdata',
                              '../data/seurat_objs/individual/linnarsson_GW8_1_processed.Rdata',
                              '../data/seurat_objs/individual/linnarsson_GW8_2_processed.Rdata',
                              '../data/seurat_objs/individual/linnarsson_GW8_3_processed.Rdata',
                              '../data/seurat_objs/individual/linnarsson_GW8-1_processed.Rdata',
                              '../data/seurat_objs/individual/linnarsson_GW8-5_1_processed.Rdata',
                              '../data/seurat_objs/individual/linnarsson_GW8-5_2_processed.Rdata',
                              '../data/seurat_objs/individual/linnarsson_GW9-2_processed.Rdata',
                              '../data/seurat_objs/individual/linnarsson_GW9-5_processed.Rdata',
                              '../data/seurat_objs/individual/linnarsson_GW10_processed.Rdata',
                              '../data/seurat_objs/individual/linnarsson_GW11-5_processed.Rdata',
                              '../data/seurat_objs/individual/linnarsson_GW12_processed.Rdata',
                              '../data/seurat_objs/individual/linnarsson_GW13_processed.Rdata',
                              '../data/seurat_objs/individual/linnarsson_GW14_processed.Rdata')

human_fetal_dataset_names = c('fan_fu','poliodakis_batch1', 'poliodakis_batch2','areal_GW14','areal_GW18_2','areal_GW19','areal_GW19_2', 'areal_GW20','areal_GW20_31','areal_GW20_34','areal_GW25',
                'linnarsson_GW5', 'linnarsson_GW5-5','linnarsson_GW6_1','linnarsson_GW6_2','linnarsson_GW6-6_1','linnarsson_GW6-6_2','linnarsson_GW6-7',
                'linnarsson_GW6-9_1','linnarsson_GW6-9_2',
                'linnarsson_GW6-9_3','linnarsson_GW6-9_4','linnarsson_GW7','linnarsson_GW7-5','linnarsson_GW8_1','linnarsson_GW8_2','linnarsson_GW8_3','linnarsson_GW8-1',
                'linnarsson_GW8-5_1',
                'linnarsson_GW8-5_2','linnarsson_GW9-2','linnarsson_GW9-5','linnarsson_GW10','linnarsson_GW11-5','linnarsson_GW12','linnarsson_GW13','linnarsson_GW14')

```


```{r}
for(i in 1:length(human_fetal_dataset_paths)){

  start = Sys.time()
  x = load(human_fetal_dataset_paths[[i]])
  dataset = get(x)

  dataset_genes = rownames(dataset)
  #Filter for the GO gene universe
  dataset_genes_in_go = dataset_genes[dataset_genes %in% go_genes]
  dataset_genes_missing_from_go = go_genes[! go_genes %in% dataset_genes]
  #Subset the gene expression for the genes present in GO
  dataset_gene_index = dataset_genes %in% dataset_genes_in_go
  exp_data = as.matrix(dataset@assays$RNA@data[dataset_gene_index, ])
  #Add zero counts for the missing GO genes
  go_missing = matrix(0, nrow = length(dataset_genes_missing_from_go), ncol = ncol(exp_data), dimnames = list(dataset_genes_missing_from_go, colnames(exp_data)))
  exp_data = rbind(exp_data, go_missing)
  r_genes = rownames(exp_data)
  
  #Subset for just the neurons
  cell_index = !dataset$labeled_classes %in% c('other','Microglia/Immune')
  exp_data = exp_data[ , cell_index]

  suppressWarnings(rm(dataset, GW18_2_merged_dataset, poliodakis_dataset_batch1,poliodakis_dataset_batch2, fan_fu_dataset, GW14_merged_dataset,
     GW19_merged_dataset,GW19_2_merged_dataset,GW20_merged_dataset,GW20_31_merged_dataset,GW20_34_merged_dataset,GW25_merged_dataset, seurat_object))
  gc()
  
  #Get coexpression data
  test_py_exp = np$array(exp_data) #Turn R matrix into numpy array
  python_script_test = get_coexpression(test_py_exp, r_genes) #Python method for spearman correlations, coexpression
  corr_matrix = as.matrix(python_script_test) #Turn to matrix
  #Convert Nans to 0, these are genes with 0 expression in any cells
  corr_matrix[!is.finite(corr_matrix)] = 0
  
  file_name = sprintf('../data/coexpression/fetal/just_neurons/%s_neuron_coexp.Rdata', human_fetal_dataset_names[i])
  save(corr_matrix, file = file_name)
  end = Sys.time()
  print(end - start)
  print(sprintf('Done getting coexpression from %i dataset...', i))
  
  #Rank the coexpression data
  #Rank everything, I want the minimum rank to be 1, so ties are minimum. Subtracting 1 sets the minimum rank to 0
  rank_corr = frank(c(corr_matrix), ties.method = 'min') - 1
  #Divide by max rank squeezes everything between 0 and 1
  norm_rank_corr = rank_corr / max(rank_corr)
  #Put back into a matrix
  rank_corr_matrix = matrix(norm_rank_corr, nrow = nrow(corr_matrix), ncol = ncol(corr_matrix), dimnames = list(rownames(corr_matrix), colnames(corr_matrix)))
  diag(rank_corr_matrix) = 1
  
  file_name = sprintf('../data/coexpression/fetal/just_neurons/ranked_%s_neuron_coexp.Rdata', human_fetal_dataset_names[i])
  save(rank_corr_matrix, file =file_name)
  end = Sys.time()
  print(end - start)
  print(sprintf('Done ranking coexpression from %i dataset...', i))  
}
```


Aggregate the just neuron co-expression networks

```{r}
path_start = '/home/werner/projects/meta_qc_organoid/data/coexpression/fetal/just_neurons/'

just_fetal_ranked_go_coexp_paths = list.files(path_start)
just_fetal_ranked_go_coexp_paths  = just_fetal_ranked_go_coexp_paths[grepl('ranked_', just_fetal_ranked_go_coexp_paths )]
just_fetal_ranked_go_coexp_paths = paste0(path_start, just_fetal_ranked_go_coexp_paths )
just_fetal_ranked_go_coexp_paths  = just_fetal_ranked_go_coexp_paths[!grepl('aggregate', just_fetal_ranked_go_coexp_paths )]

sum_rank_corr_genes = matrix(rep(0, length(go_genes)^2), nrow = length(go_genes), ncol = length(go_genes), dimnames = list(go_genes, go_genes))
for(i in 1:length(just_fetal_ranked_go_coexp_paths)){
  x = load(just_fetal_ranked_go_coexp_paths[[i]])
  rank_mat = get(x)
  rm(x)
  #Reorder to match the same gene order
  index = match(go_genes, rownames(rank_mat))
  rank_mat = rank_mat[index, index]
  sum_rank_corr_genes = sum_rank_corr_genes + rank_mat
}
#Divide sums by the number of datasets to get averages
agg_rank_mat = sum_rank_corr_genes / length(just_fetal_ranked_go_coexp_paths)

#Rank
rank_corr = frank(c(agg_rank_mat), ties.method = 'min') - 1
#Divide by max rank squeezes everything between 0 and 1
norm_rank_corr = rank_corr / max(rank_corr)
#Put back into a matrix
rank_corr_matrix = matrix(norm_rank_corr, nrow = nrow(agg_rank_mat), ncol = ncol(agg_rank_mat), 
                          dimnames = list(rownames(agg_rank_mat), colnames(agg_rank_mat)))
agg_rank_mat = rank_corr_matrix
diag(agg_rank_mat) = 1
print('Done aggregating networks...')

save(agg_rank_mat, file ='../data/coexpression/fetal/just_neurons/ranked_aggregate_annot_fetal_neuron_coexp.Rdata')

```


Do a leave one out on the preservation of co-expression with just the neurons

```{r}

human_fetal_neuron_names= c('areal_GW14','areal_GW18_2','areal_GW19_2','areal_GW19',
                              'areal_GW20_31','areal_GW20_34','areal_GW20','areal_GW25','fan_fu','linnarsson_GW10','linnarsson_GW11-5','linnarsson_GW12',
                              'linnarsson_GW13','linnarsson_GW14',
                              'linnarsson_GW5', 'linnarsson_GW5-5','linnarsson_GW6_1','linnarsson_GW6_2','linnarsson_GW6-6_1','linnarsson_GW6-6_2','linnarsson_GW6-7',
                              'linnarsson_GW6-9_1','linnarsson_GW6-9_2','linnarsson_GW6-9_3','linnarsson_GW6-9_4','linnarsson_GW7','linnarsson_GW7-5','linnarsson_GW8_1',
                              'linnarsson_GW8_2','linnarsson_GW8_3','linnarsson_GW8-1','linnarsson_GW8-5_1','linnarsson_GW8-5_2','linnarsson_GW9-2','linnarsson_GW9-5',
                              'poliodakis_batch1', 'poliodakis_batch2')
names(just_fetal_ranked_go_coexp_paths) = human_fetal_neuron_names

top_genes = 10
all_gene_fetal_neuron_pres_mat = matrix(nrow = length(go_genes), ncol = length(human_fetal_neuron_names), dimnames = list(go_genes, human_fetal_neuron_names))

for(j in 1:length(human_fetal_neuron_names)){

  start = Sys.time()
  #current_dataset = annot_marker_names[j]
  #Leave one out and get the new aggregated fetal network
  sum_rank_corr_genes = matrix(rep(0, length(go_genes)^2), nrow = length(go_genes), ncol = length(go_genes), dimnames = list(go_genes, go_genes))
  current_ranked_go_coexp_paths = just_fetal_ranked_go_coexp_paths[c(-j)]
  
  for(i in 1:length(current_ranked_go_coexp_paths)){
    x = load(current_ranked_go_coexp_paths[[i]])
    rank_mat = get(x)
    rm(x)
    #Reorder to match the same gene order
    index = match(go_genes, rownames(rank_mat))
    rank_mat = rank_mat[index, index]
    sum_rank_corr_genes = sum_rank_corr_genes + rank_mat
  }
  #Divide sums by the number of datasets to get averages
  agg_rank_mat = sum_rank_corr_genes / length(current_ranked_go_coexp_paths)
  
  #Rank
  rank_corr = frank(c(agg_rank_mat), ties.method = 'min') - 1
  #Divide by max rank squeezes everything between 0 and 1
  norm_rank_corr = rank_corr / max(rank_corr)
  #Put back into a matrix
  rank_corr_matrix = matrix(norm_rank_corr, nrow = nrow(agg_rank_mat), ncol = ncol(agg_rank_mat), 
                            dimnames = list(rownames(agg_rank_mat), colnames(agg_rank_mat)))
  agg_rank_mat = rank_corr_matrix
  diag(agg_rank_mat) = 1
  print('Done aggregating networks...')
  
  #Get the preserved coexpression aurocs for all genes
  x = load(just_fetal_ranked_go_coexp_paths[[j]])
  rank_mat = get(x)
  rm(x)
  
  test_all_genes = unlist(mclapply(1:length(go_genes), function(k) get_conserved_coexp(agg_rank_mat, rank_mat, go_genes[k],top_genes ), mc.cores = 10 ))
  all_gene_fetal_neuron_pres_mat[ ,j] = test_all_genes
  
  end = Sys.time()
  print(sprintf('Done with %i datasets...', j))
  print(end - start)
  
}



```



```{r}


save(all_gene_fetal_neuron_pres_mat, file = "../../../data/coexpression/all_gene_presCoexp/all_gene_just_neuron_presCoexp_fetal.Rdata")

```

```{r}


load(file = "../../../data/coexpression/all_gene_presCoexp/all_gene_just_neuron_presCoexp_fetal.Rdata")

```



```{r}

all_gene_fetal_neuron_pres_mat[1:10, 1:10]


ecm_genes = c("CAV1"    ,"CAV2"    ,"RUNX1"   ,"CD36"    ,"COL1A2"  ,"COL4A1" , "CTSK"   , "EFNA1"  , "FN1"  ,   "ITGA1"  , "ITGA4"  , "LAMB1"  , "LAMC1" ,  "NID1"  ,  "DDR2"  ,
              "RRAS"  ,  "DLC1",
              "MERTK"  , "EMILIN1", "DISC1" ,  "PARVB" ,  "VWA1" ,   "MMP28" ,  "SPRY4" ,  "OLFML2A")

hist(all_gene_fetal_neuron_pres_mat[ , 1], breaks = 60)

table(all_gene_fetal_neuron_pres_mat[ , 1] == 1)



rowMeans(all_gene_fetal_neuron_pres_mat[ecm_genes , ])
```



Now compute organoid preservation of co-expression using the primary tissue just neurons



```{r}

all_sample_meta = read.csv('../data/seurat_objs/all_data_just_meta_seurat_with_addl_features_v2.csv')
all_sample_meta

```

```{r}

x = load( file ='../data/coexpression/fetal/just_neurons/ranked_aggregate_annot_fetal_neuron_coexp.Rdata')
ranked_aggregate_just_neurons = get(x)

rm(agg_rank_mat, x)
gc()

ranked_aggregate_just_neurons[1:10, 1:10]

```




```{r}
org_all_gene_fetal_neuron_pres_mat = matrix(nrow = length(go_genes), ncol = nrow(all_sample_meta), dimnames = list(go_genes, all_sample_meta$sample_ids))
top = 10


for(j in 1:nrow(all_sample_meta)){

  start = Sys.time()
  load(sprintf('../data/coexpression/organoid/6_26_24_GO_coexp/ranked_%s_coexp.Rdata', all_sample_meta$sample_ids[j]))
  
  org_all_gene_fetal_neuron_pres_mat[ ,j] = unlist(mclapply(1:length(go_genes), function(i) get_conserved_coexp(ranked_aggregate_just_neurons, rank_corr_matrix, 
                                                                                     go_genes[i], num_top = top ), mc.cores = 10 ))
  
  print(Sys.time() - start)
  print(sprintf('Done with %i datasets...', j))

}


```

```{r}


save(org_all_gene_fetal_neuron_pres_mat, file = "../../../data/coexpression/all_gene_presCoexp/all_gene_just_neuron_presCoexp_organoid.Rdata")

```

```{r}


load(file = "../../../data/coexpression/all_gene_presCoexp/all_gene_just_neuron_presCoexp_organoid.Rdata")

```




```{r}

rowMeans(org_all_gene_fetal_neuron_pres_mat[ecm_genes , ])

```


```{r}
gene_means = rowMeans(all_gene_fetal_neuron_pres_mat)
gene_sds =apply(all_gene_fetal_neuron_pres_mat, MARGIN = 1, FUN = sd)
num_dataset_gene_zero_exp = rowSums(all_gene_fetal_neuron_pres_mat == .5)
table(num_dataset_gene_zero_exp == 0)

exp_genes_index = which(num_dataset_gene_zero_exp == 0)

plot(gene_means[exp_genes_index], gene_sds[exp_genes_index], cex = .5)
plot(gene_sds[exp_genes_index], gene_means[exp_genes_index], cex = .5)


fetal_all_gene_just_neuron_presCoexp_df = data.frame(gene = go_genes, mean_fetal_preCoexp = gene_means, sd_fetal_preCoexp = gene_sds, num_dataset_gene_not_expressed = num_dataset_gene_zero_exp)

fetal_all_gene_just_neuron_presCoexp_df %>% filter(num_dataset_gene_not_expressed == 0) %>% arrange(desc(mean_fetal_preCoexp))
fetal_all_gene_just_neuron_presCoexp_df %>% filter(num_dataset_gene_not_expressed == 0) %>% arrange(mean_fetal_preCoexp)

```


```{r}
gene_means = rowMeans(org_all_gene_fetal_neuron_pres_mat)
gene_sds =apply(org_all_gene_fetal_neuron_pres_mat, MARGIN = 1, FUN = sd)
num_dataset_gene_zero_exp = rowSums(org_all_gene_fetal_neuron_pres_mat == .5)
table(num_dataset_gene_zero_exp == 0)

exp_genes_index = which(num_dataset_gene_zero_exp == 0)

plot(gene_means[exp_genes_index], gene_sds[exp_genes_index], cex = .5)
plot(gene_sds[exp_genes_index], gene_means[exp_genes_index], cex = .5)


org_all_gene_just_neuron_presCoexp_df = data.frame(gene = go_genes, mean_org_preCoexp = gene_means, sd_org_preCoexp = gene_sds, num_dataset_gene_not_expressed = num_dataset_gene_zero_exp)

org_all_gene_just_neuron_presCoexp_df %>% filter(num_dataset_gene_not_expressed == 0) %>% arrange(desc(mean_org_preCoexp))
org_all_gene_just_neuron_presCoexp_df %>% filter(num_dataset_gene_not_expressed == 0) %>% arrange(mean_org_preCoexp)

```



```{r}

just_neuron_org_exp_df = org_all_gene_just_neuron_presCoexp_df %>% filter(num_dataset_gene_not_expressed <=5) %>% arrange(desc(mean_org_preCoexp))
just_neuron_fetal_exp_df = fetal_all_gene_just_neuron_presCoexp_df %>% filter(num_dataset_gene_not_expressed <= 5) %>% arrange(desc(mean_fetal_preCoexp))

shared_genes = intersect(just_neuron_fetal_exp_df$gene, just_neuron_org_exp_df$gene)

length(shared_genes)

just_neuron_org_exp_df = just_neuron_org_exp_df[just_neuron_org_exp_df$gene %in% shared_genes, ]
just_neuron_fetal_exp_df = just_neuron_fetal_exp_df[just_neuron_fetal_exp_df$gene %in% shared_genes, ]


#Reorder organoid to match fetal
index = match(just_neuron_fetal_exp_df$gene, just_neuron_org_exp_df$gene)
just_neuron_org_exp_df = just_neuron_org_exp_df[index, ]


table(just_neuron_org_exp_df$gene == just_neuron_fetal_exp_df$gene)

bad_org_index = just_neuron_org_exp_df$mean_org_preCoexp<= .75 & just_neuron_fetal_exp_df$mean_fetal_preCoexp >= .9


par(pty = 's')
plot(just_neuron_fetal_exp_df$mean_fetal_preCoexp, just_neuron_org_exp_df$mean_org_preCoexp, cex = .5, xlim = c(0,1), ylim = c(0,1))
points(just_neuron_fetal_exp_df$mean_fetal_preCoexp[bad_org_index], just_neuron_org_exp_df$mean_org_preCoexp[bad_org_index], col ='red', cex = .75, pch = 16)
abline(a = 0, b = 1,col = 'red')

#pdf(file = 'graphs/organoid_ind_gene_presCoexp/ind_global_gene_mean_presCoexp_scatter_just_neurons.pdf', height = 4, width = 4, useDingbats = F)
#par(pty = 's')
#plot(just_neuron_fetal_exp_df$mean_fetal_preCoexp, just_neuron_org_exp_df$mean_org_preCoexp, cex = .5, xlim = c(0,1), ylim = c(0,1))
#points(just_neuron_fetal_exp_df$mean_fetal_preCoexp[bad_org_index], just_neuron_org_exp_df$mean_org_preCoexp[bad_org_index], col ='red', cex = .75, pch = 16)
#abline(a = 0, b = 1,col = 'red')
#dev.off()


save(just_neuron_fetal_exp_df, just_neuron_org_exp_df, file = '/home/werner/projects/meta_qc_organoid/data/coexpression/all_gene_presCoexp/just_neuron_exp_df.Rdata')



```





```{r}
just_neuron_org_fetal_exp_df = just_neuron_fetal_exp_df
just_neuron_org_fetal_exp_df$mean_org_preCoexp = just_neuron_org_exp_df$mean_org_preCoexp

test_genes = just_neuron_org_fetal_exp_df %>% filter(mean_org_preCoexp <= .7 & mean_fetal_preCoexp >= .85) %>% pull(gene)
test_genes



go_enrich_results = gene_set_enrichment(test_genes, BP_GO_mat)


test_pvals = p.adjust(go_enrich_results$pvals, method = 'BH')

hist(go_enrich_results$pvals)


go_enrich_results_filt = filter(go_enrich_results, N_univ <= 500 & N_univ >= 10) %>% arrange(pvals)
go_enrich_results_filt 

top_10_go = go_enrich_results_filt[1:10 , ]
top_10_go = top_10_go %>% arrange(desc(adj_pvals))
top_10_go$description = factor(top_10_go$description, levels = c(top_10_go$description))
ggplot(top_10_go, aes(x = -log10(adj_pvals), y = description)) + geom_bar(stat = 'identity') +
  geom_vline(xintercept = -log10(.05), col = 'red', linetype = 'dashed')
#ggsave(filename = 'bad_org_go_terms_just_neurons_barplot_v3.pdf', path = 'graphs/organoid_ind_gene_presCoexp/', device = 'pdf', useDingbats = F, width = 6, height = 4)

top_10_go



just_neuron_go_enrich_results = go_enrich_results
save(just_neuron_go_enrich_results , file = '/home/werner/projects/meta_qc_organoid/data/coexpression/all_gene_presCoexp/just_neurons_GO_enrich_bad_org_genes.Rdata')
```



```{r}
example_genes = go_table %>% filter(gene_symbol %in% test_genes & go_id %in% c('GO:0043062','GO:0030198','GO:0045229', 'GO:0031589', 'GO:0070831')) %>% pull(gene_symbol)
ecm_genes_neurons = example_genes[!duplicated(example_genes)]





```



```{r}
intersect(ecm_genes, ecm_genes_neurons)

```



Check out DISC1 a bit further



```{r}
all_gene_fetal_neuron_pres_mat['DISC1', ]
org_all_gene_fetal_neuron_pres_mat['DISC1', ]


all_org_disc1_stats = matrix(data = NA, nrow = nrow(all_sample_meta), ncol = 4, dimnames = list(all_sample_meta$sample_ids, 
                                                                                                c('num_cells_disc1','num_cells_disc1_ratio','mean_disc1_exp','median_disc1_exp')))
for(i in 1:nrow(all_sample_meta)){

  x = load(all_sample_meta$processed_seurat_path[i])
  dataset = get(x)
  suppressWarnings(rm(x, seurat_object))
  gc()
  
  if('DISC1' %in% rownames(dataset)){
  
    disc_exp = FetchData(dataset, 'DISC1')
  
    all_org_disc1_stats[i,'num_cells_disc1'] = sum(disc_exp$DISC1 != 0)
    all_org_disc1_stats[i,'num_cells_disc1_ratio'] = sum(disc_exp$DISC1 != 0) / nrow(disc_exp)
    all_org_disc1_stats[i,'mean_disc1_exp'] = mean(filter(disc_exp, DISC1 != 0)$DISC1)
    all_org_disc1_stats[i,'median_disc1_exp'] = median(filter(disc_exp, DISC1 != 0)$DISC1)
  }else{next}
  print(i)
  
}

all_org_disc1_stats = data.frame(all_org_disc1_stats)
all_org_disc1_stats$disc1_pres_coexp = org_all_gene_fetal_neuron_pres_mat['DISC1', ]
all_org_disc1_stats$normalization = all_sample_meta$Normalization.for.this.study

norm_index = all_org_disc1_stats$normalization == 'CPM'

ggplot(all_org_disc1_stats[norm_index, ], aes(x = log10(mean_disc1_exp +1), y = disc1_pres_coexp)) + geom_point()



```

And now the primary tissue data

```{r}

human_fetal_dataset_names = c( 
                               'linnarsson_GW5','linnarsson_GW5-5','linnarsson_GW6_1','linnarsson_GW6_2','linnarsson_GW6-6_1',
                               'linnarsson_GW6-6_2','linnarsson_GW6-7','linnarsson_GW6-9_1','linnarsson_GW6-9_2','linnarsson_GW6-9_3',
                               'linnarsson_GW6-9_4','linnarsson_GW7','linnarsson_GW7-5','linnarsson_GW8_1','linnarsson_GW8_2',
                               'linnarsson_GW8_3','linnarsson_GW8-1','linnarsson_GW8-5_1','linnarsson_GW8-5_2','linnarsson_GW9-2',
                               'linnarsson_GW9-5','linnarsson_GW10','linnarsson_GW11-5','linnarsson_GW12','linnarsson_GW13', 'linnarsson_GW14',
                               'areal_GW14','poliodakis_geschwind_batch1', 'poliodakis_geschwind_batch2','areal_GW18_2','areal_GW19','areal_GW19_2',
                               'areal_GW20','areal_GW20_31','areal_GW20_34','areal_GW25','fan_fu','areal_GW16','areal_GW18', 'areal_GW22','areal_GW22T','shi_wang','zhou_lu',
                               'yu_sun_1','yu_sun_2','yu_sun_3','yu_sun_4',
                               'trevino_greenleaf_1','trevino_greenleaf_2','trevino_greenleaf_3','trevino_greenleaf_4')

human_fetal_dataset_paths = list(
                                 '../data/seurat_objs/individual/linnarsson_GW5_processed.Rdata',
                                 '../data/seurat_objs/individual/linnarsson_GW5-5_processed.Rdata',
                                 '../data/seurat_objs/individual/linnarsson_GW6_1_processed.Rdata',
                                 '../data/seurat_objs/individual/linnarsson_GW6_2_processed.Rdata',
                                 '../data/seurat_objs/individual/linnarsson_GW6-6_1_processed.Rdata',
                                 '../data/seurat_objs/individual/linnarsson_GW6-6_2_processed.Rdata',
                                 '../data/seurat_objs/individual/linnarsson_GW6-7_processed.Rdata',
                                 '../data/seurat_objs/individual/linnarsson_GW6-9_1_processed.Rdata',
                                 '../data/seurat_objs/individual/linnarsson_GW6-9_2_processed.Rdata',
                                 '../data/seurat_objs/individual/linnarsson_GW6-9_3_processed.Rdata',
                                 '../data/seurat_objs/individual/linnarsson_GW6-9_4_processed.Rdata',
                                 '../data/seurat_objs/individual/linnarsson_GW7_processed.Rdata',
                                 '../data/seurat_objs/individual/linnarsson_GW7-5_processed.Rdata',
                                 '../data/seurat_objs/individual/linnarsson_GW8_1_processed.Rdata',
                                 '../data/seurat_objs/individual/linnarsson_GW8_2_processed.Rdata',
                                 '../data/seurat_objs/individual/linnarsson_GW8_3_processed.Rdata',
                                 '../data/seurat_objs/individual/linnarsson_GW8-1_processed.Rdata',
                                 '../data/seurat_objs/individual/linnarsson_GW8-5_1_processed.Rdata',
                                 '../data/seurat_objs/individual/linnarsson_GW8-5_2_processed.Rdata',
                                 '../data/seurat_objs/individual/linnarsson_GW9-2_processed.Rdata',
                                 '../data/seurat_objs/individual/linnarsson_GW9-5_processed.Rdata',
                                 '../data/seurat_objs/individual/linnarsson_GW10_processed.Rdata',
                                 '../data/seurat_objs/individual/linnarsson_GW11-5_processed.Rdata',
                                 '../data/seurat_objs/individual/linnarsson_GW12_processed.Rdata',
                                 '../data/seurat_objs/individual/linnarsson_GW13_processed.Rdata',
                                 '../data/seurat_objs/individual/linnarsson_GW14_processed.Rdata',
                                 '../data/seurat_objs/individual/areal_kreigstein_dataset/merged_with_metadata/GW14.Rdata',
                                 '../data/seurat_objs/individual/poliodakis_geschwind_batch1_seurat_processed.Rdata',
                                 '../data/seurat_objs/individual/poliodakis_geschwind_batch2_seurat_processed.Rdata',
                                 '../data/seurat_objs/individual/areal_kreigstein_dataset/merged_with_metadata/GW18_2.Rdata', 
                                 '../data/seurat_objs/individual/areal_kreigstein_dataset/merged_with_metadata/GW19.Rdata', 
                                 '../data/seurat_objs/individual/areal_kreigstein_dataset/merged_with_metadata/GW19_2.Rdata', 
                                 '../data/seurat_objs/individual/areal_kreigstein_dataset/merged_with_metadata/GW20.Rdata', 
                                 '../data/seurat_objs/individual/areal_kreigstein_dataset/merged_with_metadata/GW20_31.Rdata', 
                                 '../data/seurat_objs/individual/areal_kreigstein_dataset/merged_with_metadata/GW20_34.Rdata',
                                 '../data/seurat_objs/individual/areal_kreigstein_dataset/merged_with_metadata/GW25.Rdata',
                                 '../data/seurat_objs/individual/fan_fu_seurat_processed.Rdata',
                                 '../data/seurat_objs/individual/areal_kreigstein_dataset/merged_with_metadata/GW16.Rdata',
                                 '../data/seurat_objs/individual/areal_kreigstein_dataset/merged_with_metadata/GW18.Rdata', 
                                 '../data/seurat_objs/individual/areal_kreigstein_dataset/merged_with_metadata/GW22.Rdata',
                                 '../data/seurat_objs/individual/areal_kreigstein_dataset/merged_with_metadata/GW22T.Rdata',
                                 '../data/seurat_objs/individual/shi_wang_seurat_processed.Rdata', 
                                 '../data/seurat_objs/individual/zhou_lu_seurat_processed.Rdata',
                                 '../data/seurat_objs/individual/yu_sun_GSM5032680_seurat_processed.Rdata',
                                 '../data/seurat_objs/individual/yu_sun_GSM5032681_seurat_processed.Rdata',
                                 '../data/seurat_objs/individual/yu_sun_GSM5032682_seurat_processed.Rdata',
                                 '../data/seurat_objs/individual/yu_sun_GSM5032683_seurat_processed.Rdata',
                                 '../data/seurat_objs/individual/trevino_greenleaf_batch1_seurat_processed.Rdata',
                                 '../data/seurat_objs/individual/trevino_greenleaf_batch2_seurat_processed.Rdata',
                                 '../data/seurat_objs/individual/trevino_greenleaf_batch3_seurat_processed.Rdata',
                                 '../data/seurat_objs/individual/trevino_greenleaf_batch4_seurat_processed.Rdata')


names(human_fetal_dataset_paths)= human_fetal_dataset_names 


all_fetal_disc1_stats = matrix(data = NA, nrow = length(human_fetal_dataset_paths ), ncol = 4, dimnames = list(human_fetal_dataset_names, 
                                                                                                c('num_cells_disc1','num_cells_disc1_ratio','mean_disc1_exp','median_disc1_exp')))
for(i in 1:length(human_fetal_dataset_paths )){

  x = load(human_fetal_dataset_paths[[i]])
  dataset = get(x)
  suppressWarnings(rm(x, seurat_object, fan_fu_dataset, poliodakis_dataset, GW14_merged_dataset, GW17_merged_dataset, GW18_2_merged_dataset,
                      GW19_merged_dataset,GW19_2_merged_dataset,
                      GW20_merged_dataset, GW20_31_merged_dataset, 
                      GW20_34_merged_dataset, GW25_merged_dataset, shi_wang_dataset,
                      zhou_lu_dataset,trevino_dataset_batch1,trevino_dataset_batch2,trevino_dataset_batch3,trevino_dataset_batch4,
                      GW16_merged_dataset, GW18_merged_dataset, GW22_merged_dataset, GW22T_merged_dataset))
  gc()
  
  if('DISC1' %in% rownames(dataset)){
  
    disc_exp = FetchData(dataset, 'DISC1')
  
    all_fetal_disc1_stats[i,'num_cells_disc1'] = sum(disc_exp$DISC1 != 0)
    all_fetal_disc1_stats[i,'num_cells_disc1_ratio'] = sum(disc_exp$DISC1 != 0) / nrow(disc_exp)
    all_fetal_disc1_stats[i,'mean_disc1_exp'] = mean(filter(disc_exp, DISC1 != 0)$DISC1)
    all_fetal_disc1_stats[i,'median_disc1_exp'] = median(filter(disc_exp, DISC1 != 0)$DISC1)
  }else{next}
  rm(dataset)
  gc()
  print(i)
  
}

all_fetal_disc1_stats = data.frame(all_fetal_disc1_stats)
#Only have data for the annotated datasets
all_fetal_disc1_stats = all_fetal_disc1_stats[1:37, ]


all_fetal_disc1_stats$disc1_pres_coexp = all_gene_fetal_neuron_pres_mat['DISC1', ]

rownames(all_fetal_disc1_stats)[rownames(all_fetal_disc1_stats) == 'poliodakis_geschwind_batch1'] = 'poliodakis_batch1'
rownames(all_fetal_disc1_stats)[rownames(all_fetal_disc1_stats) == 'poliodakis_geschwind_batch2'] = 'poliodakis_batch2'

index = match(rownames(all_fetal_disc1_stats), names(all_gene_fetal_neuron_pres_mat['DISC1', ]))
all_fetal_disc1_stats$disc1_pres_coexp = all_gene_fetal_neuron_pres_mat['DISC1', index ]

ggplot(all_fetal_disc1_stats, aes(x = mean_disc1_exp, y = disc1_pres_coexp)) + geom_point()






```

```{r}

org_for_joining = all_org_disc1_stats[norm_index, ][ ,1:5]
org_for_joining$data_label = rep('organoid', length = nrow(org_for_joining))
  
all_fetal_disc1_stats$data_label = rep('primary tissue', length = nrow(all_fetal_disc1_stats)) 

all_disc1_data = rbind(all_fetal_disc1_stats, org_for_joining)
   

save(all_disc1_data, file = "../data/coexpression/all_gene_presCoexp/disc1_data.Rdata")


ggplot(all_disc1_data, aes(x = log10(mean_disc1_exp +1), y = disc1_pres_coexp, color = data_label)) + geom_point() + xlab('log10(CPM + 1)') + 
  ylab('Preserved Coexpression DISC1 AUROC')

ggsave(filename = 'disc1_scatter.pdf', device = 'pdf', path = 'graphs/organoid_ind_gene_presCoexp/', useDingbats = F, width = 4, height = 4)


g1 = ggplot(all_disc1_data, aes(x = data_label, y = log10(mean_disc1_exp +1))) + geom_boxplot() + ylab('log10(CPM + 1)')
g2 = ggplot(all_disc1_data, aes(x = data_label, y = disc1_pres_coexp)) + geom_boxplot() + ylab('Preserved Coexpression DISC1 AUROC')
g3 = gridExtra::grid.arrange(g1, g2, nrow = 1)
ggsave(filename = 'disc1_boxplots.pdf', device = 'pdf', path = 'graphs/organoid_ind_gene_presCoexp/', useDingbats = F, width = 5, height = 4, plot = g3)
```







