

Processing the massive fetal arealization dataset from the kreigstein lab
each timepoint and area is saved in a different directory

```{r}

library(Seurat)
library(dplyr)
library(MetaMarkers)
library(data.table)
library(ggplot2)

```

Read in the metadata
cells are labeled by their age_region_barcode
This is metadata for only the filtered cells used in the analysis, so a subset of the cells present in each directory from the nemo archive

This is from Supplemental table 1 of the publication
```{r}

meta_file = '/home/werner/projects/meta_qc_organoid/datasets/arealization_fetal_kriegstein/areal_dataset_metadata.csv'

metadata = read.csv(meta_file, header = TRUE)
metadata


#Check for any mislabeled GW15 and GW16 cells
table(sapply(strsplit(metadata$cell.name, '_'), '[[',1))

```
```{r}
table(metadata$cell.type)
table(metadata$individual)

#Check for GW19_2 barcodes in the GW19 metadata

table(sapply(strsplit(filter(metadata, individual == 'GW19')$cell.name, '_', ),'[[', 2 ))

gw19_2_meta_index = which(sapply(strsplit(filter(metadata, individual == 'GW19')$cell.name, '_', ),'[[', 2 ) == 2)
gw19_meta_index = which(sapply(strsplit(filter(metadata, individual == 'GW19')$cell.name, '_', ),'[[', 2 ) != 2)

filter(metadata, individual == 'GW19')[gw19_2_meta_index, ]
filter(metadata, individual == 'GW19')[gw19_meta_index, ]

#GW19 metadata is for both GW19 and GW19_2, 54,237 cells in GW19_2

GW19_metadata = filter(metadata, individual == 'GW19')[gw19_meta_index, ]
GW19_2_metadata = filter(metadata, individual == 'GW19')[gw19_2_meta_index, ]
GW19_2_metadata$individual = rep('GW19_2', length = nrow(GW19_2_metadata))


#Some of the GW19 regions are mislabeled, GW10_hippocampus is labeled as V1
table(sapply(strsplit(GW19_metadata$cell.name, '_' ), '[[', 2 ))

hippo_index = which(sapply(strsplit(GW19_metadata$cell.name, '_' ), '[[', 2 ) == 'hippocampus')
hypothal_index = which(sapply(strsplit(GW19_metadata$cell.name, '_' ), '[[', 2 ) == 'hypothalamus')
thal_index = which(sapply(strsplit(GW19_metadata$cell.name, '_' ), '[[', 2 ) == 'thalamus')

GW19_metadata$area[hippo_index] = 'hippocampus'
GW19_metadata$area[hypothal_index] = 'hypothalamus'
GW19_metadata$area[thal_index] = 'thalamus'

#Check same for GW19_2
table(sapply(strsplit(GW19_2_metadata$cell.name, '_' ), '[[', 3 ))
table(GW19_2_metadata$area)

#Rename the GW19_2 metadata and add it all back in
metadata  = filter(metadata, individual != 'GW19')
table(metadata$individual)

metadata = rbind(metadata, GW19_metadata)
metadata = rbind(metadata, GW19_2_metadata)

table(metadata$individual)
```

individuals GW18, GW22, and GW22T did not have any matched metadata. Check for mislabeling and such
They all agree

```{r}

GW20_34_metadata = filter(metadata, individual == 'GW20_34')
table(GW20_34_metadata$area)

table(sapply(strsplit(GW20_34_metadata$cell.name, '_'),'[[',1 ))


table(sapply(strsplit(GW20_34_metadata$cell.name, '_'),'[[',2 ))

table(sapply(strsplit(GW20_34_metadata$cell.name, '_'),'[[',2 ) == GW20_34_metadata$area)
```




And this metadata is from the ucsc browser
Check for any differences between the two

```{r}

ucsc_meta_file = '/home/werner/projects/meta_qc_organoid/datasets/arealization_fetal_kriegstein/ucsc_areal_metadata.tsv'
ucsc_metadata = read.table(ucsc_meta_file, sep = '\t', header = T)
ucsc_metadata
```

change the cell_type column to match the spelling of the original metadata
and change some of the area names

```{r}
table(ucsc_metadata$cell_type)

ucsc_metadata$cell_type[ucsc_metadata$cell_type == 'astrocyte'] = 'Astrocyte'
ucsc_metadata$cell_type[ucsc_metadata$cell_type == 'dividing'] = 'Dividing'
ucsc_metadata$cell_type[ucsc_metadata$cell_type == 'endothelial'] = 'Endo'
ucsc_metadata$cell_type[ucsc_metadata$cell_type == 'intermediate progenitor cell'] = 'IPC'
ucsc_metadata$cell_type[ucsc_metadata$cell_type == 'interneuron'] = 'Interneuron'
ucsc_metadata$cell_type[ucsc_metadata$cell_type == 'microglia'] = 'Microglia'
ucsc_metadata$cell_type[ucsc_metadata$cell_type == 'neuron'] = 'Neuron'
ucsc_metadata$cell_type[ucsc_metadata$cell_type == 'oligodendrocyte'] = 'Oligo'
ucsc_metadata$cell_type[ucsc_metadata$cell_type == 'outlier'] = 'Outlier'
ucsc_metadata$cell_type[ucsc_metadata$cell_type == 'radial glia'] = 'RG'

table(ucsc_metadata$cell_type)

```
```{r}

table(ucsc_metadata$area)

ucsc_metadata$area[ucsc_metadata$area == 'caudal'] = 'CGE'
ucsc_metadata$area[ucsc_metadata$area == 'lateral'] = 'LGE'
ucsc_metadata$area[ucsc_metadata$area == 'medial'] = 'MGE'
ucsc_metadata$area[ucsc_metadata$area == 'nucleus accumbens'] = 'nucleusaccumbens'
ucsc_metadata$area[ucsc_metadata$area == 'prefrontal cortex'] = 'PFC'

table(ucsc_metadata$area)
```

all the cells in common between the two
```{r}

table(metadata$cell.name %in% ucsc_metadata$cell_name)

```


Checking barcodes, numi, ngenes, and the annotations match across the two for the different regions

the barcodes, numi, and ngenes all match across the two metadata sets, but the cell type annotations differ somewhat. The number that differ are below
The thalamus has a few thousand extra cells in the orginal metadata, thalamus cells from GW19 are missing from the UCSC metadata

CGE:  3611 cell type annotations mismatch
caudate: 1610 cell type annotations mismatch
cerebellum: 73
cingulate: 268
claustrum: 996
hippocampus: 582
hypothalamus: 4318
insula: 21
LGE: 12264
MGE: 5343
midbrain: 106
motor: 4851
mucleusaccumbens: 53
parietal: 7920
PFC: 5468
preoptic: 247
putamen: 2916
somatosensory: 5433
striatum: 98
temporal: 2129
thalamus: 4469
V1: 5358

total mismatched celltype annotations: 68,134
since the barcodes/numi/ngene all match between the two, still does not address the discrepency between the metadata and raw data onn the nnemo repository

```{r}


```

```{r}

a = 'LGE'

print('Cell barcodes...')
ori_meta = filter(metadata, area == a)$cell.name
ucsc_meta = filter(ucsc_metadata, area == a)$cell_name
length(ori_meta)
length(ucsc_meta)
index = match(ori_meta, ucsc_meta)
table(ori_meta == ucsc_meta[index])

print('Number of genes...')
ori_meta = filter(metadata, area == a)$ngene
ucsc_meta = filter(ucsc_metadata, area == a)$ngene
table(ori_meta == ucsc_meta[index])

print('Number of umi...')
ori_meta = filter(metadata, area == a)$numi
ucsc_meta = filter(ucsc_metadata, area == a)$numi
table(ori_meta == ucsc_meta[index])

print('Cell type annotations...')
ori_meta = filter(metadata, area == a)$cell.type
ucsc_meta = filter(ucsc_metadata, area == a)$cell_type
ori_meta[1:10]
ucsc_meta[1:10]
table(ori_meta == ucsc_meta[index])

```

```{r}
not_match = which(!ori_meta == ucsc_meta[index])
length(not_match)
ori_meta[not_match][1:10]
ucsc_meta[index][not_match][1:10]


table(ori_meta[not_match])
table(ucsc_meta[index][not_match])


temp_df = data.frame(publication_annotations = ori_meta[not_match], ucsc_annotations = ucsc_meta[index][not_match] )
temp_df = temp_df %>% group_by(publication_annotations, ucsc_annotations) %>% summarize(value = length(ucsc_annotations) )

ggplot(temp_df, aes(x = publication_annotations, y = value, fill = ucsc_annotations)) +
  geom_bar(position='fill', stat="identity") + ylab("% change to UCSC annotations") + 
  ggtitle(sprintf('%i cells in %s with different celltype annotations',length(not_match),a)) +
  scale_x_discrete(guide = guide_axis(n.dodge = 2))

ggplot(temp_df, aes(x = publication_annotations, y = value, fill = ucsc_annotations)) +
  geom_bar(position='stack', stat="identity") + ylab("# change to UCSC annotations") + 
  ggtitle(sprintf('%i cells in %s with different celltype annotations',length(not_match),a)) +
  scale_x_discrete(guide = guide_axis(n.dodge = 2))

temp_df
```


The ucsc metadata has a different number of cells for the thalamus region

```{r}
length(filter(metadata, area == 'thalamus')$cell.name)
length(filter(ucsc_metadata, area == 'thalamus')$cell_name)
table(filter(ucsc_metadata, area == 'thalamus')$cell_name %in% filter(metadata, area == 'thalamus')$cell.name)

#Filter for the shared barcodes and then reorder to get the same order between the two
temp_ori_thalamus = filter(metadata, area == 'thalamus')
temp_ucsc_thalamus = filter(ucsc_metadata, area == 'thalamus')

temp_ori_thalamus = temp_ori_thalamus[temp_ori_thalamus$cell.name %in% temp_ucsc_thalamus$cell_name,  ]
index = match(temp_ucsc_thalamus$cell_name,temp_ori_thalamus$cell.name )
temp_ori_thalamus = temp_ori_thalamus[index, ]

table(temp_ori_thalamus$cell.name == temp_ucsc_thalamus$cell_name )


table(temp_ori_thalamus$ngene == temp_ucsc_thalamus$ngene )
table(temp_ori_thalamus$numi == temp_ucsc_thalamus$numi )
table(temp_ori_thalamus$cell.type == temp_ucsc_thalamus$cell_type )
```

what are the missing cells

```{r}
not_present_index = which(!filter(metadata, area == 'thalamus')$cell.name %in% filter(ucsc_metadata, area == 'thalamus')$cell_name )
table(filter(metadata, area == 'thalamus')$individual[not_present_index])
table(filter(metadata, area == 'thalamus')$area[not_present_index])

```






There's also a second metadata file for the neocortex samples
Not quite sure what the difference is between the two
But there's no associated numi or feature counts for this metadata, so can't use it. Just cause the same barcodes are present doesnt mean it's the same cell, there are some mismatches

```{r}


meta_neocortex_file = '/home/werner/projects/meta_qc_organoid/datasets/arealization_fetal_kriegstein/areal_dataset_neocortex_metadata.csv'

metadata_neo_cortex = read.csv(meta_neocortex_file, header = TRUE)
metadata_neo_cortex
```
```{r}
table(metadata_neo_cortex$Age)
table(metadata_neo_cortex$Individual)
table(metadata_neo_cortex$Area)
```

and the ucsc neocortex metatdata

```{r}

ucsc_neocortex_file = '/home/werner/projects/meta_qc_organoid/datasets/arealization_fetal_kriegstein/ucsc_neocortex_metadata.tsv'
ucsc_metadata_neo_cortex = read.table(ucsc_neocortex_file, sep = '\t', header = T)
ucsc_metadata_neo_cortex

table(ucsc_metadata_neo_cortex$Individual)
table(ucsc_metadata_neo_cortex$Area)
```



Going to go through each dataset that has metadata, filter for the cells that have metadata, add the cell type identity and other metadata, save that seurat object

put them all in a single seurat object at the end of everything
```{r}
individuals = names(table(metadata$individual))
individuals
individuals[2]
table(filter(metadata, individual == individuals[2])$area)


```

GW14
somatosensory = somato
V1 = occipital

GW17
somatosensory = somato

GW18
Nothing in GW18 matches

GW18_2
Has the correct metadata. Maybe the GW18 was a bust and the _2 is a second run?

motor, parietal, somatosensory, and temporal, the metadata contains barcodes for the motor and motorVZ directories, temporal and TempVZ directories and so on
represents the different lamina, filter the metadata with the lamina column, all or VZ

somatosensory = somato
thalamus = ventral_thalamus
V1 metadata is just V1VZ data


GW_19
motor = M1_all and M1_CP, metadata for both those directories
parietal = Parietal
somatosensory = S1
temporal = Temp_all
PFC = PFC_all and PFC_CP
V1 = V1_all and V1_CP
thalamus is missing

GW19_2
motor = Motor
parietal = Parietal
somatosensory = SS
temporal = Temporal
thalamus = Thalamus


GW20
somatosensory = somata
thalamus metadata has data for the ventral, dorsal, and caudal thalamus directories

GW20_31
hypothalamus = hypo

GW20_34
hypothalamus = hypo
 PFC, thalamus, parietal metadata all have data for multiple raw data directories
 
GW22

only V1 matches, have no clue what the somato1, somato2 or PFC1, PFC2 , etc, datasets are, metadata does not match


GW22T

nothing matches

GW25

thalamus metadata is for the thalamus, anteriorthal, and pulminarthal raw data directories

Parietal metadata is for parietal, ParDL, ParMZ, ParOSVZ, ParUL, ParVZ raw data directories
PFC metadata is for PFC, PFCDL, PFCMZ, PFCUL, PFCVZ_SVZ raw data directories


```{r}
individual_index = 3
current_individual = individuals[individual_index]

regions = names(table(filter(metadata, individual == current_individual)$area))
region_index = 11

current_region = regions[region_index]


current_individual
current_region

current_dataset = paste(current_individual, current_region, sep = '_')
current_dataset
```





Need to unzip and untar each directory
Plug in directory here

```{r}
#Renaming current_dataset for when the file names and metadata don't agree in spelling

current_dataset = 'GW22_motor2'

```



```{r}
#raw_data_file = sprintf('/data/jwerner/data/nemo/u01-devhu-kriegstein-scell-10xv2/counts/%s/GRCh38/GRCh38/', current_dataset)

raw_data_file = sprintf('/data/jwerner/data/nemo/u01-devhu-kriegstein-scell-10xv2/counts/%s/%s/', current_dataset, current_dataset)
#raw_data_file 

#raw_data_file = sprintf('/data/jwerner/data/nemo/u01-devhu-kriegstein-scell-10xv2/counts/%s/filtered_feature_bc_matrix/filtered_feature_bc_matrix/', current_dataset)

raw_data_file
data_exp = Read10X(raw_data_file)
data_exp = CreateSeuratObject(counts = data_exp)
data_exp
```
```{r}
data_exp[[]]

```




Grab the metadata for that dataset

```{r}

dataset_metadata = filter(metadata, individual == current_individual & area == current_region)

#dataset_metadata = filter(metadata, individual =='GW18_2' & area == current_region)

#If working with the ____VZ or ___CP datasets
#dataset_metadata = filter(metadata, individual == current_individual & area == 'thalamus' & lamina == 'VZSVZ')
#dataset_metadata = filter(metadata, individual == current_individual & area == 'V1' & lamina == 'CP')
dataset_metadata

table(dataset_metadata$lamina)
```



For thalamus metadata, add the sub region to the area column
```{r}
dataset_metadata$area = sapply(strsplit(dataset_metadata$cell.name, split = '_', fixed = T), '[[', 2)
table(dataset_metadata$area )

gw20_thal_metadata  = dataset_metadata


```


Just get the sub thalamic regions

```{r}
dataset_metadata = filter(gw20_thal_metadata, area == 'pulminarthal')

```



Check the metadata matches up to the barcodes present
The datasets with the lamina info in their barcodes, the barcodes are inn the fourth field instead of the third

```{r}
#Get the barcodes for the cells that have metadata
cell_labels = dataset_metadata$cell.name
cell_labels[1:10]
cell_barcodes = sapply(strsplit(cell_labels, split = '_', fixed = T), '[[', 3)


#Get the barcodes from the cells present in the raw data
data_barcodes = sapply(strsplit(rownames(data_exp[[]]), split = '-', fixed = T), '[[', 1)

#Add barcodes to metadata
data_exp$barcodes = data_barcodes

table(data_exp$barcodes %in%  cell_barcodes )


#Subset data to the ones that have metadata present
data_exp_subset = subset(data_exp, subset = barcodes %in% cell_barcodes )
data_exp_subset

```



Seems like the subset of barcodes that have metadata are at least in the same order as the cell data
```{r}

table(cell_barcodes == data_exp_subset$barcodes)
table(dataset_metadata$ngene == data_exp_subset$nFeature_RNA)
table(dataset_metadata$numi == data_exp_subset$nCount_RNA)
```

But if i need to reorder the metadata...
match the barcodes
And check if the nCounts and nFeatures are the same, if not, it's not the right metadata
```{r}

index = match(data_exp_subset$barcodes, cell_barcodes)
table(data_exp_subset$barcodes == cell_barcodes[index])

table(data_exp_subset$nCount_RNA == dataset_metadata$numi[index])
table(data_exp_subset$nFeature_RNA == dataset_metadata$ngene[index])
```
Just need to subset the metadata to match

```{r}
#Reorder the metadata to match the raw data
cell_barcodes = cell_barcodes[index]
dataset_metadata = dataset_metadata[index, ]

table(data_exp_subset$barcodes == cell_barcodes)

table(data_exp_subset$nCount_RNA == dataset_metadata$numi)
table(data_exp_subset$nFeature_RNA == dataset_metadata$ngene)

```



```{r}
par(pty = 's')
plot(data_exp_subset$nCount_RNA, dataset_metadata$numi[index], cex = .5, ylab = 'Metadata nCounts_RNA', xlab = 'Raw data nCounts_RNA')
abline(a = 0, b = 1, col = 'red')

par(pty = 's')
plot(data_exp_subset$nFeature_RNA, dataset_metadata$ngene[index], cex = .5, ylab = 'Metadata nFeature_RNA', xlab = 'Raw data nFeature_RNA')
abline(a = 0, b = 1, col = 'red')

```
grab the subset of cells that match the metadata

```{r}
#Reorder the metadata to match the raw data
cell_barcodes = cell_barcodes[index]
dataset_metadata = dataset_metadata[index, ]

#Filter out the cells that don't match
#Check theyre the same cells
table(which(data_exp_subset$nCount_RNA != dataset_metadata$numi) == which(data_exp_subset$nFeature_RNA != dataset_metadata$ngene))

#Get the bad barcodes
bad_cells = cell_barcodes[which(data_exp_subset$nCount_RNA != dataset_metadata$numi)]
bad_cells

#Filter them out
data_exp_subset = data_exp_subset[ , !data_exp_subset$barcodes %in% bad_cells]
dataset_metadata = dataset_metadata[!cell_barcodes %in% bad_cells, ]
cell_barcodes = cell_barcodes[!cell_barcodes %in% bad_cells]


#Doule check everything agrees
table(data_exp_subset$barcodes == cell_barcodes)
table(data_exp_subset$nCount_RNA == dataset_metadata$numi)
table(data_exp_subset$nFeature_RNA == dataset_metadata$ngene)

```






Add the metadata to the seurat object
```{r}

data_exp_subset$age = dataset_metadata$age
data_exp_subset$individual = dataset_metadata$individual
data_exp_subset$structure = dataset_metadata$structure
data_exp_subset$area = dataset_metadata$area
data_exp_subset$clusterv1 = dataset_metadata$clusterv1
data_exp_subset$`clusterv2...final` = dataset_metadata$`clusterv2...final`
data_exp_subset$clusterv2_id = dataset_metadata$clusterv2_id
data_exp_subset$lamina = dataset_metadata$lamina
data_exp_subset$cell.type = dataset_metadata$cell.type

data_exp_subset[[]]


#save the dataset seurat object

save_file = sprintf('/home/werner/projects/meta_qc_organoid/data/seurat_objs/individual/areal_kreigstein_dataset/%s.Rdata',
                    current_dataset)
save_file
save(data_exp_subset, file = save_file)

```





Now check out the datasets with no metadata, is it possible to use them?
GW15, GW16, GW18, GW22, GW22T



```{r}

individual_index = 11
current_individual = individuals[individual_index]

regions = names(table(filter(metadata, individual == current_individual)$area))
region_index = 8
current_region = regions[region_index]
#current_region = 'somato'

current_dataset = paste(current_individual, current_region, sep = '_')
current_dataset


#GW22 and GW22T have 1 and 2 on their directory names
current_dataset = paste0(current_dataset, '1')
current_dataset


raw_data_file = sprintf('/data/jwerner/data/nemo/u01-devhu-kriegstein-scell-10xv2/counts/%s/%s/', current_dataset, current_dataset)
raw_data_file
data_exp_subset = Read10X(raw_data_file)
data_exp_subset = CreateSeuratObject(counts = data_exp_subset)

hist(data_exp_subset$nCount_RNA, main = sprintf('Num_cells: %i',ncol(data_exp_subset)))
hist(data_exp_subset$nFeature_RNA, main = sprintf('Num_cells: %i',ncol(data_exp_subset)))

#add an area metadata column for the 18, 22, and 22T datasets. Cant add one for the GW15 and GW16 datasets
data_exp_subset$area = rep(current_region, ncol(data_exp_subset))
data_exp_subset[[]]

#save the dataset seurat object
save_file = sprintf('/home/werner/projects/meta_qc_organoid/data/seurat_objs/individual/areal_kreigstein_dataset/%s.Rdata',
                    current_dataset)
save_file
save(data_exp_subset, file = save_file)

```

GW15 and GW16, no metadata available
All of 15 is very ppoor quality, in the hundreds of Features per gene, don't use
GW16 s decent

```{r}


GW15_dirs = c('GW15_hippocampus','GW15_hypothalamus','GW15_LGE','GW15_MGE','GW15_parietal','GW15_PFC_CP','GW15_PFC_VZ','GW15_somato','GW15_temporal','GW15_thalamus','GW15_wholePFC')
GW15_areas = c('hippocampus','hypothalamus','LGE','MGE','parietal','PFC_CP','PFC_VZ','somato','temporal','thalamus','wholePFC')
GW16_dirs = c('GW16_hypo','GW16_MGE','GW16_motor','GW16_MotorVZ','GW16_ParietalVZ','GW16_PFC','GW16_PFCVZ','GW16_Somato','GW16_SomatoVZ','GW16_Temporal','GW16_TemporalVZ','GW16_V1','GW16_V1VZ')
GW16_areas = c('hypo','MGE','motor','MotorVZ','ParietalVZ','PFC','PFCVZ','Somato','SomatoVZ','Temporal','TemporalVZ','V1','V1VZ')

i =13

current_dataset = GW16_dirs[i]
current_dataset

raw_data_file = sprintf('/data/jwerner/data/nemo/u01-devhu-kriegstein-scell-10xv2/counts/%s/%s/', current_dataset, current_dataset)
#Random weird file path for GW16_TemporalVZ
#raw_data_file = '/data/jwerner/data/nemo/u01-devhu-kriegstein-scell-10xv2/counts/GW16_TemporalVZ/GW16_TemporalVZ.mex.tar.gz/GW16_TemporalVZ'

raw_data_file
data_exp_subset = Read10X(raw_data_file)
data_exp_subset = CreateSeuratObject(counts = data_exp_subset)

hist(data_exp_subset$nCount_RNA, main = sprintf('Num_cells: %i',ncol(data_exp_subset)))
hist(data_exp_subset$nFeature_RNA, main = sprintf('Num_cells: %i',ncol(data_exp_subset)))

#add an area metadata column for the 18, 22, and 22T datasets. Cant add one for the GW15 and GW16 datasets
data_exp_subset$area = rep(GW16_areas[i], ncol(data_exp_subset))
data_exp_subset[[]]

#save the dataset seurat object
save_file = sprintf('/home/werner/projects/meta_qc_organoid/data/seurat_objs/individual/areal_kreigstein_dataset/%s.Rdata',
                    current_dataset)
save_file
save(data_exp_subset, file = save_file)

```


Combine all the seurat objects now that they have metadata

```{r}
all_dataset_path = '/home/werner/projects/meta_qc_organoid/data/seurat_objs/individual/areal_kreigstein_dataset/'
all_datasets = paste(all_dataset_path, list.files(all_dataset_path), sep = '')

#Get dataset identifiers
dataset_ids = sapply(strsplit(list.files(all_dataset_path), split = '.', fixed = T), '[[', 1)
dataset_ids
```

Merge all the datasets per time point, saving as separate datasets per person really

```{r}

timepoint_index = which(grepl('GW18', dataset_ids) & !grepl('GW18_2', dataset_ids))
dataset_ids[timepoint_index]

```

```{r}
#Load up all the timepoint datasets

for(i in timepoint_index){
  
  load(all_datasets[i])
  assign(dataset_ids[i], data_exp_subset)
}


#Just mmake the vector of all the dataset names
dataset_ids[timepoint_index]

```


```{r}
GW14_merged_dataset = merge(GW14_CGE, y = c(GW14_hypothalamus, GW14_LGE,GW14_MGE,GW14_motor,GW14_occipital,GW14_somato,GW14_striatum,GW14_thalamus),
                            add.cell.ids = dataset_ids[timepoint_index], project = 'GW14' )

GW17_merged_dataset = merge(GW17_motor, y = c(GW17_parietal, GW17_PFC, GW17_somato, GW17_V1), add.cell.ids = dataset_ids[timepoint_index], project = 'GW17')

GW16_merged_dataset = merge(GW16_hypo, y = c(GW16_MGE,GW16_motor,GW16_MotorVZ,GW16_ParietalVZ,GW16_PFC,GW16_PFCVZ,GW16_Somato,GW16_SomatoVZ,GW16_Temporal, GW16_TemporalVZ,GW16_V1,GW16_V1VZ), 
                            add.cell.ids = dataset_ids[timepoint_index], project = 'GW16')

GW18_merged_dataset = merge(GW18_CGE, y = c(GW18_hippocampus, GW18_hypothalamus, GW18_LGE, GW18_MGE, GW18_motor, GW18_parietal, GW18_PFC, GW18_thalamus, GW18_V1), 
                              add.cell.ids = dataset_ids[timepoint_index], project = 'GW18')

GW18_2_merged_dataset = merge(GW18_2_CGE, y = c(GW18_2_claustrum, GW18_2_hippocampus, GW18_2_hypothalamus, GW18_2_LGE, GW18_2_MGE, GW18_2_motor, GW18_2_motorVZ, GW18_2_parietal,
                                                GW18_2_ParVZ, GW18_2_somato, GW18_2_somatoVZ, GW18_2_temporal, GW18_2_TempVZ, GW18_2_V1VZ, GW18_2_ventral_thalamus), 
                              add.cell.ids = dataset_ids[timepoint_index], project = 'GW18_2')

GW19_merged_dataset = merge(GW19_thalamus, y = c(GW19_CGE, GW19_LGE, GW19_hippocampus, GW19_hypothalamus,GW19_M1_all, GW19_M1_CP, GW19_MGE, GW19_Parietal, GW19_PFC_all, GW19_PFC_CP, 
                                                 GW19_S1, GW19_Temp_all, GW19_V1_all, GW19_V1_CP), 
                            add.cell.ids = dataset_ids[timepoint_index], project = 'GW19')

GW19_2_merged_dataset = merge(GW19_2_Thalamus, y = c(GW19_2_CGE, GW19_2_Motor,  GW19_2_Parietal, GW19_2_PFC, GW19_2_SS, GW19_2_Temporal,  GW19_2_V1), 
                            add.cell.ids = dataset_ids[timepoint_index], project = 'GW19_2')

GW20_merged_dataset = merge(GW20_caudalthalamus, y = c( GW20_claustrum, GW20_dorsalthalamus, GW20_hypothalamus, GW20_motor, GW20_preoptic, GW20_putamen, GW20_somato,
                                                        GW20_ventral_thalamus, GW20CGE, GW20LGE, GW20MGE, GW20PFC, GW20V1), 
                            add.cell.ids = dataset_ids[timepoint_index], project = 'GW20')

GW20_31_merged_dataset = merge(GW20_31_cerebellum, y = c(GW20_31_CGE, GW20_31_hypo, GW20_31_LGE, GW20_31_MGE, GW20_31_parietal, GW20_31_PFC, GW20_31_temporal, GW20_31_V1), 
                               add.cell.ids = dataset_ids[timepoint_index], project = 'GW20_31')

GW20_34_merged_dataset = merge(GW20_34_caudate, y = c(GW20_34_CGE, GW20_34_dorsalthalamus, GW20_34_hypo, GW20_34_LGE, GW20_34_MGE, GW20_34_motor, GW20_34_nucleusaccumbens, 
                                                      GW20_34_parietal, GW20_34_ParVZ, GW20_34_PFC, GW20_34_PFCVZ, GW20_34_putamen, GW20_34_V1, GW20_34_ventralthalamus), 
                               add.cell.ids = dataset_ids[timepoint_index], project = 'GW20_34')

#Just the one dataset for GW22
GW22_merged_dataset = merge(GW22_V1, y = c(GW22_motor1, GW22_motor2, GW22_parietal1, GW22_parietal2, GW22_PFC1, GW22_PFC2, GW22_somato1, GW22_somato2),
                            add.cell.ids = dataset_ids[timepoint_index], project = 'GW22')


GW22T_merged_dataset = merge(GW22T_CGE, y = c(GW22T_hypo1, GW22T_hypo2, GW22T_LGE1, GW22T_LGE2, GW22T_motor1, GW22T_motor2, GW22T_parietal1, GW22T_parietal2, GW22T_PFC1, GW22T_PFC2, 
                                              GW22T_somato1, GW22T_somato2, GW22T_thalamus1, GW22T_thalamus2),
                             add.cell.ids = dataset_ids[timepoint_index], project = 'GW22T')

GW25_merged_dataset = merge(GW25_3V_hypo, y = c(GW25_anteriorthal, GW25_caudate, GW25_CGE, GW25_cingulate, GW25_insula, GW25_midbrain, GW25_motor, GW25_ParDL, GW25_parietal, 
                                                GW25_ParMZ, GW25_ParOSVZ, GW25_ParUL, GW25_ParVZ, GW25_PFC, GW25_PFCDL, GW25_PFCMZ, GW25_PFCUL, GW25_PFCVZ_SVZ, GW25_pulminarthal, 
                                                GW25_somato, GW25_temporal, GW25_thalamus), 
                            add.cell.ids = dataset_ids[timepoint_index], project = 'GW25')

```

```{r}

rm(GW25_anteriorthal, GW25_caudate, GW25_CGE, GW25_cingulate, GW25_insula, GW25_midbrain, GW25_motor, GW25_ParDL, GW25_parietal, 
                                                GW25_ParMZ, GW25_ParOSVZ, GW25_ParUL, GW25_ParVZ, GW25_PFC, GW25_PFCDL, GW25_PFCMZ, GW25_PFCUL, GW25_PFCVZ_SVZ, GW25_pulminarthal, 
                                                GW25_somato, GW25_temporal, GW25_thalamus)

```


```{r}
GW14_merged_dataset[[]]
GW16_merged_dataset[[]]
GW17_merged_dataset[[]]
GW18_merged_dataset[[]]
GW18_2_merged_dataset[[]]
GW19_merged_dataset[[]]
GW19_2_merged_dataset[[]]
GW20_merged_dataset[[]]
GW20_31_merged_dataset[[]]
GW20_34_merged_dataset[[]]
GW22_merged_dataset[[]]
GW22T_merged_dataset[[]]
GW25_merged_dataset[[]]

```





```{r}

hist(GW18_merged_dataset$nCount_RNA)
hist(GW18_merged_dataset$nFeature_RNA)

GW18_merged_dataset@assays$RNA@counts[1:10, 1:10]
```


```{r}
save(GW14_merged_dataset, file = '/home/werner/projects/meta_qc_organoid/data/seurat_objs/individual/areal_kreigstein_dataset/merged_with_metadata/GW14.Rdata')
save(GW16_merged_dataset, file = '/home/werner/projects/meta_qc_organoid/data/seurat_objs/individual/areal_kreigstein_dataset/merged_with_metadata/GW16.Rdata')
save(GW17_merged_dataset, file = '/home/werner/projects/meta_qc_organoid/data/seurat_objs/individual/areal_kreigstein_dataset/merged_with_metadata/GW17.Rdata')
save(GW18_merged_dataset, file = '/home/werner/projects/meta_qc_organoid/data/seurat_objs/individual/areal_kreigstein_dataset/merged_with_metadata/GW18.Rdata')
save(GW18_2_merged_dataset, file = '/home/werner/projects/meta_qc_organoid/data/seurat_objs/individual/areal_kreigstein_dataset/merged_with_metadata/GW18_2.Rdata')
save(GW19_merged_dataset, file = '/home/werner/projects/meta_qc_organoid/data/seurat_objs/individual/areal_kreigstein_dataset/merged_with_metadata/GW19.Rdata')
save(GW19_2_merged_dataset, file = '/home/werner/projects/meta_qc_organoid/data/seurat_objs/individual/areal_kreigstein_dataset/merged_with_metadata/GW19_2.Rdata')
save(GW20_merged_dataset, file = '/home/werner/projects/meta_qc_organoid/data/seurat_objs/individual/areal_kreigstein_dataset/merged_with_metadata/GW20.Rdata')
save(GW20_31_merged_dataset, file = '/home/werner/projects/meta_qc_organoid/data/seurat_objs/individual/areal_kreigstein_dataset/merged_with_metadata/GW20_31.Rdata')
save(GW20_34_merged_dataset, file = '/home/werner/projects/meta_qc_organoid/data/seurat_objs/individual/areal_kreigstein_dataset/merged_with_metadata/GW20_34.Rdata')
save(GW22_merged_dataset, file = '/home/werner/projects/meta_qc_organoid/data/seurat_objs/individual/areal_kreigstein_dataset/merged_with_metadata/GW22.Rdata')
save(GW22T_merged_dataset, file = '/home/werner/projects/meta_qc_organoid/data/seurat_objs/individual/areal_kreigstein_dataset/merged_with_metadata/GW22T.Rdata')
save(GW25_merged_dataset, file = '/home/werner/projects/meta_qc_organoid/data/seurat_objs/individual/areal_kreigstein_dataset/merged_with_metadata/GW25.Rdata')
```






################################################################
Load up the datasets

cpm normalizations and dim reductions from Seurat

################################################################


```{r}
load('/home/werner/projects/meta_qc_organoid/data/seurat_objs/individual/areal_kreigstein_dataset/merged_with_metadata/GW14.Rdata')
load('/home/werner/projects/meta_qc_organoid/data/seurat_objs/individual/areal_kreigstein_dataset/merged_with_metadata/GW16.Rdata')
load('/home/werner/projects/meta_qc_organoid/data/seurat_objs/individual/areal_kreigstein_dataset/merged_with_metadata/GW17.Rdata')
load('/home/werner/projects/meta_qc_organoid/data/seurat_objs/individual/areal_kreigstein_dataset/merged_with_metadata/GW18.Rdata')
load('/home/werner/projects/meta_qc_organoid/data/seurat_objs/individual/areal_kreigstein_dataset/merged_with_metadata/GW18_2.Rdata')
load('/home/werner/projects/meta_qc_organoid/data/seurat_objs/individual/areal_kreigstein_dataset/merged_with_metadata/GW19.Rdata')
load('/home/werner/projects/meta_qc_organoid/data/seurat_objs/individual/areal_kreigstein_dataset/merged_with_metadata/GW19_2.Rdata')
load('/home/werner/projects/meta_qc_organoid/data/seurat_objs/individual/areal_kreigstein_dataset/merged_with_metadata/GW20.Rdata')
load('/home/werner/projects/meta_qc_organoid/data/seurat_objs/individual/areal_kreigstein_dataset/merged_with_metadata/GW20_31.Rdata')
load('/home/werner/projects/meta_qc_organoid/data/seurat_objs/individual/areal_kreigstein_dataset/merged_with_metadata/GW20_34.Rdata')
load('/home/werner/projects/meta_qc_organoid/data/seurat_objs/individual/areal_kreigstein_dataset/merged_with_metadata/GW22.Rdata')
load('/home/werner/projects/meta_qc_organoid/data/seurat_objs/individual/areal_kreigstein_dataset/merged_with_metadata/GW22T.Rdata')
load('/home/werner/projects/meta_qc_organoid/data/seurat_objs/individual/areal_kreigstein_dataset/merged_with_metadata/GW25.Rdata')
```


```{r}

dataset_paths = c('/home/werner/projects/meta_qc_organoid/data/seurat_objs/individual/areal_kreigstein_dataset/merged_with_metadata/GW14.Rdata',
                  '/home/werner/projects/meta_qc_organoid/data/seurat_objs/individual/areal_kreigstein_dataset/merged_with_metadata/GW16.Rdata',
                  '/home/werner/projects/meta_qc_organoid/data/seurat_objs/individual/areal_kreigstein_dataset/merged_with_metadata/GW17.Rdata',
                  '/home/werner/projects/meta_qc_organoid/data/seurat_objs/individual/areal_kreigstein_dataset/merged_with_metadata/GW18.Rdata',
                  '/home/werner/projects/meta_qc_organoid/data/seurat_objs/individual/areal_kreigstein_dataset/merged_with_metadata/GW18_2.Rdata',
                  '/home/werner/projects/meta_qc_organoid/data/seurat_objs/individual/areal_kreigstein_dataset/merged_with_metadata/GW19.Rdata',
                  '/home/werner/projects/meta_qc_organoid/data/seurat_objs/individual/areal_kreigstein_dataset/merged_with_metadata/GW19_2.Rdata',
                  '/home/werner/projects/meta_qc_organoid/data/seurat_objs/individual/areal_kreigstein_dataset/merged_with_metadata/GW20.Rdata',
                  '/home/werner/projects/meta_qc_organoid/data/seurat_objs/individual/areal_kreigstein_dataset/merged_with_metadata/GW20_31.Rdata',
                  '/home/werner/projects/meta_qc_organoid/data/seurat_objs/individual/areal_kreigstein_dataset/merged_with_metadata/GW20_34.Rdata',
                  '/home/werner/projects/meta_qc_organoid/data/seurat_objs/individual/areal_kreigstein_dataset/merged_with_metadata/GW22.Rdata',
                  '/home/werner/projects/meta_qc_organoid/data/seurat_objs/individual/areal_kreigstein_dataset/merged_with_metadata/GW22T.Rdata',
                  '/home/werner/projects/meta_qc_organoid/data/seurat_objs/individual/areal_kreigstein_dataset/merged_with_metadata/GW25.Rdata')

```



```{r}
rm(GW14_merged_dataset, GW17_merged_dataset, GW18_2_merged_dataset, GW19_merged_dataset, GW20_merged_dataset, poliodakis_dataset, fan_fu_dataset, GW20_31_merged_dataset, 
   GW20_34_merged_dataset, GW22_merged_dataset, GW25_merged_dataset)

```


```{r}
dataset_names = c('GW14_merged_dataset','GW16_merged_dataset','GW17_merged_dataset','GW18_merged_dataset','GW18_2_merged_dataset','GW19_merged_dataset','GW19_2_merged_dataset','GW20_merged_dataset','GW20_31_merged_dataset',
                  'GW20_34_merged_dataset','GW22_merged_dataset','GW22T_merged_dataset', 'GW25_merged_dataset')
dataset_filenames = c('GW14', 'GW16','GW17', 'GW18', 'GW18_2','GW19','GW19_2','GW20','GW20_31','GW20_34','GW22','GW22T','GW25')

```

```{r}

for(i in 3:length(dataset_paths)){

  
  i = 12
  x = load(dataset_paths[i])
  dataset = get(x)
  
  
  dataset[["percent.mt"]] <- PercentageFeatureSet(dataset, pattern = "^MT-")
  dataset = subset(dataset, subset = percent.mt < 50 & nFeature_RNA >=200 & nFeature_RNA <=6000)
  dataset = NormalizeData(dataset, verbose = F, normalization.method = 'RC', scale.factor = 1e6)
  dataset = FindVariableFeatures(dataset, selection.method = "vst", nfeatures = 2000, verbose = F)
  all.genes <- rownames(dataset)
  dataset = ScaleData(dataset , features = all.genes, verbose = F)
  dataset = RunPCA(dataset, features = VariableFeatures(object = dataset), verbose = F)
  dataset = RunUMAP(dataset, dims = 1:20, verbose = F)
  
  
  print(DimPlot(dataset,reduction = 'umap', group.by = 'area'))
  print(DimPlot(dataset,reduction = 'umap', group.by = 'cell.type'))
  
  
  file_path = sprintf('/home/werner/projects/meta_qc_organoid/data/seurat_objs/individual/areal_kreigstein_dataset/merged_with_metadata/%s.Rdata', dataset_filenames[i])
  dataset_name = dataset_names[i]
  assign(dataset_name, dataset)
  save(list = c(dataset_name), file = file_path)
  
  rm(x,dataset,GW14_merged_dataset, GW16_merged_dataset, GW17_merged_dataset, GW18_merged_dataset, GW18_2_merged_dataset, GW19_merged_dataset, GW19_2_merged_dataset, GW20_merged_dataset, 
     GW20_31_merged_dataset, GW20_34_merged_dataset, 
     GW22_merged_dataset, GW22T_merged_dataset, GW25_merged_dataset)
  gc()

}

```

```{r}

load('/home/werner/projects/meta_qc_organoid/data/seurat_objs/individual/areal_kreigstein_dataset/merged_with_metadata/GW22T.Rdata')


GW22T_merged_dataset
```

```{r}

DimPlot(GW22T_merged_dataset, reduction = 'umap', group.by = 'area')

```


Check what the adult and fetal metamarkers look like here

Human BICCN metamarkers
Stephan got the top 1000 markers at each level of classification

```{r}
human_class_path = '/data/fischer/biccn_markers/human/class_markers_top1000.csv.gz'
human_subclass_path = '/data/fischer/biccn_markers/human/subclass_markers_top1000.csv.gz'
human_type_path = '/data/fischer/biccn_markers/human/type_markers_top1000.csv.gz'

human_class_markers = fread(human_class_path)

human_subclass_markers = fread(human_subclass_path)

human_type_markers = fread(human_type_path)
```


Get the metamarkers
```{r}
markers = list(
    poliodakis = read_markers("/home/werner/projects/meta_qc_organoid/code/poliodakis_fetal_class_markers.csv.gz"),
    fan_fu = read_markers("/home/werner/projects/meta_qc_organoid/code/fan_fu_fetal_class_markers.csv.gz")
)


```


```{r}

fetal_meta_markers = make_meta_markers(markers, detailed_stats = TRUE)

```

```{r}

top_adult_mMarkers = filter(human_class_markers, rank <= 30 )
#Run MetaMarker functions
class_test = score_cells(GW25_merged_dataset@assays$RNA@data, top_adult_mMarkers)
class_enrichment_test = compute_marker_enrichment(class_test)
class_pred_test = assign_cells(class_test)

GW25_merged_dataset[['GABAergic_enrichments']] = class_enrichment_test['all|GABAergic' , ]
GW25_merged_dataset[['Glutamatergic_enrichments']] = class_enrichment_test['all|Glutamatergic' , ]
GW25_merged_dataset[['Non_neuronal_enrichments']] = class_enrichment_test['all|Non-neuronal' , ]
GW25_merged_dataset[['predicted_class']] = class_pred_test$predicted


GW25_merged_dataset[['GABAergic_scores']] = class_test['all|GABAergic' , ]
GW25_merged_dataset[['Glutamatergic_scores']] = class_test['all|Glutamatergic' , ]
GW25_merged_dataset[['Non_neuronal_scores']] = class_test['all|Non-neuronal' , ]


DimPlot(GW25_merged_dataset,reduction = 'umap', group.by = 'area')
DimPlot(GW25_merged_dataset,reduction = 'umap', group.by = 'cell.type')
FeaturePlot(GW25_merged_dataset, features = c('GABAergic_enrichments'))
FeaturePlot(GW25_merged_dataset, features = c('Glutamatergic_enrichments'))
FeaturePlot(GW25_merged_dataset, features = c('Non_neuronal_enrichments'))
```


```{r}

top_fetal_mMarkers = filter(fetal_meta_markers, rank <= 30 )
#Run MetaMarker functions
class_test = score_cells(GW25_merged_dataset@assays$RNA@data, top_fetal_mMarkers)
class_enrichment_test = compute_marker_enrichment(class_test)
class_pred_test = assign_cells(class_test)

GW25_merged_dataset[['GABAergic_enrichments']] = class_enrichment_test['all|GABAergic' , ]
GW25_merged_dataset[['Glutamatergic_enrichments']] = class_enrichment_test['all|Glutamatergic' , ]
GW25_merged_dataset[['Neural_prog_enrichments']] = class_enrichment_test['all|Neural_Progenitor' , ]
GW25_merged_dataset[['Non_neuronal_enrichments']] = class_enrichment_test['all|Non-neuronal' , ]
GW25_merged_dataset[['predicted_class']] = class_pred_test$predicted

GW25_merged_dataset$MM_predicted_class =  class_pred_test$predicted

GW25_merged_dataset[['GABAergic_scores']] = class_test['all|GABAergic' , ]
GW25_merged_dataset[['Glutamatergic_scores']] = class_test['all|Glutamatergic' , ]
GW25_merged_dataset[['Neural_prog_enrichments']] = class_test['all|Neural_Progenitor' , ]
GW25_merged_dataset[['Non_neuronal_scores']] = class_test['all|Non-neuronal' , ]


DimPlot(GW25_merged_dataset,reduction = 'umap', group.by = 'area')
DimPlot(GW25_merged_dataset,reduction = 'umap', group.by = 'cell.type')
DimPlot(GW25_merged_dataset, reduction = 'umap', group.by = 'MM_predicted_class')
FeaturePlot(GW25_merged_dataset, features = c('GABAergic_enrichments'))
FeaturePlot(GW25_merged_dataset, features = c('Glutamatergic_enrichments'))
FeaturePlot(GW25_merged_dataset, features = c('Neural_prog_enrichments'))
FeaturePlot(GW25_merged_dataset, features = c('Non_neuronal_enrichments'))
```








